### 2022-11-06 reports/report1/main.tex
Create main.tex

\documentclass[12pt]{article}
\usepackage[left=1in, right=1in, top=1in, bottom=1in]{geometry}
\usepackage{graphicx}
\usepackage{url}
\usepackage{cite}
\usepackage{float}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{hyperref}
\hypersetup{
    colorlinks,
    citecolor=black,
    filecolor=black,
    linkcolor=black,
    urlcolor=black
}

\renewcommand{\familydefault}{\sfdefault}

\begin{document}
\begin{titlepage}
    \begin{center}
        \LARGE
        \textbf{ELE 401 - GRADUATION PROJECT I}

        \Large
        \textbf{FIRST INTERIM REPORT}

        \vspace{70pt}

        \textit{
            Hacettepe University \\
            Department of Electrical and Electronics Engineering
        }
    \end{center}

    \vspace{90pt}

    \large

    \textbf{Project Title:} Sensor Fusion in the Cloud \\

    \textbf{Project Group Members:} Ertuğrul Tiyek, Ahmet Yusuf Şirin

    \vspace{30pt}

    \textbf{Project Supervisor:} Asst. Prof. Dr. İsmail Uyanık \\

    \textbf{Submission Date:} 05.11.2022

    \vspace{\fill}

    \begin{center}
        \textit{FALL 2022-2023}
    \end{center}
\end{titlepage}

\clearpage

\tableofcontents
\listoffigures
\listoftables

\clearpage

\section{INTRODUCTION}

One of the most important aim of the modern technology is the sensing the environment. Living beings have sense organs to know the environment. They have brains to make fusion the sensor data and augment the environmental information taken from sense organs. For example, human standing in balance, we can say that they use the data coming from their eyes and ears together.

People are studying to make this sensing and augmenting artificially, too. They use computers to achieve this. But computers cannot sense the environment directly. They need some devices called sensors. There are all kinds of sensors to measure different environmental quantities. For same quantity, there are different sensors using different technologies, have different precision and accuracy levels. Each sensor has its advantages and disadvantages. Even sensors that are serving a different purpose can help to augment same information other than their main. This is known as "sensor fusion"~\cite{enwiki:1115352853}. Just like humans, robots are use camera and inertial sensors to stand in balance.

In our project, sensor fusion handles extracting depth information. Sensor fusion can be implemented using specific algorithms like EKF, CNN etc. according to quantity that will be calculated, number and types of sensors used, accuracy and performance needed. We are planning to develop a sensor fusion algorithm with the help of living organisms. As we mentioned before, living beings perform sensor fusion perfectly. We will be examining the zebra fish and how it processes data of its sense organs to produce depth information. Then we will try to develop similar algorithm or improve our algorithm with data taken from zebra fish.

Depth information can be used for many applications like automatic driven vehicles, UAVs, Augmented Reality devices and 3D modeling of areas. Our concept in this project will be autonomous controlled SLAM Robot. Simultaneous localization and mapping (SLAM) is the computational problem of constructing or updating a map of an unknown environment while simultaneously keeping track of an agent's location within it~\cite{enwiki:1120084976}. We use a robot to move the system.

Another concept we aim to use is cloud computing. Mobile computing devices may not match the performance requirements in some cases. To achieve more accurate depth information, complex control algorithms or neural networks will be used in our project which needs high computing power. As an alternative to the mobile computing devices, we may use cloud computing.

\section{PROJECT DESCRIPTION}

This project aims to construct an unmanned ground vehicle that senses the environment and decides its movements. As mentioned in the "Introduction" section, the sensor fusion and mapping algorithms take a lot of space in this project. In addition to that, there are other parts that need to be implemented like robot controls, data acquisition from sensors and cameras, decision making algorithms etc.

    \subsubsection*{A. Hardware Architecture of the Project}

    The hardware component choices are explained with consideration of advantages and disadvantages in the section 3.2 Design Constraints. Here we have our main components and specifications of them:

    \begin{enumerate}
        \item \textbf{Robot – Unitree Go1 Edu} \\
        This robot is a ready-to-go quadruped robot with built-in 5 Sets of Fish-eye Stereo Depth Cameras, Ai Post-processing, 3 Sets of Hypersonic Sensors. There are three Nvidia Jetson Nano, one Raspberry Pi 4, one STM32 Microcontroller devices inside of the Unitree Go1 Edu. These devices located in the robot to perform certain tasks. STM32 Microcontroller drives joint motors of the robot. Raspberry Pi 4 is used for both communicate with user and send commands to STM32. Two of Nvidia Jetson Nano 4GB kits and one Nvidia Jetson Nano 2 GB development kit are used for getting data from camera and sensors, processing raw data to extract dept information of stereo cameras.

        \begin{figure}[H]
            \centering
            \includegraphics[width=0.9\textwidth]{Go1CommFram_E.png}    
            \caption{Infrastructure of the Unitree Go1 Robot}
        \end{figure}

        The manufacturer provides two Software Development Kits. "unitree\_legged\_sdk” for Controlling the robot, “unitree\_camera\_sdk” for getting camera data and calculating depth information.

        \item \textbf{Stereo Cameras – Unitree Camera } \\
        There are 5 sets of fish-eye stereo depth cameras located on different important locations of the robot. The known features are:

        \begin{itemize}
            \item Lens Angle $\approx$ 150x170° 
            \item HD (1280x720) , 30fps video streaming 
            \item Raw Frame 
            \item Rectangular Frame 
            \item Depth Frame 
            \item Point Cloud Image 
            \item Provided SDK 
        \end{itemize}

        \item \textbf{3D LIDAR – Velodyne 80-VLP-16-A~\cite{VelodyneLiDAR}} \\
        The VLP-16 has a range of 100m, and the sensor's low power consumption (~8W), light weight (830 grams), compact footprint (~Ø103mm x 72mm), and dual return capability make it ideal for UAVs and other mobile applications.

        Velodyne’s LiDAR Puck supports 16 channels, ~300,000 points/sec, a 360° horizontal field of view and a 30° vertical field of view, with +/- 15° up and down. The Velodyne LiDAR Puck does not have visible rotating parts, making it highly resilient in challenging environments (Rated IP67).

        Automotive Mapping UAV Security Robotics Automation

        The module provides up to 0.3 million points/second over 100Mbps ethernet cable and UDP interface. UDP packages contain distances, calibrated reflectivity information, rotation angles, synchronized time stamps (microseconds resolution).

        \item \textbf{Computation Unit – Nvidia Jetson Nano 4GB} \\
        Unitree Go1 Edu has three Nvidia Jetson Nano Development kits inside it to perform tasks like point cloud, human skeleton tracking etc. We don’t use most of the features robot has, so we decided on using one of the Nvidia Jetson Nano 4GB models in the robot with reducing its 
        workload.

        Here is some specifications of Nvidia Jetson Nano 4GB~\cite{JatsonNano}:

        \begin{itemize}
            \item GPU		:	128-core Maxwell 
            \item CPU		:	Quad-core ARM A57 @ 1.43 GHz 
            \item Memory	:	4 GB 64-bit LPDDR4 25.6 GB/s 
            \item Storage	:	microSD (not included) 
            \item Camera	:	2x MIPI CSI-2 DPHY lanes 
            \item Connectivity	:	Gigabit Ethernet, M.2 Key E 
            \item Display	:	HDMI and display port 
            \item USB		:	4x USB 3.0, USB 2.0 Micro-B 
            \item Others		:	GPIO, I2C, I2S, SPI, UART 
        \end{itemize}

        As mentioned before, one of the aims of this project is performing heavy computation on cloud. The point of implementing the algorithm on a mobile platform is to keep things simple in the first step. That provides us both simple implementation and ability to make a comparison between mobile computing and cloud computing in terms of performance, accuracy and cost.

    \end{enumerate}

    \subsubsection*{B. Software Architecture of the Project}

    Laying on the keep it simple stupid (KISS) principle, we separate our project into smaller and independent parts of development.

    \begin{enumerate}
        \item \textbf{Controlling the Robot Movement and Getting Sensor Data Using Software} \\
        There is a SDK provided by Unitree to control the robot with software called unitree\_legged\_sdk. We used the python programming language with unitree\_legged\_sdk. The details about choice of programming language are explained in the section 3.2 design constraints.

        Connection between robot and controller computer established via UDP protocol. There are two options that can provide UDP connection. One is the ethernet port of switch that connects all devices in the robot and the other is Wi-Fi access point that connected to the Raspberry Pi 4 in the robot. All devices are using a network mask 255.255.255.0, so it is important to use a static IP on the user computer.

        \begin{figure}[H]
            \centering
            \includegraphics[width=0.8\textwidth]{UserRobotConnection.png}
            \caption{User Robot Connection: Network Mask Settings}
        \end{figure}

        \begin{figure}[H]
            \centering
            \includegraphics[width=0.8\textwidth]{UserRPiConn.png}
            \caption{User Robot Connection: IP of Raspberry Pi}
        \end{figure}

        Unitree\_legged\_sdk allows both low-level and high-level controls. Low-level control allows user to adjust each joint manually. High-level control provides user to control robot with high-level commands like walk, turn, jump etc. We will be using high-level control for this project.

        SDK has udp class that provides us an object constructs a UDP connection with Raspberry Pi of Robot. UDP object has functions to receive, send, decode and encode udp packages.

        Another class included in the SDK is HighState. HighState objects store state information of the robot. It is updated using udp.receive(HighState) function. HighState objects includes IMU measurement information, odometer information, 2D speed information, yaw speed information, position information, body height information etc.

        To send the commands to the robot, there is a class called CMD in the SDK. CMD object is a command package that tell the robot what to do. CMD package consists of command mode, velocity, yaw speed, target position, required body height variables. Then udp.SetSend(cmd) redyes the cmd object  with encoding it to be sent. Finally the udp object sends the command to the Raspberry Pi using udp.send().

        Raspberry Pi 4 also uses unitree\_legged\_sdk to receive commands from and send state information to the user computer. It also communicates with the STM32 Microcontroller which controls all joint motors of the robot.

        \item \textbf{Getting RGB and Depth Frames from Stereo Cameras (unitree\_camera\_sdk)} \\
        As mentioned before, there are 5 sets of fish-eye stereo depth camera units to both provide an RGB frame and depth frame around the robot.

        \begin{figure}[H]
            \centering
            \includegraphics[width=0.8\textwidth]{RobotComponentLocations.jpeg}
            \caption{Legged Robot Sensors and Other Components}
        \end{figure}

        \begin{figure}[H]
            \centering
            \includegraphics[width=0.8\textwidth]{RobotFront.png}
            \caption{Legged Robot Head Supersensory}
        \end{figure}

        Unitree provides unitree\_camera\_sdk to access camera data of the robot. The SDK allows getting raw frames, rectangular frames, depth frames and point cloud frames.

        SDK is based on GStreamer framework and works on Nvidia Jetson Nano devices. GStreamer is a library for constructing graphs of media-handling components. Applications can take advantage of advances in codec and filter technology transparently. Developers can add new codecs and filters by writing a simple plugin with a clean, gener~\cite{gstreamer}. GStreamer creates a media pipeline that allows developers to implement their applications simple and modular.

        unitree\_camera\_sdk creates a pipeline on GStreamer and a virtual video device for each camera connected to the Jetson Nano. Developer can easily read camera information, raw frames, rectangular frames, depth frames and point cloud frames using SDK in their apps.

        \begin{figure}[H]
            \centering
            \includegraphics[width=\textwidth]{RawStereo.png}
            \caption{Raw Frame}
        \end{figure}

        \begin{figure}[H]
            \centering
            \begin{subfigure}[b]{0.4\textwidth}
                \centering
                \includegraphics[width=\textwidth]{StereoRectFrame.png}
                \caption{Rectangular Frame}
            \end{subfigure}
            \hfill
            \begin{subfigure}[b]{0.4\textwidth}
                \centering
                \includegraphics[width=\textwidth]{StereoDeptFrame.png}
                \caption{Depth Frame}
            \end{subfigure}
            \caption{Worked Frames}
        \end{figure}

        Robot has some built in features like human skeleton tracking that uses cameras automatically. There are some services implemented to run these applications automatically while system is starting up. Developer should be whether kill these processes after startup or edit the services not to run processes that uses cameras.

        Dept information is calculated from stereo camera using stereo vision algorithms. First step is matching each pixel of left frame with corresponding pixel in right frame. This step includes computer vision algorithms like Sum of Squared Differences (SSD) and Sum of Absolute Differences (SAD).

        \begin{figure}[H]
            \centering
            \includegraphics[width=0.8\textwidth]{SSDSAD.png}
            \caption{Sum of Squared Differences (SSD) and Sum of Absolute Differences (SAD)}
        \end{figure}

        After every point matched, distance of every point can be calculated using basic geometric calculations like Thales Theorem.

        \begin{figure}[H]
            \centering
            \includegraphics[width=0.8\textwidth]{Thales.png}
            \caption{Thales Theorem}
        \end{figure}

        Stereo cameras provide three dimensional images without using moving parts and high costs. But they have disadvantages as they have advantages. Each camera has lenses to converge and diverge the light coming from different angles and make them fall onto the sensor. That means each camera has a visual angle. Other than that visual angle cannot enter the frame. As we think of the closest points of the baseline of the stereo camera that enters visual angle, it is hard to both match these points in both frames and calculate their distances from camera. Figure shows disparity with respect to distance of the point from baseline of the stereo camera.

        \begin{figure}[H]
            \centering
            \includegraphics[width=0.6\textwidth]{DisparityOnStereo.png}
            \caption{Disparity with respect to Distance}
        \end{figure}

        \item \textbf{Developing Sensor Fusion Algorithm for 3D LIDAR and Stereo Cameras}
        \item \textbf{Simultaneously Localization and Mapping (SLAM) }
        \item \textbf{Automating SLAM with Decision Algorithm }
        \item \textbf{Moving Sensor Fusion Algorithm to Cloud}
        \item \textbf{Further Applications Like Swarm Robots with Centralized Cloud System }
    \end{enumerate}

    To accomplish this much work successfully the proof-of-concept parts should be made in a professional manner. In our earlier course work we acquired these experiences by the leading of our course supervisors. Although we will only implement software level designs, the understanding of mathematics and other engineering knowledge cannot be denied. Since we always advice with our supervisor.

    \begin{table}[H]
        \centering
        \includegraphics[width=0.65\textwidth]{GanttChart.png}
        \caption{Gantt Chart of Processes}
    \end{table}

\section{ENGINEERING STANDARDS AND DESIGN \\ CONSTRAINTS}

The design constraints should be identified and some discussion on how these apply to the design project and their realization are to be included in this section. You may refer to the list of some realistic design constraints that can be found in the Project Design Constraints document. Other constraints can be identified and discussed, if applicable.

    \subsection{ENGINEERING STANDARDS}

    We use the IEEE standards in our project. From the related pages~\cite{enwiki:1120084976} we found the good to have rules as follows:

    IEEE 829-2008, also known as the 829 Standard for Software and System Test Documentation, was an IEEE standard that specified the form of a set of documents for use in eight defined stages of software testing and system testing, each stage potentially producing its own separate type of document. The standard specified the format of these documents, but did not stipulate whether they must all be produced, nor did it include any criteria regarding adequate content for these documents. These were a matter of judgment outside the purview of the standard~\cite{enwiki:1105776890}. There will be a lot of test cases for the design. Since the system is complex, the test scenarios should be documented with known standard in order to maintenance independently from project contributors.

    A software requirements specification (SRS) is a description of a software system to be developed. It is modeled after business requirements specification (CONOPS). The software requirements specification lays out functional and non-functional requirements, and it may include a set of use cases that describe user interactions that the software must provide to the user for perfect interaction~\cite{enwiki:1117244157}. Since we have such a software system will be developed, it should lay on some standards to make the codes and all implementations reusable and modular.

    A software design description (a.k.a. software design document or SDD; just design document; also Software Design Specification) is a representation of a software design that is to be used for recording design information, addressing various design concerns, and communicating that information to the design’s stakeholders~\cite{enwiki:1105986397}. Similarly, the developed system should have standardized documentation as we said in earlier two standards. The three keywords for a good software design are: modular, reusable, and flexible.

    Of course, as the project progresses, it will be important that we comply with more detailed standards. We will continue to mention these in future reports.

    \subsection{DESIGN CONSTRAINTS}

    From the "Project Design Constraints" document, our constraints are:

    \begin{itemize}
        \item \textbf{Economy:} \\
        For the term project duration we have no spending other then the supervisors did. In case of further production, maintenance, we should spend around 18 dollars without counting the costs of the work put forward. We cannot find this such project on the market. But educational project like ours exist.

        \item \textbf{Environment:} \\
        Our robot is battery powered and now on it is only in the development process by different cases. Hence for now we do not have any power consumption calculations (also because we are focused on the software). The radiation we uncovered are on the ISM RF band, legally. And there is no serious pollution of noise, hopefully.

        \item \textbf{Manufacturability:} \\
        Physical implementation will consist of collecting materials and combining them due to use of ready made products. Since, there is no need for high level technologies of manufacturing.

        \item \textbf{Sustainability:} \\
        If we look at the reliability and durability of the design, we can say the legged robot that we use is waterproof but there are a few of external components those are not. Our project is a comprehensive study in which the mathematical foundations of a product that can serve humanity in many areas such as personal use, industrial use and use in disasters are laid and upper engineering problems are solved. In this way, its sustainability is high because it sits on a very comprehensive basis. Also upgrades for scaling purposes are possible to implement since the software we are developing has an abstraction from the hardware.

    \end{itemize}

    and the other constraints for our project specifically due to usage of the some bought products are:

    \begin{itemize}
        \item \textbf{Computing Power}
            \begin{itemize}
                \item \textbf{Internal Nvidia Jetson Nano 4GB:} \\
                Mobile computing devices tend to have less computing power due to their size limits. There are three Nvidia Jetson Nano devices inside of the Unitree Go1 Edu. These devices located in the robot to perform certain tasks. Two of Nvidia Jetson Nano 4GB and one Nvidia Jetson Nano 2 GB development kits are used for getting data from camera and sensors, processing raw data to extract dept information of stereo cameras, human skeleton tracking and so on.

                We don’t need most of the features of the robot like human tracking in our project, so we can lighten its workload and use one of them as our computing unit.

                \item \textbf{External Nvidia Jetson:} \\
                In case of having a trouble with using internal Nvidia Jetson device, we can also use an external one. It can be again a Nvidia Jetson Nano with all resources dedicated to our sensor fusion algorithm or a more advanced model like Xavier Nx. This option will rise the cost of the project, so this is an option only if actually needed.

                \item \textbf{Cloud Computing:} \\
                Computing is any goal-oriented activity requiring, benefiting from, or creating computing machinery~\cite{enwiki:1120224488}. There are different computing elements. CPU’s have ability to work high clock frequencies and have large set of sequential instructions. GPU’s are mostly  Mobile computing devices like Nvidia’s Jetson series development kits may not match the performance requirements in some cases. There are GPU servers to provide high parallel computing power.

            \end{itemize}

            \vspace{10pt}

            \begin{table}[H]
                \centering
                \includegraphics[width=0.8\textwidth]{ComputerComparison.png}
                \caption{Computing Power Comparison}
            \end{table}

            \textbf{Result:}

            We decided to use Internal Jetson Development kits of the robot for sake of cost reduction. But as mentioned before, one of the aims of this project is moving execution of algorithms that need high computing power to the cloud. So, we will be using both internal Jetson devices and cloud computing. That will be give us an opportunity to compare results in both alternatives.

        \item \textbf{Depth Sensor \& Stereo Camera}
            \begin{itemize}
                \item \textbf{Internal Unitree Camera – Super Sensory System:} \\
                Lens Angle ~150°×170°

                \begin{figure}[H]
                    \centering
                    \includegraphics[width=0.5\textwidth]{SuperSensorySystem.png}
                    \caption{Super Sensory System}
                \end{figure}

                \item \textbf{External ZED Stereo Camera:} \\

                \begin{figure}[H]
                    \centering
                    \includegraphics[width=0.5\textwidth]{ZEDCam.png}
                    \caption{ZED Stereo Camera~\cite{ZEDStereoCamera}}
                \end{figure}

                Some specifications of ZED:

                \begin{itemize}
                    \item High-Resolution and High Frame-rate 3D Video Capture (1080p 30fps) 
                    \item Depth Perception indoors and outdoors at up to 20m 
                    \item 6-DoF Positional Tracking 
                    \item Spatial Mapping 
                    \item 110° Wide Angle Cameras 
                \end{itemize}

                \item \textbf{Xbox 360 Kinect:} \\
                \begin{figure}[H]
                    \centering
                    \includegraphics[width=0.5\textwidth]{XBoxKinect.png}
                    \caption{Xbox 360 Kinect~\cite{XboxKinect}}
                \end{figure}

                \begin{enumerate}
                    \item 3D Depth sensor (IR Emitter + IR Camera / Depth Sensor) 
                    \item RGB camera (Color Sensor) 
                    \item Microphone array 
                    \item Tilt motor (for detecting floor and players in the play space) 
                \end{enumerate}

                \item \textbf{Velodyne 80-VLP-16-A 3D LiDAR:} \\
                \begin{figure}[H]
                    \centering
                    \includegraphics[width=0.3\textwidth]{LiDAR.png}
                    \caption{Velodyne LiDAR}
                \end{figure}

                Some specifications of LiDAR:

                \begin{itemize}
                    \item 100 m Range 
                    \item 360° x 30° Viewing Angle 
                    \item 0.3 Million Points/Second 
                    \item 100Mbps Ethernet \& UDP Interface 
                    \item Rated IP67 
                \end{itemize}
            \end{itemize}

        \vspace{10pt}

        \begin{table}[H]
            \centering
            \includegraphics[width=0.8\textwidth]{SensorProsCons.png}
            \caption{Sensor Comparison}
        \end{table}

        \textbf{Result:}

        As we consider the trade-offs that are given at table-x, we decided to move on with Unitree Stereo Cameras and Velodyne VLP-16 3D Lidar.

        \item \textbf{Programming Languages} \\
        There are different tools we need to use to control the robot, access its sensor measurements and camera data etc. The robot manufacturer provides SDKs to access and control the robot. These SDKs are implemented using C++ programming language.

        There is an SDK provided by Unitree to control robots with software. The SDK is implemented in C/C++ languages, but it supports programming in C, C++ and Python languages. So, at some point, we had to use those languages.

        Besides these, ROS technology has a lot of programming languages provided. Since, we have no restriction from this side, hopefully.

        \vspace{10pt}

        \textbf{Result:}

        As we consider the performance and easy usage, we are planning to use Python only for algorithm development. After that we will be implement same algorithms in C++ to improve performance. SDK handles most of the ROS work, but we use ROS commands if it needed.

    \end{itemize}

\section{SUSTAINABLE DEVELOPMENT GOALS}

Our project is a comprehensive study in which the mathematical foundations of a product that can serve humanity in many areas such as personal use, industrial use and use in disasters are laid and upper engineering problems are solved. In this way, its sustainability is high because it sits on a very comprehensive basis. If we look at our project in the context of global goals, it is a good example of Goal 9: Industry, Innovation, and Infrastructure. In addition, our design successfully complies with Goal 12: Responsible consumption and production, as abstractions are kept neatly. For example, scaling in design using the same software infrastructure is quite possible and simple.

\section{BACKGROUND}

For this project the need of strong background is irrefutable. Since there is software oriented design for our case, the reasearchs are more leading. Background comes from the earlier courses in our department, additional research, professional life, and even daily life. Also tricky parts are coming from experiences of our supervisors (know-hows, etc.).

    \subsection{BACKGROUND ACQUIRED IN EARLIER COURSE WORK}

    The sensor fusion operation require a lot mathematical theory and its practical implementations. At this point, the importance of our Calculus (MAT123/124) and Engineering Mathematics (MAT235/236), Signals and Systems (ELE301), Control Systems (ELE354), and Digital Signal Processing (ELE407) courses takes place.

    But they are not enough to come up with a practical project. To design the system of this large computing environment and the network we should lay on the Computers and Programming (ELE107), Computer Programming (ELE118/120), Data Structures (ELE411), and Data Communication (ELE412) courses. For the sensor fusion operation we also use the knowledge from the Image Processing (ELE492) course for filtering bases and the hands-on programming purposes.

    \subsection{BACKGROUND ACQUIRED THROUGH ADDITIONAL\\ RESEARCH}

    Although our courses that we took in our department are pretty important for the theoretical base and introductory for the implementation, the know-hows of the programming languages, computer architectures, design patterns, and the protocols that we use are beyond from the course curriculums. To get those knowledge, we made a lot of research about those. Besides this, specifically the programming language practices are coming from our part-time jobs, internships, extracurricular studies, and further studies, and coding practices which coming from our supervisors suggestions.

    To be more specific, the CPP and the Python programming languages are not taught but we should already know those to go further in our design. Another example is ROS (robot operating system) concept that is special for our purposes has nothing to do with our courses. Sensor fusion is also a specific research area of the digital signal processing, control systems, mathematics, and Computer Vision~\cite{enwiki:1120271364} for ur case, obviously. Hence, we also made lots of preliminary study in this topic, other than courses.

\addcontentsline{toc}{section}{REFERENCES}
\bibliographystyle{plain}
\bibliography{refs}
\end{document}

### 2023-01-22 reports/report2/main.tex
Interim report 2 added.

\documentclass[12pt]{article}
\usepackage[left=1in, right=1in, top=1in, bottom=1in]{geometry}
\usepackage{graphicx}
\usepackage{url}
\usepackage{cite}
\usepackage{float}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{hyperref}
\hypersetup{
    colorlinks,
    citecolor=black,
    filecolor=black,
    linkcolor=black,
    urlcolor=black
}

\renewcommand{\familydefault}{\sfdefault}

\begin{document}
\begin{titlepage}
    \begin{center}
        \LARGE
        \textbf{ELE 401 - GRADUATION PROJECT I}

        \Large
        \textbf{SECOND INTERIM REPORT}

        \vspace{70pt}

        \textit{
            Hacettepe University \\
            Department of Electrical and Electronics Engineering
        }
    \end{center}

    \vspace{90pt}

    \large

    \textbf{Project Title:} Sensor Fusion in the Cloud \\

    \textbf{Project Group Members:} Ertuğrul Tiyek, Ahmet Yusuf Şirin

    \vspace{30pt}

    \textbf{Project Supervisor:} Asst. Prof. Dr. İsmail Uyanık \\

    \textbf{Submission Date:} 11.12.2022

    \vspace{\fill}

    \begin{center}
        \textit{FALL 2022-2023}
    \end{center}
\end{titlepage}

\clearpage

\tableofcontents
\listoffigures
\listoftables

\clearpage

\begin{abstract}
    From the point on we are studying on the sensor fusion. In this report we will briefly emphasize the legged robot usage for our final decisions. On the other hand, the concept ot the sensor fusion and the alternative methodologies to implement it correctly to serve the main purpose of sensor fusion, that is reliable odometry data extraction.
\end{abstract}

\section{INTRODUCTION}

One of the most important aim of the modern technology is the sensing the environment. Living beings have sense organs to know the environment. They have brains to make fusion the sensor data and augment the environmental information taken from sense organs. For example, human standing in balance, we can say that they use the data coming from their eyes and ears together.

People are studying to make this sensing and augmenting artificially, too. They use computers to achieve this. But computers cannot sense the environment directly. They need some devices called sensors. There are all kinds of sensors to measure different environmental quantities. For same quantity, there are different sensors using different technologies, have different precision and accuracy levels. Each sensor has its advantages and disadvantages. Even sensors that are serving a different purpose can help to augment same information other than their main. This is known as "sensor fusion"~\cite{enwiki:1115352853}. Just like humans, robots are use camera and inertial sensors to stand in balance.

In our project, sensor fusion handles extracting depth information. Sensor fusion can be implemented using specific algorithms like EKF, CNN etc. according to quantity that will be calculated, number and types of sensors used, accuracy and performance needed. We are planning to develop a sensor fusion algorithm with the help of living organisms. As we mentioned before, living beings perform sensor fusion perfectly. We will be examining the zebra fish and how it processes data of its sense organs to produce depth information. Then we will try to develop similar algorithm or improve our algorithm with data taken from zebra fish.

Depth information can be used for many applications like automatic driven vehicles, UAVs, Augmented Reality devices and 3D modeling of areas. Our concept in this project will be autonomous controlled SLAM Robot. Simultaneous localization and mapping (SLAM) is the computational problem of constructing or updating a map of an unknown environment while simultaneously keeping track of an agent's location within it~\cite{enwiki:1120084976}. We use a robot to move the system.

Another concept we aim to use is cloud computing. Mobile computing devices may not match the performance requirements in some cases. To achieve more accurate depth information, complex control algorithms or neural networks will be used in our project which needs high computing power. As an alternative to the mobile computing devices, we may use cloud computing.

\section{METHODS}

The definition of our project is explained in detail in the first interim report. In the light of the definition, we try to make the decisions and the design parameters clear. To sake of easiness we draw some paths going same destination and arguing to reach the optimal path on the alternatives. Those alternatives are listed and described below. Then we will clearly elucidate our method now.

Since the project includes plenty of areas to work those are distinct, we divide them into subtopics. Those are:

\begin{enumerate}
    \item Sensor Choosing Methods
    \item Sensor Fusion Methods
    \item Sensor Data Enhancement Methods
    \item Fusion Data Enhancement Methods
\end{enumerate}

\subsection{Sensor Choosing Methods}

Our aim in this project is to achieve high reliability by using (fusing) all the sensors we have together. However, when we want to do all these together, we foresee that the project will reach very different dimensions. Therefore, while continuing the project, we decided to select a number of sensors in order to focus on algorithm development.

We set ourselves criteria when choosing sensors. Easy to use sensors is one of our main priorities due to our job description. In addition, considering the algorithmic complexity, working with sensors with linear system response is also an important criterion. In addition, the frequency of receiving data from the sensor is of great importance. Because if the sensor data we will fuse does not come at the same frequency, we will need to add operations to our algorithm to make the sensor frequencies equal by upsampling and downsampling.

We basically divided the sensors we have into two groups as those producing RGBD data or those producing Depth Point Cloud data.

\subsubsection{Method 1: Internal Sensor Using}
The internal sensors of the legged robot include five pairs of stereo cameras that produce RGBD data, and ultrasonic sensors that produce Depth Point Cloud data when viewed cumulatively, SuperSensorySystem. In addition to these, each foot has pressure sensors and an internal IMU.

\subsubsection{Method 2: External Sensor Using}
Our external sensors are ZED Stereo camera which produces RGBD data, Velodyne 3D LiDAR which produces Depth Point Cloud data and Xbox 360 Kinect having 3D depth sensor, RGB camera, microphone array, and also tilt motor. In the future, it is also possible to add sensors such as IMU and GPS to support odometry information.

\subsubsection{Method 3: Mixed Sensor Using}
We can also use our sensors mixed. However, complexity will increase in this method and the workload to be done before the sensor fusion algorithm will become more.

\begin{itemize}
    \item \textbf{Internal Unitree Camera – Super Sensory System:} \\
          Lens Angle ~150°×170°

          \begin{figure}[H]
              \centering
              \includegraphics[width=0.5\textwidth]{SuperSensorySystem.png}
              \caption{Super Sensory System}
          \end{figure}

    \item \textbf{External ZED Stereo Camera:} \\

          \begin{figure}[H]
              \centering
              \includegraphics[width=0.5\textwidth]{ZEDCam.png}
              \caption{ZED Stereo Camera~\cite{ZEDStereoCamera}}
          \end{figure}

          Some specifications of ZED:

          \begin{itemize}
              \item High-Resolution and High Frame-rate 3D Video Capture (1080p 30fps)
              \item Depth Perception indoors and outdoors at up to 20m
              \item 6-DoF Positional Tracking
              \item Spatial Mapping
              \item 110° Wide Angle Cameras
          \end{itemize}

    \item \textbf{Xbox 360 Kinect:} \\
          \begin{figure}[H]
              \centering
              \includegraphics[width=0.5\textwidth]{XBoxKinect.png}
              \caption{Xbox 360 Kinect~\cite{XboxKinect}}
          \end{figure}

          \begin{enumerate}
              \item 3D Depth sensor (IR Emitter + IR Camera / Depth Sensor)
              \item RGB camera (Color Sensor)
              \item Microphone array
              \item Tilt motor (for detecting floor and players in the play space)
          \end{enumerate}

    \item \textbf{Velodyne 80-VLP-16-A 3D LiDAR:} \\
          \begin{figure}[H]
              \centering
              \includegraphics[width=0.3\textwidth]{LiDAR.png}
              \caption{Velodyne LiDAR}
          \end{figure}

          Some specifications of LiDAR:

          \begin{itemize}
              \item 100 m Range
              \item 360° x 30° Viewing Angle
              \item 0.3 Million Points/Second
              \item 100Mbps Ethernet \& UDP Interface
              \item Rated IP67
          \end{itemize}
\end{itemize}

We have taken all these into consideration while doing the preliminary design, and together with the evaluations under the following headings, we have determined a starting point for ourselves. You can find the details under the Preliminary design heading.

% Stereo Camera Decision Method
% LiDAR Decision Method
% Internal IMU usage on legged robot
% Documentation about sensors in marketplace and we owned (engineering decisions on costs)

\subsection{Sensor Fusion Methods}

\subsubsection{Method 1: Complementary Filter}
We wanted to set out using known methods for sensor fusion. In this way, we aimed to improve the performance we want to achieve with methods that we understand fundamentally, instead of using the resources in the literature directly. In fact, we started out by examining the usages of Kalman filter and extended Kalman filter, which are quite commonly used for state estimations made from aircraft IMU sensors, in this field. We predicted that if you continue on your way by choosing the most primitive complementary filter among them, we will be able to understand more clearly that we are moving away from the target in every development we make.

\begin{figure}[H]
    \centering
    \includegraphics[width=0.7\textwidth]{CF.png}
    \caption{Complementary Filter Example for Aircraft Attitude}
\end{figure}

\subsubsection{Method 2: Kalman Filter}
Using the Kalman filter is also a method between using the extended Kalman filter and using a complementary filter. Since we want to use it with some methods that will enhance compelemntery filtery, we have already produced our own gain function. However, we also considered the Kalman filter as a method.

\begin{figure}[H]
    \centering
    \includegraphics[width=0.7\textwidth]{KF.png}
    \caption{Kalman Filter Example for Aircraft Attitude}
\end{figure}

\subsubsection{Method 3: Extended Kalman Filter Method}
In addition to previous methods, we also evaluated the opinion that preliminary design can speed up the process by directly using the most advanced extenden Kalman filter. However, unlike in the aircraft scenario, we could not find an example use for depth data to be suitable for our scenario~\cite{9307398}.

\begin{figure}[H]
    \centering
    \includegraphics[width=0.7\textwidth]{EKF.png}
    \caption{Extended Kalman Filter Example for Aircraft Attitude}
\end{figure}

A comparison in which the models in which the state estimator methods are effective and the noise characteristics are given together is given in the X table, cost function comparisons for further analysis~\cite{The_cost_function_of_the_data_fusion_process_and_i}.

\begin{table}[H]
    \centering
    \includegraphics[width=0.8\textwidth]{CostComparison.jpeg}
    \caption{Cost Function Comparison}
\end{table}

\subsection{Sensor Data Enhancement Methods}

We compare the following methods for data enhancement to following data fusion.

\subsubsection{Method 1: Only Data Rate and Range Matching}
Since the data we want to obtain is odometry, we evaluated creating parametric inputs in the methods we used while continuing the project. Because we sould think about Sensor Data Rate - State Estimation Relation. In this way, we aimed to analyze whether we could establish a relationship between the quality of the raw data and the esimated odometry data. After some tests we will briefly report our results about early sensor data enhancement for sensor fusion.

\subsubsection{Metod 2: Papoulis-Gerchberg Algorithm for Depth Point Cloud}
We are searching about Papoulis-Gerchberg algorithm effects on LiDAR and Stereo depth map results to enhance the data resolution before data fusion. An image super-resolution method, the Papoulis–Gerchberg (P–G) algorithm, on range data represented in the form of a greyscale image. However, the low convergence rate of the original P–G algorithm impedes its use for online applications~\cite{ozbay2015high, kuzucu2018enhancing}. By referencing this method we also proposed as an enhancement method.~\cite{6280128}

\section{PRELIMINARY DESIGN}

ZED stereo camera and Velodyne 3D LiDAR were chosen for the preliminary design, which are sensors that are well-documented, have a large community, and are easier to use and access resources than the internal sensors of the legged robot. This sensor will also minimize the complexity of the algorithm to be used in sensor fusion, as they offer a wide resolution and data rate band.

As the sensor fusion algorithm, it was ordered from simple to complex or from low to high performance, and it was decided to make a start with a complementary filter. Again, it was decided to use methods such as edge preservation~\cite{9307398} as an aid to the main algorithm in order to increase the sensor fusion performance.

We choose the complementary filter for fusing the depth information of two sensors. Each sensor has its own properties, advantages and disadvantages. According to these characteristics of sensors we tune our sensor fusion algorithm. But like all control systems, we first regulate our input signals to work with.

Thing we have to match in these two sensors data is viewing angle. Stereo vision camera has a viewing angle upto 110°. We simply reduced the viewing angle of the 3D lidar to match the stereo vision cameras viewing angle. But further steps we are planning to use also 360° lidar information to enhance environmental data while performing SLAM.

Raw data provided by lidar and stereo vision cameras in different formats. Stereo vision camera extracts depth frames as well as RGBD (Red Green Blue Depth) in cartesian form. 3D lidar extracts data in point cloud form which is basically in spherical form. In the other hand the resolution of stereo vision data is available in 2208x1242 at 15 fps, 1920x1080 at 30fps, 1280x720 at 60 fps, 672x376 at 100 fps. The 3D lidar works completely different and samples much less data. In order to match the sampled data from two sensors, we decided upsampling the lidar data and converting it to cartesian form. We use dynamic Papoulis–Gerchberg algorithm to augment the lidar data and then linearly interpolate this information while converting point cloud data to depth frame.

Iterative Papoulis–Gerchberg algorithm is the method used to recover the lost samples of the signal. It is mostly used for super resolution images in image processing. In a single frame manner, iterative Papoulis–Gerchberg algorithm will be enough to increase data in a single point cloud of a lidar. But our aim is continuously matching the frames, we will use the dynamic Papoulis–Gerchberg algorithm. This algorithm make use of stillness of background and continuity of movement to reduce computational effort in Papoulis–Gerchberg.

There is also difference between sample rates of two sensors. Stereo camera samples frames in range of 15 to 100 frames per second. Lidar samples point clouds by rotating its lidar continuously. Each rotation provides one whole 360° sample (30° vertical). Lidar works in range 300 rpms to 600 rpms. That provides us 5-to-10 point clouds per second. We aim to up-sample the  number of point cloud per second by again linear interpolation.

There is one more problem with the lidar. Ability of lidar is not limited by the only rotation but also sample rate of sensor. Increasing rotation speed reduces the quality of the point clouds. One part of the project is experiencing the different data rates and calculating the processing effort, transmission capabilities and response times. So we will optimize resolution and data rates in the content of this project.

\subsection*{Sensor Fusion Algorithm}

We matched the formats of the input signals to keep the sensor fusion simple. There are two depth images as input signals to get more accurate depth image.

As a preliminary design, we are planning to calculate the depth image in per pixel fashion. As mentioned before, the complementary filter is formulated as:

\begin{equation}
    \hat{x} = \hat{x}_1 \alpha + \hat{x}_2 (1 - \alpha)
\end{equation}

We will generate $\alpha$ generation function to calculate $\alpha$ for each pixel and condition. Input sources are not functioning perfect in every conditions. There are some internal and external variables that effects the accuracy of measurement. Let’s observe these variables:

\subsubsection*{Variables That Effect Lidar:}

\begin{itemize}
    \item Beam Angle:

          3D lidar scans the environment in a three-dimensional manner. Velodyne VLP-16 scans 360° horizontally and 30° vertically. In different beam angles, Lidar accuracy depends. There is an article that contains more information about beam angle effect.

          \begin{figure}[H]
              \centering
              \includegraphics[width=0.5\textwidth]{LidarAngles.jpeg}
              \caption{Velodyne LiDAR Beam Angles~\cite{VelodynePerformance}}
          \end{figure}

    \item Rotation Speed:

          Sample rate of the lidar decreases as rotation speed of the lidar increases. It effects the quality of the depth image. But increasing the rotation speed provides us approximation to real time application.

    \item Distance:

          3D lidars can work in a range that is specified in datasheet. Out of this range, lidar cannot work properly. As we think of the working principle of the lidar, the processing time is not enough to measure too close points. The Velodyne VLP-16 model has a working range between 1m to 100m.

    \item Rough and Uneven Views:

          The Lidar is using wavelengths very close to visible light to observe visible objects. In some environments like forests, rough areas, rainy and foggy weathers there are lots of indentations like branches, twigs and leaves or water drops. Each of this bits and pieces prevents the infrared laser lights coming from lidar and creates canopies. Instead of a clean point cloud, lidar gets a noisy one.

          We can measure the effect of the bits and pieces in the environment to accuracy of lidar measurement by calculating the gradient of depth frame around target pixel.

          \begin{figure}[H]
              \centering
              \includegraphics[width=0.9\textwidth]{LidarViews.png}
              \caption{Gradient of Depth Image Around Target Pixel Gives Roughness of the Environment~\cite{https://doi.org/10.48550/arxiv.2010.09076}}
          \end{figure}
\end{itemize}

\subsubsection*{Variables That Effect Stereo Vision Camera:}

\begin{itemize}
    \item Distance:

          According to viewing angles and algorithm used to extract depth information from stereo image, there is an uncertainty appears. And this uncertainty increases as distance of the pixel increases. Uncertainty reduces the confidence of the input as it increased.

          \begin{figure}[H]
              \centering
              \includegraphics[width=0.4\textwidth]{CamDistance.png}
              \caption{Stereo Camera Depth Error Calculation~\cite{DisparityCalculator, gallup2008variable}}
          \end{figure}

    \item Brightness:

          The brightness of the environment is important for stereo vision depth estimation. As considering the working principles of the stereo vision depth estimation, there are two frames taken from different angles. First step is matching each pixel of left frame with corresponding pixel in right frame. This step includes computer vision algorithms like Sum of Squared Differences (SSD) and Sum of Absolute Differences (SAD). To perform matching properly, the frames should have well distributed histograms and good contrast values. If there is the brightness of frames are too high or too low, the algorithms cannot match the corresponding pixels.

          We can calculate the effect of brightness by observing the histogram of the RGB frames.
    \item Straight and no-Contrast Views:

          Like the brightness, the content of the frame is also important. If the frame includes low contrast, straight or similar patterns, the matching algorithms won’t work properly. For example, If the frame includes a flat white wall, the algorithm can not distinguish pixels in different frames. Likewise, if the frame has a periodic pattern like damas pattern, the algorithm also not work well.

          In this case we can calculate the effect of straightness by calculating gradient around target pixel in RGB frames.
\end{itemize}

According to these parameters, we will derive confidence generator functions that will calculate a confidence value for each input. By normalizing one of the confidence values will give us $\alpha$ value.

\begin{equation}
    \alpha = \frac{C_{LIDAR}}{C_{LIDAR} + C_{STEREO}}
\end{equation}

\begin{figure}[H]
    \centering
    \includegraphics[width=0.9\textwidth]{AlphaFunction.jpeg}
    \caption{Alpha Function for Complementary Filter Enhancement}
\end{figure}

After extracting $\alpha$ value for each pixel of depth frame, we can simply apply complementary filter.

\begin{figure}[H]
    \centering
    \includegraphics[width=0.9\textwidth]{CompFiltBD.jpeg}
    \caption{Complementary Filter Block Diagram}
\end{figure}

\addcontentsline{toc}{section}{REFERENCES}
\bibliographystyle{plain}
\bibliography{refs}
\end{document}

### 2023-01-22 reports/report3/main.tex
Term report (report3) added

--- 2022-11-06 reports/report1/main.tex
@@
         \Large
-        \textbf{FIRST INTERIM REPORT}
+        \textbf{TERM REPORT}
 
@@
 
-    \textbf{Submission Date:} 05.11.2022
+    \textbf{Submission Date:} 22.01.2023
 
@@
 \clearpage
+
+\begin{abstract}
+    One of the most important aim of the modern technology is the sensing the environment.
+Living beings have sense organs to know the environment.
+They have brains to make fusion the sensor data and augment the environmental information taken from sense organs.
+For example, human standing in balance, we can say that they use the data coming from their eyes and ears together.
+
+    People are studying to make this sensing and augmenting artificially, too.
+They use computers to achieve this.
+But computers cannot sense the environment directly.
+They need some devices called sensors.
+There are all kinds of sensors to measure different environmental quantities.
+For same quantity, there are different sensors using different technologies, have different precision and accuracy levels.
+Each sensor has its advantages and disadvantages.
+Even sensors that are serving a different purpose can help to augment same information other than their main.
+This is known as "sensor fusion"~\cite{enwiki:1115352853}.
+Just like humans, robots are use camera and inertial sensors to stand in balance.
+
+    In our project, sensor fusion handles extracting depth information.
+Sensor fusion can be implemented using specific algorithms like EKF, CNN etc.
+according to quantity that will be calculated, number and types of sensors used, accuracy and performance needed.
+We are planning to develop a sensor fusion algorithm with the help of living organisms.
+As we mentioned before, living beings perform sensor fusion perfectly.
+We will be examining the zebra fish and how it processes data of its sense organs to produce depth information.
+Then we will try to develop similar algorithm or improve our algorithm with data taken from zebra fish.
+
+    Depth information can be used for many applications like automatic driven vehicles, UAVs, Augmented Reality devices and 3D modeling of areas.
+Our concept in this project will be autonomous controlled SLAM Robot.
+Simultaneous localization and mapping (SLAM) is the computational problem of constructing or updating a map of an unknown environment while simultaneously keeping track of an agent's location within it~\cite{enwiki:1120084976}.
+We use a robot to move the system.
+
+    Another concept we aim to use is cloud computing.
+Mobile computing devices may not match the performance requirements in some cases.
+To achieve more accurate depth information, complex control algorithms or neural networks will be used in our project which needs high computing power.
+As an alternative to the mobile computing devices, we may use cloud computing.
+
+    From the point on we are studying on the sensor fusion.
+In this report we will briefly emphasize the legged robot usage for our final decisions.
+On the other hand, the concept ot the sensor fusion and the alternative methodologies to implement it correctly to serve the main purpose of sensor fusion, that is reliable odometry data extraction.
+\end{abstract}
 
@@
 
+\section{METHODS}
+
+The definition of our project is explained in detail in the first interim report.
+In the light of the definition, we try to make the decisions and the design parameters clear.
+To sake of easiness we draw some paths going same destination and arguing to reach the optimal path on the alternatives.
+Those alternatives are listed and described below.
+Then we will clearly elucidate our method now.
+
+Since the project includes plenty of areas to work those are distinct, we divide them into subtopics. Those are:
+
+\begin{enumerate}
+    \item Sensor Choosing Methods
+    \item Sensor Fusion Methods
+    \item Sensor Data Enhancement Methods
+    \item Fusion Data Enhancement Methods
+\end{enumerate}
+
+\subsection{Sensor Choosing Methods}
+
+Our aim in this project is to achieve high reliability by using (fusing) all the sensors we have together.
+However, when we want to do all these together, we foresee that the project will reach very different dimensions.
+Therefore, while continuing the project, we decided to select a number of sensors in order to focus on algorithm development.
+
+We set ourselves criteria when choosing sensors.
+Easy to use sensors is one of our main priorities due to our job description.
+In addition, considering the algorithmic complexity, working with sensors with linear system response is also an important criterion.
+In addition, the frequency of receiving data from the sensor is of great importance.
+Because if the sensor data we will fuse does not come at the same frequency, we will need to add operations to our algorithm to make the sensor frequencies equal by upsampling and downsampling.
+
+We basically divided the sensors we have into two groups as those producing RGBD data or those producing Depth Point Cloud data.
+
+\subsubsection{Method 1: Internal Sensor Using}
+The internal sensors of the legged robot include five pairs of stereo cameras that produce RGBD data, and ultrasonic sensors that produce Depth Point Cloud data when viewed cumulatively, SuperSensorySystem.
+In addition to these, each foot has pressure sensors and an internal IMU.
+
+\subsubsection{Method 2: External Sensor Using}
+Our external sensors are ZED Stereo camera which produces RGBD data, Velodyne 3D LiDAR which produces Depth Point Cloud data and Xbox 360 Kinect having 3D depth sensor, RGB camera, microphone array, and also tilt motor.
+In the future, it is also possible to add sensors such as IMU and GPS to support odometry information.
+
+\subsubsection{Method 3: Mixed Sensor Using}
+We can also use our sensors mixed. However, complexity will increase in this method and the workload to be done before the sensor fusion algorithm will become more.
+
+\begin{itemize}
+    \item \textbf{Internal Unitree Camera – Super Sensory System:} \\
+          Lens Angle ~150°×170°
+
+          \begin{figure}[H]
+              \centering
+              \includegraphics[width=0.5\textwidth]{SuperSensorySystem.png}
+              \caption{Super Sensory System}
+          \end{figure}
+
+    \item \textbf{External ZED Stereo Camera:} \\
+
+          \begin{figure}[H]
+              \centering
+              \includegraphics[width=0.5\textwidth]{ZEDCam.png}
+              \caption{ZED Stereo Camera~\cite{ZEDStereoCamera}}
+          \end{figure}
+
+          Some specifications of ZED:
+
+          \begin{itemize}
+              \item High-Resolution and High Frame-rate 3D Video Capture (1080p 30fps)
+              \item Depth Perception indoors and outdoors at up to 20m
+              \item 6-DoF Positional Tracking
+              \item Spatial Mapping
+              \item 110° Wide Angle Cameras
+          \end{itemize}
+
+    \item \textbf{Xbox 360 Kinect:} \\
+          \begin{figure}[H]
+              \centering
+              \includegraphics[width=0.5\textwidth]{XBoxKinect.png}
+              \caption{Xbox 360 Kinect~\cite{XboxKinect}}
+          \end{figure}
+
+          \begin{enumerate}
+              \item 3D Depth sensor (IR Emitter + IR Camera / Depth Sensor)
+              \item RGB camera (Color Sensor)
+              \item Microphone array
+              \item Tilt motor (for detecting floor and players in the play space)
+          \end{enumerate}
+
+    \item \textbf{Velodyne 80-VLP-16-A 3D LiDAR:} \\
+          \begin{figure}[H]
+              \centering
+              \includegraphics[width=0.3\textwidth]{LiDAR.png}
+              \caption{Velodyne LiDAR}
+          \end{figure}
+
+          Some specifications of LiDAR:
+
+          \begin{itemize}
+              \item 100 m Range
+              \item 360° x 30° Viewing Angle
+              \item 0.3 Million Points/Second
+              \item 100Mbps Ethernet \& UDP Interface
+              \item Rated IP67
+          \end{itemize}
+\end{itemize}
+
+We have taken all these into consideration while doing the preliminary design, and together with the evaluations under the following headings, we have determined a starting point for ourselves.
+You can find the details under the Preliminary design heading.
+
+% Stereo Camera Decision Method
+% LiDAR Decision Method
+% Internal IMU usage on legged robot
+% Documentation about sensors in marketplace and we owned (engineering decisions on costs)
+
+\subsection{Sensor Fusion Methods}
+
+\subsubsection{Method 1: Complementary Filter}
+We wanted to set out using known methods for sensor fusion.
+In this way, we aimed to improve the performance we want to achieve with methods that we understand fundamentally, instead of using the resources in the literature directly.
+In fact, we started out by examining the usages of Kalman filter and extended Kalman filter, which are quite commonly used for state estimations made from aircraft IMU sensors, in this field.
+We predicted that if you continue on your way by choosing the most primitive complementary filter among them, we will be able to understand more clearly that we are moving away from the target in every development we make.
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.7\textwidth]{CF.png}
+    \caption{Complementary Filter Example for Aircraft Attitude}
+\end{figure}
+
+\subsubsection{Method 2: Kalman Filter}
+Using the Kalman filter is also a method between using the extended Kalman filter and using a complementary filter.
+Since we want to use it with some methods that will enhance compelemntery filtery, we have already produced our own gain function.
+However, we also considered the Kalman filter as a method.
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.7\textwidth]{KF.png}
+    \caption{Kalman Filter Example for Aircraft Attitude}
+\end{figure}
+
+\subsubsection{Method 3: Extended Kalman Filter Method}
+In addition to previous methods, we also evaluated the opinion that preliminary design can speed up the process by directly using the most advanced extenden Kalman filter.
+However, unlike in the aircraft scenario, we could not find an example use for depth data to be suitable for our scenario~\cite{9307398}.
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.7\textwidth]{EKF.png}
+    \caption{Extended Kalman Filter Example for Aircraft Attitude}
+\end{figure}
+
+A comparison in which the models in which the state estimator methods are effective and the noise characteristics are given together is given in the X table, cost function comparisons for further analysis~\cite{The_cost_function_of_the_data_fusion_process_and_i}.
+
+\begin{table}[H]
+    \centering
+    \includegraphics[width=0.8\textwidth]{CostComparison.jpeg}
+    \caption{Cost Function Comparison}
+\end{table}
+
+\subsection{Sensor Data Enhancement Methods}
+
+We compare the following methods for data enhancement to following data fusion.
+
+\subsubsection{Method 1: Only Data Rate and Range Matching}
+Since the data we want to obtain is odometry, we evaluated creating parametric inputs in the methods we used while continuing the project.
+Because we sould think about Sensor Data Rate - State Estimation Relation.
+In this way, we aimed to analyze whether we could establish a relationship between the quality of the raw data and the esimated odometry data.
+After some tests we will briefly report our results about early sensor data enhancement for sensor fusion.
+
+\subsubsection{Metod 2: Papoulis-Gerchberg Algorithm for Depth Point Cloud}
+We are searching about Papoulis-Gerchberg algorithm effects on LiDAR and Stereo depth map results to enhance the data resolution before data fusion.
+An image super-resolution method, the Papoulis–Gerchberg (P–G) algorithm, on range data represented in the form of a greyscale image.
+However, the low convergence rate of the original P–G algorithm impedes its use for online applications~\cite{ozbay2015high, kuzucu2018enhancing}.
+By referencing this method we also proposed as an enhancement method.~\cite{6280128}
+
+\section{PRELIMINARY DESIGN}
+
+ZED stereo camera and Velodyne 3D LiDAR were chosen for the preliminary design, which are sensors that are well-documented, have a large community, and are easier to use and access resources than the internal sensors of the legged robot.
+This sensor will also minimize the complexity of the algorithm to be used in sensor fusion, as they offer a wide resolution and data rate band.
+
+As the sensor fusion algorithm, it was ordered from simple to complex or from low to high performance, and it was decided to make a start with a complementary filter.
+Again, it was decided to use methods such as edge preservation~\cite{9307398} as an aid to the main algorithm in order to increase the sensor fusion performance.
+
+We choose the complementary filter for fusing the depth information of two sensors.
+Each sensor has its own properties, advantages and disadvantages.
+According to these characteristics of sensors we tune our sensor fusion algorithm.
+But like all control systems, we first regulate our input signals to work with.
+
+Thing we have to match in these two sensors data is viewing angle.
+Stereo vision camera has a viewing angle upto 110°.
+We simply reduced the viewing angle of the 3D lidar to match the stereo vision cameras viewing angle.
+But further steps we are planning to use also 360° lidar information to enhance environmental data while performing SLAM.
+
+Raw data provided by lidar and stereo vision cameras in different formats.
+Stereo vision camera extracts depth frames as well as RGBD (Red Green Blue Depth) in cartesian form.
+3D lidar extracts data in point cloud form which is basically in spherical form.
+In the other hand the resolution of stereo vision data is available in 2208x1242 at 15 fps, 1920x1080 at 30fps, 1280x720 at 60 fps, 672x376 at 100 fps.
+The 3D lidar works completely different and samples much less data.
+In order to match the sampled data from two sensors, we decided upsampling the lidar data and converting it to cartesian form.
+We use dynamic Papoulis–Gerchberg algorithm to augment the lidar data and then linearly interpolate this information while converting point cloud data to depth frame.
+
+Iterative Papoulis–Gerchberg algorithm is the method used to recover the lost samples of the signal.
+It is mostly used for super resolution images in image processing.
+In a single frame manner, iterative Papoulis–Gerchberg algorithm will be enough to increase data in a single point cloud of a lidar.
+But our aim is continuously matching the frames, we will use the dynamic Papoulis–Gerchberg algorithm.
+This algorithm make use of stillness of background and continuity of movement to reduce computational effort in Papoulis–Gerchberg.
+
+There is also difference between sample rates of two sensors.
+Stereo camera samples frames in range of 15 to 100 frames per second.
+Lidar samples point clouds by rotating its lidar continuously.
+Each rotation provides one whole 360° sample (30° vertical).
+Lidar works in range 300 rpms to 600 rpms.
+That provides us 5-to-10 point clouds per second.
+We aim to up-sample the  number of point cloud per second by again linear interpolation.
+
+There is one more problem with the lidar.
+Ability of lidar is not limited by the only rotation but also sample rate of sensor.
+Increasing rotation speed reduces the quality of the point clouds.
+One part of the project is experiencing the different data rates and calculating the processing effort, transmission capabilities and response times.
+So we will optimize resolution and data rates in the content of this project.
+
+\subsection*{Sensor Fusion Algorithm}
+
+We matched the formats of the input signals to keep the sensor fusion simple. There are two depth images as input signals to get more accurate depth image.
+
+As a preliminary design, we are planning to calculate the depth image in per pixel fashion. As mentioned before, the complementary filter is formulated as:
+
+\begin{equation}
+    \hat{x} = \hat{x}_1 \alpha + \hat{x}_2 (1 - \alpha)
+\end{equation}
+
+We will generate $\alpha$ generation function to calculate $\alpha$ for each pixel and condition.
+Input sources are not functioning perfect in every conditions.
+There are some internal and external variables that effects the accuracy of measurement.
+Let’s observe these variables:
+
+\subsubsection*{Variables That Effect Lidar:}
+
+\begin{itemize}
+    \item Beam Angle:
+
+          3D lidar scans the environment in a three-dimensional manner.
+Velodyne VLP-16 scans 360° horizontally and 30° vertically.
+In different beam angles, Lidar accuracy depends.
+There is an article that contains more information about beam angle effect.
+
+          \begin{figure}[H]
+              \centering
+              \includegraphics[width=0.5\textwidth]{LidarAngles.jpeg}
+              \caption{Velodyne LiDAR Beam Angles~\cite{VelodynePerformance}}
+          \end{figure}
+
+    \item Rotation Speed:
+
+          Sample rate of the lidar decreases as rotation speed of the lidar increases.
+It effects the quality of the depth image.
+But increasing the rotation speed provides us approximation to real time application.
+
+    \item Distance:
+
+          3D lidars can work in a range that is specified in datasheet.
+Out of this range, lidar cannot work properly.
+As we think of the working principle of the lidar, the processing time is not enough to measure too close points.
+The Velodyne VLP-16 model has a working range between 1m to 100m.
+
+    \item Rough and Uneven Views:
+
+          The Lidar is using wavelengths very close to visible light to observe visible objects.
+In some environments like forests, rough areas, rainy and foggy weathers there are lots of indentations like branches, twigs and leaves or water drops.
+Each of this bits and pieces prevents the infrared laser lights coming from lidar and creates canopies.
+Instead of a clean point cloud, lidar gets a noisy one.
+
+          We can measure the effect of the bits and pieces in the environment to accuracy of lidar measurement by calculating the gradient of depth frame around target pixel.
+
+          \begin{figure}[H]
+              \centering
+              \includegraphics[width=0.9\textwidth]{LidarViews.png}
+              \caption{Gradient of Depth Image Around Target Pixel Gives Roughness of the Environment~\cite{https://doi.org/10.48550/arxiv.2010.09076}}
+          \end{figure}
+\end{itemize}
+
+\subsubsection*{Variables That Effect Stereo Vision Camera:}
+
+\begin{itemize}
+    \item Distance:
+
+          According to viewing angles and algorithm used to extract depth information from stereo image, there is an uncertainty appears.
+And this uncertainty increases as distance of the pixel increases.
+Uncertainty reduces the confidence of the input as it increased.
+
+          \begin{figure}[H]
+              \centering
+              \includegraphics[width=0.4\textwidth]{CamDistance.png}
+              \caption{Stereo Camera Depth Error Calculation~\cite{DisparityCalculator, gallup2008variable}}
+          \end{figure}
+
+    \item Brightness:
+
+          The brightness of the environment is important for stereo vision depth estimation.
+As considering the working principles of the stereo vision depth estimation, there are two frames taken from different angles.
+First step is matching each pixel of left frame with corresponding pixel in right frame.
+This step includes computer vision algorithms like Sum of Squared Differences (SSD) and Sum of Absolute Differences (SAD).
+To perform matching properly, the frames should have well distributed histograms and good contrast values.
+If there is the brightness of frames are too high or too low, the algorithms cannot match the corresponding pixels.
+
+          We can calculate the effect of brightness by observing the histogram of the RGB frames.
+    \item Straight and no-Contrast Views:
+
+          Like the brightness, the content of the frame is also important.
+If the frame includes low contrast, straight or similar patterns, the matching algorithms won’t work properly.
+For example, If the frame includes a flat white wall, the algorithm can not distinguish pixels in different frames.
+Likewise, if the frame has a periodic pattern like damas pattern, the algorithm also not work well.
+
+          In this case we can calculate the effect of straightness by calculating gradient around target pixel in RGB frames.
+\end{itemize}
+
+According to these parameters, we will derive confidence generator functions that will calculate a confidence value for each input.
+By normalizing one of the confidence values will give us $\alpha$ value.
+
+\begin{equation}
+    \alpha = \frac{C_{LIDAR}}{C_{LIDAR} + C_{STEREO}}
+\end{equation}
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.9\textwidth]{AlphaFunction.jpeg}
+    \caption{Alpha Function for Complementary Filter Enhancement}
+\end{figure}
+
+After extracting $\alpha$ value for each pixel of depth frame, we can simply apply complementary filter.
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.9\textwidth]{CompFiltBD.jpeg}
+    \caption{Complementary Filter Block Diagram}
+\end{figure}
+
+\section{TEAM WORK}
+
+Teamwork is an essential aspect of any successful project, and this project is no exception.
+Our team is composed of multiple members with different backgrounds, expertise, and skills, working together to achieve a common goal.
+To ensure effective collaboration and communication among team members, we have chosen to use Github as our primary source control and project management platform.
+This allows us to easily share code, track changes, and collaborate on different parts of the project.
+
+In addition to Github, we are also using MATLAB, Python, and the Robot Operating System (ROS) environment for our project.
+These tools provide us with the necessary functionality and flexibility to implement and test our sensor fusion algorithm.
+MATLAB, in particular, is useful for prototyping and simulating our algorithm, while Python and ROS provide us with the ability to implement and test our algorithm on real-world systems.
+
+We also keep a logbook during the project which contains our daily progress, issues encountered, and solutions implemented.
+This logbook acts as a reference for team members and helps keep track of the project's history and progress.
+Keeping a logbook is a best practice in project management, as it helps to stay organized, keep track of progress, and document any issues that may arise.
+
+Other papers in the field also highlighted the importance of team work, communication, and collaboration in the development of sensor fusion algorithms.
+For example, in their paper on sensor fusion for mobile robots, authors underline that the development of a sensor fusion algorithm is a complex task that requires a multidisciplinary team with expertise in various fields such as robotics, control systems, and computer vision.
+They also point out that effective communication and collaboration among team members is crucial to the success of the project.
+
+According to \cite{daniilidis2001sensor}, sensor fusion is a complex task that has been extensively studied and is widely used in various applications.
+In \cite{durrant2006simultaneous}, it was stated that simultaneous localization and mapping (SLAM) is an important problem in the field of robotics that requires the integration of sensor data through sensor fusion.
+Additionally, in \cite{silva2020edge} the authors proposed to use edge preservation methods to increase the performance of depth map filtering for stereo vision, which can be useful for our sensor fusion algorithm.
+
+In conclusion, effective teamwork is crucial to the success of this project, and we are utilizing a variety of tools and best practices to ensure that our team can work together effectively.
+Our use of Github, MATLAB, Python, and the ROS environment, as well as our logbook and regular team meetings, will help us to stay organized and work effectively together to achieve our goals.
+
+\section{COMMENTS AND CONCLUSIONS}
+
+In conclusion, sensor fusion is an essential component of modern technology that is used to sense and augment the environment.
+Our proposed project aims to develop a sensor fusion algorithm for extracting depth information using data from living organisms, specifically studying the way in which zebra fish process sensor data to produce depth information.
+We plan to use a legged robot for the system and cloud computing for high computing power requirements.
+The preliminary design of the project includes using the ZED stereo camera and Velodyne 3D LiDAR as sensors and implementing a complementary filter algorithm to fuse the depth information from the two sensors, with edge preservation methods being used as an aid to the main algorithm to increase performance.
+
+Effective teamwork and collaboration among team members are crucial to the success of this project, and we are utilizing a variety of tools and best practices such as Github, MATLAB, Python, the Robot Operating System (ROS) environment, and a logbook to ensure that our team can work together effectively.
+We also have been studying the sensor fusion and alternative methodologies to implement it correctly to serve the main purpose of sensor fusion, that is reliable odometry data extraction.
+
+In future work, we plan to improve the sensor fusion algorithm by incorporating data from the zebra fish, optimize the resolution and data rates, and test the algorithm on a legged robot with cloud computing.
+Additionally, we plan to implement the sensor fusion algorithm on an autonomous SLAM robot to demonstrate the usefulness of the depth information in real-world applications.
+
+The preliminary design of our sensor fusion algorithm includes the use of the ZED stereo camera and Velodyne 3D LiDAR sensors as they are well-documented, have a large community, and are easier to use and access resources than the internal sensors of the legged robot.
+The sensor fusion algorithm chosen for the preliminary design is a complementary filter, which is a simple and low-performance algorithm, but it can be improved by using edge preservation methods to increase the sensor fusion performance.
+We have also considered the different properties, advantages, and disadvantages of the sensors to optimize the sensor fusion algorithm.
+The stereo vision camera and 3D lidar have different viewing angles, data formats, and sample rates, thus we have to match them by reducing the viewing angle of the 3D lidar, upsampling the lidar data, converting it to cartesian form, and using dynamic Papoulis-Gerchberg algorithm to augment the lidar data.
+We also aim to optimize resolution and data rates in the future work to increase the quality of the point clouds and reduce the computational effort.
+
+In addition to the previous conclusions, it is worth noting that the preliminary design of our sensor fusion algorithm was realized using the MATLAB programming environment.
+This choice was made because of the wide range of tools and functions available in MATLAB for image processing and sensor data manipulation, as well as its user-friendly interface and compatibility with other programming languages such as Python and C++.
+
+By using MATLAB, we were able to effectively implement the sensor fusion algorithm, including the complementary filter, edge preservation methods, and dynamic Papoulis-Gerchberg algorithm.
+We also were able to test the algorithm with sample data and evaluate its performance.
+
+It is important to note that while the preliminary design has been realized in MATLAB, we plan to implement the final version of the sensor fusion algorithm in an appropriate programming environment for the legged robot, such as ROS, to ensure the compatibility and efficiency of the algorithm in the real-world application.
+
 \addcontentsline{toc}{section}{REFERENCES}

### 2023-01-23 MATLAB/data_import.m
.m files added.

@@ -1,0 +1,107 @@
+% 3D Lidar point cloud
+veloReader = velodyneFileReader("lidar\300\2022-12-10-16-53-10_Velodyne-VLP-16-Data.pcap","VLP16");
+lidar_pc = readFrame(veloReader);
+
+% Stereo camera point cloud
+camera_pc = pcread("ZED\vga\point_cloud_PLY_3029_376_06-12-2022-19-40-31.ply");
+%%
+savePointCloud(lidar_pc, 'lidar_pc.mat');
+savePointCloud(camera_pc, 'camera_pc.mat');
+%%
+pcshow(lidar_pc);
+pcshow(camera_pc);
+%%
+lidar_pc_den = pcdenoise(lidar_pc);
+camera_pc_den = pcdenoise(camera_pc);
+%% 
+% The reason that your Location argument is 16xNx3 instead of 16Nx3 is likely 
+% because the Velodyne LiDAR sensor captures data in multiple "sweeps" or "scans" 
+% of the environment, each with a different vertical field of view (FOV). Each 
+% of these sweeps or scans is stored as a separate "slice" or "layer" in the 3D 
+% matrix, with each slice representing a different FOV. Therefore, the matrix 
+% is 16 layers (sweeps) by N points by 3 dimensions (x,y,z) to reflect the 16 
+% different FOVs.
+
+lidar_pts = reshape(lidar_pc_den.Location, [], 3);
+%% 
+% Camera point cloud is around 90 degree rotated along the x-axis. Hence, I 
+% re-rotate to the desired location to match the domains. I use the affine transformation 
+% to achieve this.
+% 
+% 
+
+% Rotate x-axis
+a = pi/2;
+
+A = [1 0 0 0;...
+     0 cos(a) sin(a) 0;...
+     0 -sin(a) cos(a) 0;...
+     0 0 0 1];
+
+tform = affine3d(A);
+
+camera_pc_rot = pctransform(camera_pc_den, tform);
+
+% Rotate z-axis
+a = pi/20;
+
+A = [cos(a) sin(a) 0 0;...
+     -sin(a) cos(a) 0 0;...
+     0 0 1 0;...
+     0 0 0 1];
+
+tform = affine3d(A);
+
+camera_pc_rot = pctransform(camera_pc_rot, tform);
+
+% Get Nx3 matrix
+camera_pts = camera_pc_rot.Location;
+%%
+% Visualize Lidar point cloud
[…]

### 2023-01-23 MATLAB/preliminaryDesign/main.m

@@ -1,0 +1,250 @@
+% Load point cloud data from Lidar and stereo camera
+lidar_pts = loadLidarPointCloud();
+camera_pts = loadStereoCameraPointCloud();
+
+% or Generate synthetic point cloud data
+% [lidar_pts, camera_pts] = generatePointCloudData();
+%% 
+% Firstly I plot the preprocessed an denoised point clouds on the same figure. 
+% After that, I apply the proposed sensor fusion method.
+
+% Create figure
+figure;
+
+% Visualize Lidar point cloud
+scatter3(lidar_pts(:,1), lidar_pts(:,2), lidar_pts(:,3), 'r', 'Marker', '.');
+hold on;
+
+% Visualize camera point cloud
+scatter3(camera_pts(:,1), camera_pts(:,2), camera_pts(:,3), 'b', 'Marker', '.');
+hold off;
+
+% Add legend and labels
+legend('Lidar', 'Camera');
+xlabel('X');
+ylabel('Y');
+zlabel('Z');
+%%
+% Upsample Lidar point cloud with Interpolation
+lidar_pts_upsampled = upsampleLidar(lidar_pts, camera_pts);
+
+% Create figure
+figure;
+
+% Visualize original Lidar point cloud
+scatter3(lidar_pts(:,1), lidar_pts(:,2), lidar_pts(:,3), 'k', 'Marker', '.');
+hold on;
+
+% Visualize filtered point cloud
+scatter3(lidar_pts_upsampled(:,1), lidar_pts_upsampled(:,2), lidar_pts_upsampled(:,3), 'r', 'Marker', '.');
+hold off;
+
+% Add legend and labels
+legend('Original Lidar', 'Upsampled');
+xlabel('X');
+ylabel('Y');
+zlabel('Z');
+%% 
+% The Papoulis-Gerchberg algorithm is a method for reconstructing an image from 
+% its Fourier transform magnitude and a known phase constraint. It can be used 
+% to improve the resolution of a lidar point cloud by iteratively updating the 
+% phase of the Fourier transform of the point cloud while enforcing a constraint 
+% on the known magnitude of the point cloud.
+
+% Upsample Lidar point cloud with Papoulis-Gerchberg algorithm
+N = length(camera_pts);
+M = length(lidar_pts);
+
+lidar_pts_pg = papoulis_gerchberg(lidar_pts, N/M);
+%%
[…]

### 2023-05-09 docker/zed/run.sh
docker directory organized

@@ -1,0 +1,15 @@
+echo "source /opt/ros/noetic/setup.bash" >> ~/.bashrc
+source ~/.bashrc
+
+sudo apt install -y python3-rosdep python3-rosinstall python3-rosinstall-generator python3-wstool build-essential
+
+sudo rosdep init
+rosdep update
+
+# Install Cuda
+wget https://developer.download.nvidia.com/compute/cuda/repos/ubuntu2004/x86_64/cuda-ubuntu2004.pin
+sudo mv cuda-ubuntu2004.pin /etc/apt/preferences.d/cuda-repository-pin-600
+sudo apt-key adv --fetch-keys https://developer.download.nvidia.com/compute/cuda/repos/ubuntu2004/x86_64/3bf863cc.pub
+sudo add-apt-repository "deb https://developer.download.nvidia.com/compute/cuda/repos/ubuntu2004/x86_64/ /"
+sudo apt-get update
+sudo apt-get -y install cuda

### 2023-05-09 reports/report4/main.tex
2nd term interim report added

\documentclass[12pt]{article}
\usepackage[left=1in, right=1in, top=1in, bottom=1in]{geometry}
\usepackage{graphicx}
\usepackage{url}
\usepackage{cite}
\usepackage{float}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{hyperref}
\hypersetup{
    colorlinks,
    citecolor=black,
    filecolor=black,
    linkcolor=black,
    urlcolor=black
}

\renewcommand{\familydefault}{\sfdefault}

\begin{document}
\begin{titlepage}
    \begin{center}
        \LARGE
        \textbf{ELE 402 - GRADUATION PROJECT II}

        \Large
        \textbf{INTERIM REPORT}

        \vspace{70pt}

        \textit{
            Hacettepe University \\
            Department of Electrical and Electronics Engineering
        }
    \end{center}

    \vspace{90pt}

    \large

    \textbf{Project Title:} Sensor Fusion in the Cloud \\

    \textbf{Project Group Members:} Ertuğrul Tiyek, Ahmet Yusuf Şirin

    \vspace{30pt}

    \textbf{Project Supervisor:} Asst. Prof. Dr. İsmail Uyanık \\

    \textbf{Submission Date:} 07.05.2023

    \vspace{\fill}

    \begin{center}
        \textit{SPRING 2022-2023}
    \end{center}
\end{titlepage}

\clearpage

\tableofcontents
\listoffigures
\listoftables

\clearpage

\begin{abstract}
    One of the most important aim of the modern technology is the sensing the environment. Living beings have sense organs to know the environment. They have brains to make fusion the sensor data and augment the environmental information taken from sense organs. For example, human standing in balance, we can say that they use the data coming from their eyes and ears together.

    People are studying to make this sensing and augmenting artificially, too. They use computers to achieve this. But computers cannot sense the environment directly. They need some devices called sensors. There are all kinds of sensors to measure different environmental quantities. For same quantity, there are different sensors using different technologies, have different precision and accuracy levels. Each sensor has its advantages and disadvantages. Even sensors that are serving a different purpose can help to augment same information other than their main. This is known as "sensor fusion"~\cite{enwiki:1115352853}. Just like humans, robots are use camera and inertial sensors to stand in balance.

    In our project, sensor fusion handles extracting depth information. Sensor fusion can be implemented using specific algorithms like EKF, CNN etc. according to quantity that will be calculated, number and types of sensors used, accuracy and performance needed. We are planning to develop a sensor fusion algorithm with the help of living organisms. As we mentioned before, living beings perform sensor fusion perfectly. We will be examining the zebra fish and how it processes data of its sense organs to produce depth information. Then we will try to develop similar algorithm or improve our algorithm with data taken from zebra fish.

    Depth information can be used for many applications like automatic driven vehicles, UAVs, Augmented Reality devices and 3D modeling of areas. Our concept in this project will be autonomous controlled SLAM Robot. Simultaneous localization and mapping (SLAM) is the computational problem of constructing or updating a map of an unknown environment while simultaneously keeping track of an agent's location within it~\cite{enwiki:1120084976}. We use a robot to move the system.

    Another concept we aim to use is cloud computing. Mobile computing devices may not match the performance requirements in some cases. To achieve more accurate depth information, complex control algorithms or neural networks will be used in our project which needs high computing power. As an alternative to the mobile computing devices, we may use cloud computing.

    From the point on we are studying on the sensor fusion. In this report we will briefly emphasize the legged robot usage for our final decisions. On the other hand, the concept ot the sensor fusion and the alternative methodologies to implement it correctly to serve the main purpose of sensor fusion, that is reliable odometry data extraction.
\end{abstract}

\section{INTRODUCTION}

This report presents the progress and design iterations of a fully ROS-based robot, focusing on sensor fusion and SLAM operations. The project aims to develop a robotic platform capable of autonomously navigating and mapping an environment with the help of a variety of sensors. The purpose of this report is to provide an overview of the prototype design and the design process, highlighting the key features, testing results, and evaluations from each iteration. This document is intended to serve as a guide for future work and improvements on the project.

The report is organized as follows: Section 2 introduces the prototype, including its purpose, functionalities, and key features. Section 3 details the design process, covering the different iterations carried out to refine the prototype, along with their respective testing and evaluation.

\section{PROTOTYPE}

The prototype developed for this project is a fully ROS-based robot designed to perform sensor fusion and SLAM operations. The primary purpose of building this prototype is to demonstrate the effectiveness of integrating various sensors for accurate and efficient mapping and navigation in an unknown environment. The prototype serves as a platform for testing different sensor fusion algorithms and SLAM techniques to identify the most suitable combination for real-world applications.

The prototype consists of a mobile robotic base equipped with multiple sensors, including LiDAR, camera, inertial measurement unit (IMU), and wheel encoders. These sensors work in tandem to gather data from the environment, which is then processed and combined using sensor fusion algorithms. The fused data is fed into the SLAM system, allowing the robot to construct a map of its surroundings while simultaneously keeping track of its position within that map.

Key features of the prototype include:

\begin{enumerate}
    \item ROS integration: The robot's software architecture is built on the Robot Operating System (ROS), a widely used, flexible framework for robotics development. This allows for easy integration of existing ROS packages and ensures compatibility with a wide range of sensors and algorithms.

    \item Sensor fusion: The prototype employs various sensor fusion techniques to combine the data from different sensors, resulting in a more accurate and robust representation of the environment. This helps improve the robot's overall performance in mapping and navigation tasks.

    \item SLAM implementation: The prototype integrates state-of-the-art SLAM algorithms to simultaneously localize the robot within the environment and construct a map based on the fused sensor data. This enables the robot to navigate autonomously and efficiently, even in previously unexplored areas.

    \item Modular design: The hardware and software components of the prototype are designed to be modular, allowing for easy modification, replacement, and upgrading of individual elements. This ensures that the robot can be easily adapted to accommodate new sensors or algorithms as they become available.
\end{enumerate}

LOAM (Lidar Odometry and Mapping) is a state-of-the-art algorithm for simultaneous localization and mapping (SLAM) that leverages the advantages of lidar sensors. The LOAM algorithm is designed to work in real-time and has been widely adopted in various robotic applications, including autonomous vehicles and drones, due to its robust performance and high accuracy.

The main idea behind LOAM is to separate the point cloud data generated by the lidar sensor into two components: ground points and edge points. Ground points represent the flat surfaces of the environment, while edge points correspond to the sharp features such as corners and edges. By analyzing the relative motion of these two sets of points, LOAM can estimate the robot's pose and build a detailed map of the environment.

LOAM is particularly well-suited for robotic platforms equipped with high-resolution lidar sensors, as it can handle large amounts of data efficiently and maintain accurate localization even in complex environments. Integrating LOAM into the prototype enables the robot to perform SLAM tasks with high precision, making it an ideal choice for the sensor fusion system.

\begin{figure}[h]
    \centering
    \includegraphics[width=0.9\textwidth]{loam_ros_graph.png}
    \caption{Local Host LOAM ROS Node/Topic Schematic}
\end{figure}

The Robot Operating System (ROS) is a flexible and powerful framework for developing robotic applications, providing a wide range of libraries, tools, and software components for various robotic tasks. Running ROS on a Raspberry Pi (RPI) is an efficient and cost-effective way to create a compact, versatile robotic system that can take advantage of ROS's capabilities.

By utilizing an RPI as the central processing unit for the robot, the system can benefit from the RPI's low power consumption, small form factor, and extensive community support. This allows for a lightweight, portable robotic platform that can be easily extended and customized according to specific application requirements. Moreover, the RPI can run a Linux-based operating system, which is compatible with ROS, making it an ideal choice for integrating ROS into the robot's control system.

Installing and running ROS on an RPI involves setting up the necessary dependencies, configuring the ROS environment, and deploying the desired ROS packages. Once the ROS system is up and running on the RPI, the robot can utilize various ROS functionalities, such as communication between different nodes, sensor data processing, and control algorithms implementation.

Integrating ROS with the RPI enables seamless communication with other hardware components, such as sensors and actuators, through the use of ROS topics, services, and actions. This allows for efficient sensor fusion, SLAM operations, and overall system coordination. The combination of ROS and RPI offers a powerful, compact solution for developing advanced robotic applications while maintaining a low-cost and accessible platform.

\begin{figure}[h]
    \centering
    \includegraphics[width=0.9\textwidth]{rpi_ros_graph.png}
    \caption{Legged Robot RPi ROS Node/Topic Schematic}
\end{figure}

\section{DESIGN PROCESS}

The design process of the prototype followed an iterative approach, involving multiple stages of development, testing, and evaluation. This methodology allowed the team to identify shortcomings in the initial design, implement modifications, and continuously improve the prototype's performance. Each iteration focused on addressing specific challenges, refining the sensor fusion techniques, and optimizing the SLAM implementation. In the following subsections, we present the steps of the design process performed in constructing the prototype, including the objectives, testing, results, and evaluation for each iteration.

    \subsection{ITERATION 1: INITIAL SENSOR INTEGRATION AND ROS SETUP}

    The first iteration aimed to establish a solid foundation for the project by integrating the different sensors and setting up the ROS framework for the robot. The main objectives of this iteration were:

    \begin{enumerate}
        \item Selection and integration of appropriate sensors (LiDAR, camera, IMU, and wheel encoders) for the robot.
        \item Installation and configuration of ROS on the robot's embedded computer system.
        \item Setup of the communication infrastructure between the sensors, ROS nodes, and the robot's control system.
    \end{enumerate}

    In addition to the aforementioned advantages, the integration of the Raspberry Pi (RPI) with the Robot Operating System (ROS) allows for enhanced communication capabilities with micro-ROS systems. Micro-ROS is a lightweight version of ROS specifically designed for microcontrollers, enabling resource-constrained devices to become part of the ROS ecosystem. By incorporating micro-ROS into the system, it becomes possible to extend the functionality and modularity of the robotic platform even further.

    The RPI serves as the central communication hub between the ROS nodes and the micro-ROS system. It manages the flow of data between the different subsystems and ensures seamless interaction among the various components of the robot. To facilitate this communication, the RPI is connected to a network switch, which enables efficient data transmission and distribution throughout the system.

    The network switch plays a crucial role in the overall performance and reliability of the robotic system. It ensures that all components, including the RPI, micro-ROS nodes, and other connected devices, can exchange information effectively and maintain synchronized operation. By leveraging a network switch in conjunction with the RPI, the system can handle multiple data streams, balance network traffic, and reduce latency, thereby improving the overall performance of the robot.

    In summary, the combination of the Raspberry Pi, ROS, micro-ROS, and a network switch provides a powerful and versatile robotic platform. The RPI acts as the central processing unit and communication hub, allowing for efficient coordination between the ROS nodes and micro-ROS systems. The use of a network switch ensures seamless data exchange among the various components of the robot, resulting in a robust, scalable, and high-performance robotic system capable of tackling a wide range of applications.

    \begin{figure}[H]
        \centering
        \includegraphics[width=0.9\textwidth]{Go1CommFram_E.png}    
        \caption{Infrastructure of the Unitree Go1 Robot}
    \end{figure}

        \subsubsection{TESTING AND RESULTS}

        The robot's hardware and software components were assembled, and the integration of sensors with the ROS framework was tested. The communication between the sensors and ROS nodes was verified using the provided visualization tools (such as RViz) to ensure proper data flow. Preliminary tests were conducted to evaluate the robot's ability to gather data from the environment using the sensors.

        \subsubsection{EVALUATION}

        The results from the first iteration revealed the successful integration of sensors and the establishment of the ROS environment. However, some issues were identified, such as occasional data loss due to communication bottlenecks and the need for better synchronization between the sensors. These issues were noted and addressed in the subsequent iterations.

    \subsection{ITERATION 2: SENSOR FUSION AND SLAM IMPLEMENTATION}

    In the second iteration, the focus was on incorporating sensor fusion techniques to combine the data from the different sensors and implementing SLAM algorithms to enable mapping and localization capabilities. The main objectives of this iteration were:

    \begin{enumerate}
        \item Development and implementation of sensor fusion algorithms to combine data from the LiDAR, camera, IMU, and wheel encoders.
        \item Selection and integration of a suitable SLAM algorithm for mapping and localization.
        \item Evaluation of the robot's performance in mapping and navigation tasks using the fused sensor data and SLAM system.
    \end{enumerate}

        \subsubsection{TESTING AND RESULTS}

        The sensor fusion algorithms were tested to ensure the proper combination of data from the different sensors. The robot was then deployed in a controlled environment to evaluate its mapping and navigation capabilities using the SLAM system. The resulting maps were compared to the ground truth to assess the accuracy of the robot's localization and mapping.

        \subsubsection{EVALUATION}

        The results from the second iteration showed significant improvement in the robot's mapping and localization performance. The sensor fusion algorithms effectively combined the data from the different sensors, resulting in a more accurate and robust representation of the environment. However, some challenges remained, such as occasional drift in the robot's position estimation and the need for further optimization of the SLAM algorithm. These issues were addressed in the subsequent iterations.

    \subsubsection{ITERATION 3: IMPROVING SENSOR SYNCHRONIZATION AND ROBUSTNESS}

    In the third iteration, the team focused on enhancing sensor synchronization and overall system robustness to improve the robot's mapping and localization capabilities. The main objectives of this iteration were:

    \begin{enumerate}
        \item Refine sensor synchronization to minimize data loss and inconsistencies between sensor readings.
        \item Implement error handling and recovery mechanisms to increase the system's robustness.
        \item Optimize the SLAM algorithm to reduce drift and improve localization accuracy.
    \end{enumerate}

        \subsubsection{TESTING AND RESULTS}

        The improved sensor synchronization was tested by analyzing the consistency of sensor data and verifying the data alignment using visualization tools. The robot was then deployed in a controlled environment, including areas with challenging conditions, such as uneven terrain, dynamic obstacles, and poor lighting. The robot's performance in mapping, localization, and navigation tasks was assessed, taking into consideration the implemented error handling and recovery mechanisms.

        \subsubsection{EVALUATION}

        The results of the third iteration showed a notable improvement in sensor synchronization and overall system robustness. The robot was able to handle challenging environmental conditions more effectively and recover from errors more efficiently. However, some areas for further optimization were identified, such as refining the SLAM algorithm for better performance in dynamic environments and enhancing the robot's ability to cope with sensor noise.

    \subsection{ITERATION 4: DYNAMIC ENVIRONMENT HANDLING AND NOISE REDUCTION}

    The fourth iteration aimed to address the remaining challenges by improving the robot's performance in dynamic environments and reducing the impact of sensor noise. The main objectives of this iteration were:

    \begin{enumerate}
        \item Adapt the SLAM algorithm to better handle dynamic environments and moving obstacles.
        \item Implement noise filtering techniques to reduce the impact of sensor noise on the robot's mapping and localization performance.
        \item Evaluate the overall performance of the robot, including its robustness, adaptability, and accuracy in various environmental conditions.
    \end{enumerate}

        \subsubsection{TESTING AND RESULTS}

        The modified SLAM algorithm and noise filtering techniques were tested in multiple environments, including static and dynamic settings. The robot was deployed in a controlled environment with moving obstacles and varying lighting conditions to assess its ability to adapt and maintain accurate mapping and localization performance.

        \subsubsection{EVALUATION}

        The results from the fourth iteration demonstrated the robot's improved performance in dynamic environments and its enhanced ability to handle sensor noise. The robot was able to maintain accurate mapping and localization despite the presence of moving obstacles and varying environmental conditions. The noise filtering techniques effectively reduced the impact of sensor noise on the robot's performance. With these improvements, the prototype was considered ready for further real-world testing and potential future applications.

\addcontentsline{toc}{section}{REFERENCES}
\bibliographystyle{plain}
\bibliography{refs}
\end{document}

### 2023-07-18 reports/report5/main.tex
Report 5 added and logbook updated

--- 2023-01-22 reports/report3/main.tex
@@
         \LARGE
-        \textbf{ELE 401 - GRADUATION PROJECT I}
+        \textbf{ELE 402 - GRADUATION PROJECT II}
 
@@
 
-    \textbf{Submission Date:} 22.01.2023
+    \textbf{Submission Date:} 18.06.2023
 
@@
 \begin{abstract}
-    One of the most important aim of the modern technology is the sensing the environment.
-Living beings have sense organs to know the environment.
-They have brains to make fusion the sensor data and augment the environmental information taken from sense organs.
-For example, human standing in balance, we can say that they use the data coming from their eyes and ears together.
-
-    People are studying to make this sensing and augmenting artificially, too.
-They use computers to achieve this.
-But computers cannot sense the environment directly.
-They need some devices called sensors.
-There are all kinds of sensors to measure different environmental quantities.
-For same quantity, there are different sensors using different technologies, have different precision and accuracy levels.
-Each sensor has its advantages and disadvantages.
-Even sensors that are serving a different purpose can help to augment same information other than their main.
-This is known as "sensor fusion"~\cite{enwiki:1115352853}.
-Just like humans, robots are use camera and inertial sensors to stand in balance.
-
-    In our project, sensor fusion handles extracting depth information.
-Sensor fusion can be implemented using specific algorithms like EKF, CNN etc.
-according to quantity that will be calculated, number and types of sensors used, accuracy and performance needed.
-We are planning to develop a sensor fusion algorithm with the help of living organisms.
-As we mentioned before, living beings perform sensor fusion perfectly.
-We will be examining the zebra fish and how it processes data of its sense organs to produce depth information.
-Then we will try to develop similar algorithm or improve our algorithm with data taken from zebra fish.
-
-    Depth information can be used for many applications like automatic driven vehicles, UAVs, Augmented Reality devices and 3D modeling of areas.
-Our concept in this project will be autonomous controlled SLAM Robot.
-Simultaneous localization and mapping (SLAM) is the computational problem of constructing or updating a map of an unknown environment while simultaneously keeping track of an agent's location within it~\cite{enwiki:1120084976}.
-We use a robot to move the system.
-
-    Another concept we aim to use is cloud computing.
-Mobile computing devices may not match the performance requirements in some cases.
-To achieve more accurate depth information, complex control algorithms or neural networks will be used in our project which needs high computing power.
-As an alternative to the mobile computing devices, we may use cloud computing.
-
-    From the point on we are studying on the sensor fusion.
-In this report we will briefly emphasize the legged robot usage for our final decisions.
-On the other hand, the concept ot the sensor fusion and the alternative methodologies to implement it correctly to serve the main purpose of sensor fusion, that is reliable odometry data extraction.
+    Our project centers around sensor fusion, a critical aspect of artificially perceiving and enhancing the environment.
+By leveraging sensors, computers can acquire data on diverse environmental parameters.
+While each sensor has its strengths and weaknesses, even sensors designed for different purposes can contribute to augmenting the same information.
+Our project specifically focuses on extracting depth information through sensor fusion techniques, including EKF, CNN, and a novel iterative algorithm inspired by Papoulis Gerchberg.
+This algorithm operates in the frequency domain, reducing the processing workload compared to per-pixel control algorithms.
+The depth information obtained through our approach holds tremendous potential for a wide range of applications, including autonomous vehicles, UAVs, augmented reality devices, and 3D modeling.
+Additionally, we explore the integration of cloud computing to meet the computational demands of complex control algorithms or neural networks, ensuring accurate depth estimation.
+Our work places particular emphasis on the utilization of legged robots and highlights the significance of reliable odometry data extraction for simultaneous localization and mapping (SLAM).
 \end{abstract}
@@
 
-One of the most important aim of the modern technology is the sensing the environment.
-Living beings have sense organs to know the environment.
-They have brains to make fusion the sensor data and augment the environmental information taken from sense organs.
-For example, human standing in balance, we can say that they use the data coming from their eyes and ears together.
-
-People are studying to make this sensing and augmenting artificially, too.
-They use computers to achieve this.
-But computers cannot sense the environment directly.
-They need some devices called sensors.
-There are all kinds of sensors to measure different environmental quantities.
-For same quantity, there are different sensors using different technologies, have different precision and accuracy levels.
-Each sensor has its advantages and disadvantages.
-Even sensors that are serving a different purpose can help to augment same information other than their main.
-This is known as "sensor fusion"~\cite{enwiki:1115352853}.
-Just like humans, robots are use camera and inertial sensors to stand in balance.
-
-In our project, sensor fusion handles extracting depth information.
-Sensor fusion can be implemented using specific algorithms like EKF, CNN etc.
-according to quantity that will be calculated, number and types of sensors used, accuracy and performance needed.
-We are planning to develop a sensor fusion algorithm with the help of living organisms.
-As we mentioned before, living beings perform sensor fusion perfectly.
-We will be examining the zebra fish and how it processes data of its sense organs to produce depth information.
-Then we will try to develop similar algorithm or improve our algorithm with data taken from zebra fish.
-
-Depth information can be used for many applications like automatic driven vehicles, UAVs, Augmented Reality devices and 3D modeling of areas.
-Our concept in this project will be autonomous controlled SLAM Robot.
-Simultaneous localization and mapping (SLAM) is the computational problem of constructing or updating a map of an unknown environment while simultaneously keeping track of an agent's location within it~\cite{enwiki:1120084976}.
-We use a robot to move the system.
-
-Another concept we aim to use is cloud computing.
-Mobile computing devices may not match the performance requirements in some cases.
-To achieve more accurate depth information, complex control algorithms or neural networks will be used in our project which needs high computing power.
-As an alternative to the mobile computing devices, we may use cloud computing.
+This report provides an overview of the completed project, which focuses on sensor fusion and SLAM operations within a fully ROS-based robot.
+The project aimed to develop an autonomous robotic platform capable of mapping and navigating its environment using a diverse set of sensors.
+The previous interim report outlined the prototype design, iterations, and evaluation results.
+In this final report, we will discuss the progress made since the interim report, including the additional work carried out.
+One significant accomplishment is the development of our sensor fusion algorithm, which has been further accelerated through the implementation of CUDA libraries.
+Cloud computing was also integrated into the project, and a modular Docker container structure was adopted to enhance system independence and scalability.
+Additionally, the SLAM process was performed using various algorithms and real-time as well as recorded datasets.
+Overall, this report presents a comprehensive overview of the project's accomplishments, including the development of the sensor fusion algorithm, cloud computing integration, SLAM implementation, and adherence to engineering standards and design constraints.
+The subsequent sections will delve into the specific details of each aspect, providing a comprehensive understanding of the completed project.
 
@@
 
-This project aims to construct an unmanned ground vehicle that senses the environment and decides its movements.
-As mentioned in the "Introduction" section, the sensor fusion and mapping algorithms take a lot of space in this project.
-In addition to that, there are other parts that need to be implemented like robot controls, data acquisition from sensors and cameras, decision making algorithms etc.
+Our project focuses on the development of a fully autonomous robotic platform that utilizes sensor fusion and SLAM techniques to navigate and map its environment.
+The primary objective is to create a robust system capable of perceiving and augmenting the surrounding environment by combining data from various sensors.
+By leveraging sensor fusion algorithms such as EKF, CNN, and an iterative approach inspired by Papoulis Gerchberg, we aim to extract accurate depth information.
 
-    \subsubsection*{A. Hardware Architecture of the Project}
+The extracted depth information holds significant potential for a wide range of applications, including autonomous vehicles, unmanned aerial vehicles (UAVs), augmented reality devices, and 3D modeling.
+To handle the computational demands of complex control algorithms and neural networks for precise depth estimation, we explore the integration of cloud computing.
+This allows us to leverage high computing power and achieve enhanced accuracy.
 
-    The hardware component choices are explained with consideration of advantages and disadvantages in the section 3.2 Design Constraints. Here we have our main components and specifications of them:
+Our project incorporates the use of legged robots, which offer increased mobility and adaptability in diverse environments.
+Additionally, reliable odometry data extraction is emphasized, as it plays a crucial role in simultaneous localization and mapping (SLAM).
+By ensuring accurate odometry, we can effectively construct and update a map of the unknown environment while keeping track of the robot's location within it.
 
-    \begin{enumerate}
-        \item \textbf{Robot – Unitree Go1 Edu} \\
-        This robot is a ready-to-go quadruped robot with built-in 5 Sets of Fish-eye Stereo Depth Cameras, Ai Post-processing, 3 Sets of Hypersonic Sensors.
-There are three Nvidia Jetson Nano, one Raspberry Pi 4, one STM32 Microcontroller devices inside of the Unitree Go1 Edu.
-These devices located in the robot to perform certain tasks.
-STM32 Microcontroller drives joint motors of the robot.
-Raspberry Pi 4 is used for both communicate with user and send commands to STM32.
-Two of Nvidia Jetson Nano 4GB kits and one Nvidia Jetson Nano 2 GB development kit are used for getting data from camera and sensors, processing raw data to extract dept information of stereo cameras.
+Modularity, system independence, and scalability are crucial considerations for our project.
+To achieve these goals, we have implemented a Docker container structure that facilitates easy integration and maintenance of the system components.
+This approach allows for seamless expansion and upgrades as needed.
 
-        \begin{figure}[H]
-            \centering
-            \includegraphics[width=0.9\textwidth]{Go1CommFram_E.png}    
-            \caption{Infrastructure of the Unitree Go1 Robot}
-        \end{figure}
-
-        The manufacturer provides two Software Development Kits. "unitree\_legged\_sdk” for Controlling the robot, “unitree\_camera\_sdk” for getting camera data and calculating depth information.
-
-        \item \textbf{Stereo Cameras – Unitree Camera } \\
-        There are 5 sets of fish-eye stereo depth cameras located on different important locations of the robot. The known features are:
-
-        \begin{itemize}
-            \item Lens Angle $\approx$ 150x170° 
-            \item HD (1280x720) , 30fps video streaming 
-            \item Raw Frame 
-            \item Rectangular Frame 
-            \item Depth Frame 
-            \item Point Cloud Image 
-            \item Provided SDK 
-        \end{itemize}
-
-        \item \textbf{3D LIDAR – Velodyne 80-VLP-16-A~\cite{VelodyneLiDAR}} \\
-        The VLP-16 has a range of 100m, and the sensor's low power consumption (~8W), light weight (830 grams), compact footprint (~Ø103mm x 72mm), and dual return capability make it ideal for UAVs and other mobile applications.
-
-        Velodyne’s LiDAR Puck supports 16 channels, ~300,000 points/sec, a 360° horizontal field of view and a 30° vertical field of view, with +/- 15° up and down.
-The Velodyne LiDAR Puck does not have visible rotating parts, making it highly resilient in challenging environments (Rated IP67).
-
-        Automotive Mapping UAV Security Robotics Automation
-
-        The module provides up to 0.3 million points/second over 100Mbps ethernet cable and UDP interface.
-UDP packages contain distances, calibrated reflectivity information, rotation angles, synchronized time stamps (microseconds resolution).
-
-        \item \textbf{Computation Unit – Nvidia Jetson Nano 4GB} \\
-        Unitree Go1 Edu has three Nvidia Jetson Nano Development kits inside it to perform tasks like point cloud, human skeleton tracking etc.
-We don’t use most of the features robot has, so we decided on using one of the Nvidia Jetson Nano 4GB models in the robot with reducing its
-        workload.
-
-        Here is some specifications of Nvidia Jetson Nano 4GB~\cite{JatsonNano}:
-
-        \begin{itemize}
-            \item GPU		:	128-core Maxwell 
-            \item CPU		:	Quad-core ARM A57 @ 1.43 GHz 
-            \item Memory	:	4 GB 64-bit LPDDR4 25.6 GB/s 
-            \item Storage	:	microSD (not included) 
-            \item Camera	:	2x MIPI CSI-2 DPHY lanes 
-            \item Connectivity	:	Gigabit Ethernet, M.2 Key E 
-            \item Display	:	HDMI and display port 
-            \item USB		:	4x USB 3.0, USB 2.0 Micro-B 
-            \item Others		:	GPIO, I2C, I2S, SPI, UART 
-        \end{itemize}
-
-        As mentioned before, one of the aims of this project is performing heavy computation on cloud.
-The point of implementing the algorithm on a mobile platform is to keep things simple in the first step.
-That provides us both simple implementation and ability to make a comparison between mobile computing and cloud computing in terms of performance, accuracy and cost.
-
-    \end{enumerate}
-
-    \subsubsection*{B. Software Architecture of the Project}
-
-    Laying on the keep it simple stupid (KISS) principle, we separate our project into smaller and independent parts of development.
-
-    \begin{enumerate}
-        \item \textbf{Controlling the Robot Movement and Getting Sensor Data Using Software} \\
-        There is a SDK provided by Unitree to control the robot with software called unitree\_legged\_sdk.
-We used the python programming language with unitree\_legged\_sdk.
-The details about choice of programming language are explained in the section 3.2 design constraints.
-
-        Connection between robot and controller computer established via UDP protocol.
-There are two options that can provide UDP connection.
-One is the ethernet port of switch that connects all devices in the robot and the other is Wi-Fi access point that connected to the Raspberry Pi 4 in the robot.
-All devices are using a network mask 255.255.255.0, so it is important to use a static IP on the user computer.
-
-        \begin{figure}[H]
-            \centering
-            \includegraphics[width=0.8\textwidth]{UserRobotConnection.png}
-            \caption{User Robot Connection: Network Mask Settings}
-        \end{figure}
-
-        \begin{figure}[H]
-            \centering
-            \includegraphics[width=0.8\textwidth]{UserRPiConn.png}
-            \caption{User Robot Connection: IP of Raspberry Pi}
-        \end{figure}
-
-        Unitree\_legged\_sdk allows both low-level and high-level controls.
-Low-level control allows user to adjust each joint manually.
-High-level control provides user to control robot with high-level commands like walk, turn, jump etc.
-We will be using high-level control for this project.
-
-        SDK has udp class that provides us an object constructs a UDP connection with Raspberry Pi of Robot. UDP object has functions to receive, send, decode and encode udp packages.
-
-        Another class included in the SDK is HighState.
-HighState objects store state information of the robot.
-It is updated using udp.receive(HighState) function.
-HighState objects includes IMU measurement information, odometer information, 2D speed information, yaw speed information, position information, body height information etc.
-
-        To send the commands to the robot, there is a class called CMD in the SDK.
-CMD object is a command package that tell the robot what to do.
-CMD package consists of command mode, velocity, yaw speed, target position, required body height variables.
-Then udp.SetSend(cmd) redyes the cmd object  with encoding it to be sent.
-Finally the udp object sends the command to the Raspberry Pi using udp.send().
-
-        Raspberry Pi 4 also uses unitree\_legged\_sdk to receive commands from and send state information to the user computer.
-It also communicates with the STM32 Microcontroller which controls all joint motors of the robot.
-
-        \item \textbf{Getting RGB and Depth Frames from Stereo Cameras (unitree\_camera\_sdk)} \\
-        As mentioned before, there are 5 sets of fish-eye stereo depth camera units to both provide an RGB frame and depth frame around the robot.
-
-        \begin{figure}[H]
-            \centering
-            \includegraphics[width=0.8\textwidth]{RobotComponentLocations.jpeg}
-            \caption{Legged Robot Sensors and Other Components}
-        \end{figure}
-
-        \begin{figure}[H]
-            \centering
-            \includegraphics[width=0.8\textwidth]{RobotFront.png}
-            \caption{Legged Robot Head Supersensory}
-        \end{figure}
-
-        Unitree provides unitree\_camera\_sdk to access camera data of the robot. The SDK allows getting raw frames, rectangular frames, depth frames and point cloud frames.
-
-        SDK is based on GStreamer framework and works on Nvidia Jetson Nano devices.
-GStreamer is a library for constructing graphs of media-handling components.
-Applications can take advantage of advances in codec and filter technology transparently.
-Developers can add new codecs and filters by writing a simple plugin with a clean, gener~\cite{gstreamer}.
-GStreamer creates a media pipeline that allows developers to implement their applications simple and modular.
-
-        unitree\_camera\_sdk creates a pipeline on GStreamer and a virtual video device for each camera connected to the Jetson Nano.
-Developer can easily read camera information, raw frames, rectangular frames, depth frames and point cloud frames using SDK in their apps.
-
-        \begin{figure}[H]
-            \centering
-            \includegraphics[width=\textwidth]{RawStereo.png}
-            \caption{Raw Frame}
-        \end{figure}
-
-        \begin{figure}[H]
-            \centering
-            \begin{subfigure}[b]{0.4\textwidth}
-                \centering
-                \includegraphics[width=\textwidth]{StereoRectFrame.png}
-                \caption{Rectangular Frame}
-            \end{subfigure}
-            \hfill
-            \begin{subfigure}[b]{0.4\textwidth}
-                \centering
-                \includegraphics[width=\textwidth]{StereoDeptFrame.png}
-                \caption{Depth Frame}
-            \end{subfigure}
-            \caption{Worked Frames}
-        \end{figure}
-
-        Robot has some built in features like human skeleton tracking that uses cameras automatically.
-There are some services implemented to run these applications automatically while system is starting up.
-Developer should be whether kill these processes after startup or edit the services not to run processes that uses cameras.
-
-        Dept information is calculated from stereo camera using stereo vision algorithms.
-First step is matching each pixel of left frame with corresponding pixel in right frame.
-This step includes computer vision algorithms like Sum of Squared Differences (SSD) and Sum of Absolute Differences (SAD).
-
-        \begin{figure}[H]
-            \centering
-            \includegraphics[width=0.8\textwidth]{SSDSAD.png}
-            \caption{Sum of Squared Differences (SSD) and Sum of Absolute Differences (SAD)}
-        \end{figure}
-
-        After every point matched, distance of every point can be calculated using basic geometric calculations like Thales Theorem.
-
-        \begin{figure}[H]
-            \centering
-            \includegraphics[width=0.8\textwidth]{Thales.png}
-            \caption{Thales Theorem}
-        \end{figure}
-
-        Stereo cameras provide three dimensional images without using moving parts and high costs.
-But they have disadvantages as they have advantages.
-Each camera has lenses to converge and diverge the light coming from different angles and make them fall onto the sensor.
-That means each camera has a visual angle.
-Other than that visual angle cannot enter the frame.
-As we think of the closest points of the baseline of the stereo camera that enters visual angle, it is hard to both match these points in both frames and calculate their distances from camera.
-Figure shows disparity with respect to distance of the point from baseline of the stereo camera.
-
-        \begin{figure}[H]
-            \centering
-            \includegraphics[width=0.6\textwidth]{DisparityOnStereo.png}
-            \caption{Disparity with respect to Distance}
-        \end{figure}
-
-        \item \textbf{Developing Sensor Fusion Algorithm for 3D LIDAR and Stereo Cameras}
-        \item \textbf{Simultaneously Localization and Mapping (SLAM) }
-        \item \textbf{Automating SLAM with Decision Algorithm }
-        \item \textbf{Moving Sensor Fusion Algorithm to Cloud}
-        \item \textbf{Further Applications Like Swarm Robots with Centralized Cloud System }
-    \end{enumerate}
-
-    To accomplish this much work successfully the proof-of-concept parts should be made in a professional manner.
-In our earlier course work we acquired these experiences by the leading of our course supervisors.
-Although we will only implement software level designs, the understanding of mathematics and other engineering knowledge cannot be denied.
-Since we always advice with our supervisor.
-
-    \begin{table}[H]
-        \centering
-        \includegraphics[width=0.65\textwidth]{GanttChart.png}
-        \caption{Gantt Chart of Processes}
-    \end{table}
+Throughout the project, we have made significant advancements in sensor fusion, algorithm development, cloud computing integration, and SLAM implementation.
+The combination of these elements forms a comprehensive robotic platform capable of autonomously navigating, perceiving, and mapping its surroundings.
+Our project not only demonstrates the practical application of sensor fusion and SLAM but also provides valuable insights and a foundation for future developments in the field.
 
@@
 
-The design constraints should be identified and some discussion on how these apply to the design project and their realization are to be included in this section.
-You may refer to the list of some realistic design constraints that can be found in the Project Design Constraints document.
-Other constraints can be identified and discussed, if applicable.
-
 \subsection{ENGINEERING STANDARDS}
 
-    We use the IEEE standards in our project. From the related pages~\cite{enwiki:1120084976} we found the good to have rules as follows:
+Several engineering standards were followed throughout the project to ensure efficient and reliable communication, interoperability, and data transmission between the various components of the robotic system.
+These standards include:
 
-    IEEE 829-2008, also known as the 829 Standard for Software and System Test Documentation, was an IEEE standard that specified the form of a set of documents for use in eight defined stages of software testing and system testing, each stage potentially producing its own separate type of document.
-The standard specified the format of these documents, but did not stipulate whether they must all be produced, nor did it include any criteria regarding adequate content for these documents.
-These were a matter of judgment outside the purview of the standard~\cite{enwiki:1105776890}.
-There will be a lot of test cases for the design.
-Since the system is complex, the test scenarios should be documented with known standard in order to maintenance independently from project contributors.
+\begin{itemize}
+    \item TCP/IP (Transmission Control Protocol/Internet Protocol):
 
-    A software requirements specification (SRS) is a description of a software system to be developed.
-It is modeled after business requirements specification (CONOPS).
-The software requirements specification lays out functional and non-functional requirements, and it may include a set of use cases that describe user interactions that the software must provide to the user for perfect interaction~\cite{enwiki:1117244157}.
-Since we have such a software system will be developed, it should lay on some standards to make the codes and all implementations reusable and modular.
+          The SSH (Secure Shell) protocol, based on TCP/IP, was used for secure communication between internal computers of the robot.
+It provides encrypted data transmission, ensuring confidentiality and integrity.
 
-    A software design description (a.k.a.
-software design document or SDD; just design document; also Software Design Specification) is a representation of a software design that is to be used for recording design information, addressing various design concerns, and communicating that information to the design’s stakeholders~\cite{enwiki:1105986397}.
-Similarly, the developed system should have standardized documentation as we said in earlier two standards.
-The three keywords for a good software design are: modular, reusable, and flexible.
+    \item UDP (User Datagram Protocol):
 
-    Of course, as the project progresses, it will be important that we comply with more detailed standards. We will continue to mention these in future reports.
+          UDP was utilized for transmitting Lidar data. UDP is a lightweight and connectionless protocol, well-suited for real-time applications where low latency is crucial.
+
+    \item ROS (Robot Operating System):
+
+          ROS was employed as the communication framework for transmitting stereo camera information and facilitating communication between different modules of the system.
+ROS provides a standardized messaging protocol, allowing seamless integration of various sensors and modules.
+
+    \item Wi-Fi 802.11.ax:
+
+          The Wi-Fi 802.11.ax standard was utilized for wireless communication.
+This standard offers high-speed, reliable wireless connectivity, enabling communication between the robotic platform and external devices.
+\end{itemize}
+
+Adhering to these engineering standards ensures compatibility, efficient data transmission, and secure communication within the robotic system.
+By employing TCP/IP, UDP, ROS, and Wi-Fi 802.11.ax, the project meets industry best practices for communication protocols, enabling seamless interaction between components and modules of the robot.
 
@@
 
+\section{PROTOTYPE}
+
+The prototype developed for this project is a fully ROS-based robot designed to perform sensor fusion and SLAM operations.
+The primary purpose of building this prototype is to demonstrate the effectiveness of integrating various sensors for accurate and efficient mapping and navigation in an unknown environment.
+The prototype serves as a platform for testing different sensor fusion algorithms and SLAM techniques to identify the most suitable combination for real-world applications.
+
+The prototype consists of a mobile robotic base equipped with multiple sensors, including LiDAR, camera, inertial measurement unit (IMU), and wheel encoders.
+These sensors work in tandem to gather data from the environment, which is then processed and combined using sensor fusion algorithms.
+The fused data is fed into the SLAM system, allowing the robot to construct a map of its surroundings while simultaneously keeping track of its position within that map.
+
+Key features of the prototype include:
+
+\begin{enumerate}
+    \item ROS integration: The robot's software architecture is built on the Robot Operating System (ROS), a widely used, flexible framework for robotics development.
+This allows for easy integration of existing ROS packages and ensures compatibility with a wide range of sensors and algorithms.
+
+    \item Sensor fusion: The prototype employs various sensor fusion techniques to combine the data from different sensors, resulting in a more accurate and robust representation of the environment.
+This helps improve the robot's overall performance in mapping and navigation tasks.
+
+    \item SLAM implementation: The prototype integrates state-of-the-art SLAM algorithms to simultaneously localize the robot within the environment and construct a map based on the fused sensor data.
+This enables the robot to navigate autonomously and efficiently, even in previously unexplored areas.
+
+    \item Modular design: The hardware and software components of the prototype are designed to be modular, allowing for easy modification, replacement, and upgrading of individual elements.
+This ensures that the robot can be easily adapted to accommodate new sensors or algorithms as they become available.
+\end{enumerate}
+
+LOAM (Lidar Odometry and Mapping) is a state-of-the-art algorithm for simultaneous localization and mapping (SLAM) that leverages the advantages of lidar sensors.
+The LOAM algorithm is designed to work in real-time and has been widely adopted in various robotic applications, including autonomous vehicles and drones, due to its robust performance and high accuracy.
+
+The main idea behind LOAM is to separate the point cloud data generated by the lidar sensor into two components: ground points and edge points.
+Ground points represent the flat surfaces of the environment, while edge points correspond to the sharp features such as corners and edges.
+By analyzing the relative motion of these two sets of points, LOAM can estimate the robot's pose and build a detailed map of the environment.
+
+LOAM is particularly well-suited for robotic platforms equipped with high-resolution lidar sensors, as it can handle large amounts of data efficiently and maintain accurate localization even in complex environments.
+Integrating LOAM into the prototype enables the robot to perform SLAM tasks with high precision, making it an ideal choice for the sensor fusion system.
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=0.9\textwidth]{loam_ros_graph.png}
+    \caption{Local Host LOAM ROS Node/Topic Schematic}
+\end{figure}
+
+The Robot Operating System (ROS) is a flexible and powerful framework for developing robotic applications, providing a wide range of libraries, tools, and software components for various robotic tasks.
+Running ROS on a Raspberry Pi (RPI) is an efficient and cost-effective way to create a compact, versatile robotic system that can take advantage of ROS's capabilities.
+
+By utilizing an RPI as the central processing unit for the robot, the system can benefit from the RPI's low power consumption, small form factor, and extensive community support.
+This allows for a lightweight, portable robotic platform that can be easily extended and customized according to specific application requirements.
+Moreover, the RPI can run a Linux-based operating system, which is compatible with ROS, making it an ideal choice for integrating ROS into the robot's control system.
+
+Installing and running ROS on an RPI involves setting up the necessary dependencies, configuring the ROS environment, and deploying the desired ROS packages.
+Once the ROS system is up and running on the RPI, the robot can utilize various ROS functionalities, such as communication between different nodes, sensor data processing, and control algorithms implementation.
+
+Integrating ROS with the RPI enables seamless communication with other hardware components, such as sensors and actuators, through the use of ROS topics, services, and actions.
+This allows for efficient sensor fusion, SLAM operations, and overall system coordination.
+The combination of ROS and RPI offers a powerful, compact solution for developing advanced robotic applications while maintaining a low-cost and accessible platform.
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=0.9\textwidth]{rpi_ros_graph.png}
+    \caption{Legged Robot RPi ROS Node/Topic Schematic}
+\end{figure}
+
+\section{DESIGN PROCESS}
+
+The design process of the prototype followed an iterative approach, involving multiple stages of development, testing, and evaluation.
+This methodology allowed the team to identify shortcomings in the initial design, implement modifications, and continuously improve the prototype's performance.
+Each iteration focused on addressing specific challenges, refining the sensor fusion techniques, and optimizing the SLAM implementation.
+In the following subsections, we present the steps of the design process performed in constructing the prototype, including the objectives, testing, results, and evaluation for each iteration.
+
+\subsection{ITERATION 1: INITIAL SENSOR INTEGRATION AND ROS SETUP}
+
+The first iteration aimed to establish a solid foundation for the project by integrating the different sensors and setting up the ROS framework for the robot.
+The main objectives of this iteration were:
+
+\begin{enumerate}
+    \item Selection and integration of appropriate sensors (LiDAR, camera, IMU, and wheel encoders) for the robot.
+    \item Installation and configuration of ROS on the robot's embedded computer system.
+    \item Setup of the communication infrastructure between the sensors, ROS nodes, and the robot's control system.
+\end{enumerate}
+
+In addition to the aforementioned advantages, the integration of the Raspberry Pi (RPI) with the Robot Operating System (ROS) allows for enhanced communication capabilities with micro-ROS systems.
+Micro-ROS is a lightweight version of ROS specifically designed for microcontrollers, enabling resource-constrained devices to become part of the ROS ecosystem.
+By incorporating micro-ROS into the system, it becomes possible to extend the functionality and modularity of the robotic platform even further.
+
+The RPI serves as the central communication hub between the ROS nodes and the micro-ROS system.
+It manages the flow of data between the different subsystems and ensures seamless interaction among the various components of the robot.
+To facilitate this communication, the RPI is connected to a network switch, which enables efficient data transmission and distribution throughout the system.
+
+The network switch plays a crucial role in the overall performance and reliability of the robotic system.
+It ensures that all components, including the RPI, micro-ROS nodes, and other connected devices, can exchange information effectively and maintain synchronized operation.
+By leveraging a network switch in conjunction with the RPI, the system can handle multiple data streams, balance network traffic, and reduce latency, thereby improving the overall performance of the robot.
+
+In summary, the combination of the Raspberry Pi, ROS, micro-ROS, and a network switch provides a powerful and versatile robotic platform.
+The RPI acts as the central processing unit and communication hub, allowing for efficient coordination between the ROS nodes and micro-ROS systems.
+The use of a network switch ensures seamless data exchange among the various components of the robot, resulting in a robust, scalable, and high-performance robotic system capable of tackling a wide range of applications.
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.9\textwidth]{Go1CommFram_E.png}
+    \caption{Infrastructure of the Unitree Go1 Robot}
+\end{figure}
+
+\subsubsection{TESTING AND RESULTS}
+
+The robot's hardware and software components were assembled, and the integration of sensors with the ROS framework was tested.
+The communication between the sensors and ROS nodes was verified using the provided visualization tools (such as RViz) to ensure proper data flow.
+Preliminary tests were conducted to evaluate the robot's ability to gather data from the environment using the sensors.
+
+\subsubsection{EVALUATION}
+
+The results from the first iteration revealed the successful integration of sensors and the establishment of the ROS environment.
+However, some issues were identified, such as occasional data loss due to communication bottlenecks and the need for better synchronization between the sensors.
+These issues were noted and addressed in the subsequent iterations.
+
+\subsection{ITERATION 2: SENSOR FUSION AND SLAM IMPLEMENTATION}
+
+In the second iteration, the focus was on incorporating sensor fusion techniques to combine the data from the different sensors and implementing SLAM algorithms to enable mapping and localization capabilities.
+The main objectives of this iteration were:
+
+\begin{enumerate}
+    \item Development and implementation of sensor fusion algorithms to combine data from the LiDAR, camera, IMU, and wheel encoders.
+    \item Selection and integration of a suitable SLAM algorithm for mapping and localization.
+    \item Evaluation of the robot's performance in mapping and navigation tasks using the fused sensor data and SLAM system.
+\end{enumerate}
+
+\subsubsection{TESTING AND RESULTS}
+
+The sensor fusion algorithms were tested to ensure the proper combination of data from the different sensors.
+The robot was then deployed in a controlled environment to evaluate its mapping and navigation capabilities using the SLAM system.
+The resulting maps were compared to the ground truth to assess the accuracy of the robot's localization and mapping.
+
+\subsubsection{EVALUATION}
+
+The results from the second iteration showed significant improvement in the robot's mapping and localization performance.
+The sensor fusion algorithms effectively combined the data from the different sensors, resulting in a more accurate and robust representation of the environment.
+However, some challenges remained, such as occasional drift in the robot's position estimation and the need for further optimization of the SLAM algorithm.
+These issues were addressed in the subsequent iterations.
+
+\subsubsection{ITERATION 3: IMPROVING SENSOR SYNCHRONIZATION AND ROBUSTNESS}
+
+In the third iteration, the team focused on enhancing sensor synchronization and overall system robustness to improve the robot's mapping and localization capabilities.
+The main objectives of this iteration were:
+
+\begin{enumerate}
+    \item Refine sensor synchronization to minimize data loss and inconsistencies between sensor readings.
+    \item Implement error handling and recovery mechanisms to increase the system's robustness.
+    \item Optimize the SLAM algorithm to reduce drift and improve localization accuracy.
+\end{enumerate}
+
+\subsubsection{TESTING AND RESULTS}
+
+The improved sensor synchronization was tested by analyzing the consistency of sensor data and verifying the data alignment using visualization tools.
+The robot was then deployed in a controlled environment, including areas with challenging conditions, such as uneven terrain, dynamic obstacles, and poor lighting.
+The robot's performance in mapping, localization, and navigation tasks was assessed, taking into consideration the implemented error handling and recovery mechanisms.
+
+\subsubsection{EVALUATION}
+
+The results of the third iteration showed a notable improvement in sensor synchronization and overall system robustness.
+The robot was able to handle challenging environmental conditions more effectively and recover from errors more efficiently.
+However, some areas for further optimization were identified, such as refining the SLAM algorithm for better performance in dynamic environments and enhancing the robot's ability to cope with sensor noise.
+
+\subsection{ITERATION 4: DYNAMIC ENVIRONMENT HANDLING AND NOISE REDUCTION}
+
+The fourth iteration aimed to address the remaining challenges by improving the robot's performance in dynamic environments and reducing the impact of sensor noise.
+The main objectives of this iteration were:
+
+\begin{enumerate}
+    \item Adapt the SLAM algorithm to better handle dynamic environments and moving obstacles.
+    \item Implement noise filtering techniques to reduce the impact of sensor noise on the robot's mapping and localization performance.
+    \item Evaluate the overall performance of the robot, including its robustness, adaptability, and accuracy in various environmental conditions.
+\end{enumerate}
+
+\subsubsection{TESTING AND RESULTS}
+
+The modified SLAM algorithm and noise filtering techniques were tested in multiple environments, including static and dynamic settings.
+The robot was deployed in a controlled environment with moving obstacles and varying lighting conditions to assess its ability to adapt and maintain accurate mapping and localization performance.
+
+\subsubsection{EVALUATION}
+
+The results from the fourth iteration demonstrated the robot's improved performance in dynamic environments and its enhanced ability to handle sensor noise.
+The robot was able to maintain accurate mapping and localization despite the presence of moving obstacles and varying environmental conditions.
+The noise filtering techniques effectively reduced the impact of sensor noise on the robot's performance.
+With these improvements, the prototype was considered ready for further real-world testing and potential future applications.
+
+\section{FINAL DESIGN}
+
+    \subsection{MEETING THE CONSTRAINTS AND ENGINEERING STANDARDS}
+
+    The final design of our project successfully met the design constraints outlined in the project description.
+We ensured compliance with the specified economy, environment, manufacturability, and sustainability constraints.
+
+        \subsubsection{Economy}
+
+        Throughout the project, we maintained strict control over costs to meet the project's economic constraints.
+By leveraging the Unitree Go1 Edu robot's internal Nvidia Jetson Nano 4GB devices for computing power, we were able to reduce expenses compared to using external computing units.
+Additionally, the use of readily available sensors and components helped keep the overall costs within budget.
+
+        \subsubsection{Environment}
+
+        The environmental impact of our project was considered at various stages of development.
+We focused on using battery-powered systems, minimizing power consumption, and adhering to legal regulations regarding radiation and noise pollution.
+By selecting energy-efficient components and optimizing algorithms, we aimed to reduce the ecological footprint of the robotic platform.
+
+        \subsubsection{Manufacturability}
+
+        The prototype design prioritized manufacturability by utilizing readily available products and components.
+This approach minimized the need for high-level manufacturing technologies and complex assembly processes.
+The modular design of the robot and its compatibility with existing ROS packages also facilitated the integration of different hardware and software components.
+
+        \subsubsection{Sustainability}
+
+        Our project aimed to establish a sustainable foundation for future developments.
+By emphasizing the mathematical foundations, software abstractions, and comprehensive engineering solutions, we laid the groundwork for future advancements and upgrades.
+The modular design of the robot, adherence to engineering standards, and consideration of safety measures ensured the long-term sustainability and reliability of the robotic platform.
+
+        The project also met the engineering standards and best practices to ensure the reliability, maintainability, and safety of the robotic platform.
+We followed software engineering principles, such as modular design, code documentation, version control, and code reviews.
+The adoption of the ROS framework provided a standardized and modular approach to software development, enhancing interoperability and code reusability.
+Safety measures were implemented to prevent potential hazards, and documentation and reporting were maintained throughout the project to ensure traceability and knowledge dissemination.
+
+    \subsection{COST ANALYSIS}
+
+    The cost analysis of our project includes both initial costs and potential future costs for production and maintenance.
+The initial costs include the purchase of components and materials required for building the prototype, while the future costs involve scaling up production and maintaining the robotic platform.
+
+        \subsubsection{Initial Costs}
+
+        The initial costs for our project primarily consisted of the Unitree Go1 Edu robot and additional sensors.
+The Unitree Go1 Edu robot, including its embedded computing units and sensors, had a cost of approximately \$ 18,000.
+The cost breakdown of the additional sensors used in the project is as follows:
+
+        \begin{itemize}
+            \item LiDAR: Velodyne 80-VLP-16-A 3D LiDAR - \$ 2500.
+            \item Camera: StereoLabs ZED Stereo Camera - \$ 1000.
+        \end{itemize}
+
+        The total initial cost of the prototype, including the robot and additional sensors, amounted to approximately \$ 22,500.
+
+        \subsubsection{Future Costs}
+
+        Future costs for production and maintenance may vary depending on the scale of deployment and specific application requirements. Some potential future costs to consider include:
+
+        \begin{itemize}
+            \item Production Costs: Scaling up production would involve the cost of manufacturing additional robotic platforms, including the necessary components and assembly processes.
+These costs would depend on the production volume and any customization required for specific applications.
+            \item Maintenance Costs: Maintenance costs would include periodic hardware inspections, software updates, and component replacements.
+The costs would depend on the frequency and complexity of maintenance tasks, as well as any warranty or support agreements.
+            \item Upgrades and Expansion: As technology advances and new sensors or algorithms become available, there may be additional costs associated with upgrading or expanding the capabilities of the robotic platform.
+These costs would depend on the specific upgrades and their integration complexity.
+        \end{itemize}
+
+        It is important to note that the costs mentioned above are estimates based on current market prices and may vary depending on factors such as location, supplier, and future market conditions.
+
+        \subsubsection{Cost-Benefit Analysis}
+
+        While the cost of the project may be significant, it is essential to consider the potential benefits and applications of the developed robotic platform.
+The integration of sensor fusion, SLAM techniques, and cloud computing provides opportunities for various industries, including autonomous vehicles, UAVs, augmented reality devices, and 3D modeling.
+The accuracy and efficiency of the depth estimation and mapping capabilities make the platform valuable for navigation in unknown environments.
+The scalability and modularity of the design also allow for future upgrades and customization, ensuring long-term usability and adaptability.
+
+        The cost-benefit analysis should take into account the potential economic, environmental, and societal impact of deploying such a robotic platform.
+While the costs may be higher initially, the benefits derived from improved navigation, mapping, and perception capabilities can outweigh the expenses in various domains.
+Furthermore, the sustainability and modularity of the design contribute to long-term cost savings by enabling upgrades and maintenance rather than complete system replacements.
+
+    \subsection{SUMMARY}
+
+    The final design of our project successfully met the specified design constraints, including economic, environmental, manufacturability, and sustainability considerations.
+The adherence to engineering standards ensured the reliability, maintainability, and safety of the robotic platform.
+The cost analysis provided an overview of the initial costs and potential future costs for production, maintenance, and upgrades.
+By considering the potential benefits and applications of the developed robotic platform, the cost-benefit analysis demonstrates the value and long-term viability of the project.
+Overall, the final design represents a comprehensive and practical solution for sensor fusion, SLAM implementation, and cloud computing integration in autonomous robotic systems.
+
 \section{TEAM WORK}

### 2024-02-27 paper/main.tex
Create main.tex

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

\title{Bismillahirrahmanirrahim iSLAM\\
{\footnotesize \textsuperscript{*}Note: Sub-titles are not captured in Xplore and
should not be used}
\thanks{Identify applicable funding agency here. If none, delete this.}
}

\author{
\IEEEauthorblockN{1\textsuperscript{st} Assc. Prof. İsmail Uyanık}
\IEEEauthorblockA{\textit{Department of Electrical and Electronics Engineering} \\
\textit{Hacettepe University}\\
Ankara, Turkiye \\
uyanik@ee.hacettepe.edu.tr}
\and
\IEEEauthorblockN{2\textsuperscript{st} Ahmet Yusuf Şirin}
\IEEEauthorblockA{\textit{Department of Electrical and Electronics Engineering} \\
\textit{Hacettepe University}\\
Ankara, Turkiye \\
ayusufsirin@gmail.com}
\and
\IEEEauthorblockN{3\textsuperscript{rd} Ertuğrul Tiyek}
\IEEEauthorblockA{\textit{Department of Electrical and Electronics Engineering} \\
\textit{Hacettepe University}\\
Ankara, Turkiye \\
ertugrultyk@gmail.com}
}

\maketitle

\begin{abstract}
Depth perception in autonomous systems is critically enhanced by sensor fusion techniques, which traditionally rely on the combination of stereo vision and LiDAR data. However, the challenge of integrating sparse LiDAR data with dense stereo imagery remains. This paper introduces a novel fusion approach that applies the Papoulis-Gerchberg algorithm to iteratively refine stereo depth maps with LiDAR data. Our method uses LiDAR measurements as anchor points to guide the enhancement of depth information, resulting in superior resolution and accuracy. We evaluate our approach using the KITTI benchmark suite, demonstrating significant improvements over existing methods such as FastFusion, particularly in edge definition and object boundary sharpness. The results suggest our method not only achieves higher fidelity in depth estimation but also offers a new perspective on sensor data integration for real-time applications in autonomous navigation.
\end{abstract}

\begin{IEEEkeywords}
Sensor Fusion, Stereo Vision, LiDAR Depth Sensing, Papoulis-Gerchberg Algorithm, Autonomous Vehicles, Iterative Enhancement, Depth Map, Super-Resolution, Real-Time Depth Estimation
\end{IEEEkeywords}

\section{Introduction}

The quest for accurate depth perception in autonomous systems has given rise to various sensor fusion methodologies, predominantly those employing neural networks. While these data-driven techniques have demonstrated remarkable performance, they often require extensive training on large datasets and substantial computational resources, limiting their practicality in dynamic real-world scenarios.

Contrary to the prevalent neural network-based approaches, this paper introduces a novel sensor fusion framework that does not rely on learning from data. By adapting the Papoulis-Gerchberg algorithm, traditionally used for signal reconstruction and super-resolution, our method iteratively refines stereo depth maps using LiDAR data without the need for training. This non-learning-based approach circumvents the complexities of neural models, resulting in a lightweight and versatile solution suitable for real-time applications.

Our methodology capitalizes on the complementary strengths of stereo vision and LiDAR, embedding accurate depth measurements into dense stereo maps to produce superior environmental representations. By eschewing neural networks, we avoid their inherent demand for computational power and data, presenting an approach that is not only more practical but also readily deployable in diverse operational contexts.

The following sections will outline the current landscape of depth sensing and fusion techniques, detail our fusion algorithm, and present a comparative analysis with existing neural network-based methods, showcasing the practical and performance benefits of our approach.

\section{Related Works}

Stereo matching and depth estimation are pivotal in applications such as autonomous driving and robotic navigation. Traditional methods, as outlined by Scharstein and Szeliski (2002), have evolved with the adoption of Deep Neural Network (DNN) techniques, significantly enhancing accuracy (Zbontar & LeCun, 2016). However, the computational demands of these DNN solutions often outpace the capabilities of onboard systems in real-time scenarios (Shamsafar et al., 2022; Rahim et al., 2021).

In the realm of stereo-LiDAR fusion, approaches like those proposed by Parks et al. (2018) and Park et al. (2019) augment stereo disparity with LiDAR data but fail to fully exploit the precision of LiDAR. Choe et al. (2021) and Yang et al. (2019) introduce architectures that incorporate LiDAR data into early stages of depth estimation but often at a sacrifice of either depth integration or accuracy. Other methods, such as those by Mai et al. (2021) and Li and Cao (2020), achieve higher accuracy but at the cost of increased computational load, which is not conducive to real-time application.

Our work diverges from these DNN-heavy strategies by employing the Papoulis-Gerchberg algorithm, a technique historically utilized in signal processing for super-resolution, adapted here for sensor fusion. This approach circumvents the prohibitive computational demands of DNNs, offering a lightweight yet robust alternative for real-time, accurate depth estimation from stereo-LiDAR data fusion.

\section{Technical Approach}
\subsection{Pre-Processing}
\subsubsection{Data Augmentation}
\subsubsection{Stereo Infill}
\subsubsection{Lidar Null Mask}
\subsection{PG: Papoulis GerchBerg}
\subsubsection{PG Background}
\subsubsection{our Approach}
\subsubsection{Use of Band Reject Filter in PG}
\subsubsection{Evaluated PG for Our Work}
\subsection{Post-Proc}
\subsubsection{Removal of Trivial Aumented Data}

\section{Implementation}
\subsection{Adopting the Algorithm to Linear Algebra}

\section{Optimization}
\subsection{GPU Acceleration}
\subsection{Cloud Computing}

\section{Experiment}

\subsection{Experimenting our Algorithm With Kitti Benchmark Suite}
%% Mention the dataset preparation picking data from different datasets

\subsubsection{Effect of Shape and Format of Lidar Data}
\subsubsection{Experiment Scenarios}

\subsection{Experiments With Our Sensor Suit}
%% Mention the moving system and challenges, calibration and synchronization
%% Realtime

\subsubsection{Sensor Suit}
\subsubsubsection{Upsampling of lidar data} %% Further improvements dynamic lidar upsampling with kalman

\subsubsection{Experiment Scenarios}

\section{Results}

\subsection{Abbreviations and Acronyms}\label{AA}

\subsection{Units}
\begin{itemize}
\item Use either SI (MKS) or CGS as primary units. (SI units are encouraged.) English units may be used as secondary units (in parentheses). An exception would be the use of English units as identifiers in trade, such as ``3.5-inch disk drive''.
\item Avoid combining SI and CGS units, such as current in amperes and magnetic field in oersteds. This often leads to confusion because equations do not balance dimensionally. If you must use mixed units, clearly state the units for each quantity that you use in an equation.
\item Do not mix complete spellings and abbreviations of units: ``Wb/m\textsuperscript{2}'' or ``webers per square meter'', not ``webers/m\textsuperscript{2}''. Spell out units when they appear in text: ``. . . a few henries'', not ``. . . a few H''.
\item Use a zero before decimal points: ``0.25'', not ``.25''. Use ``cm\textsuperscript{3}'', not ``cc''.)
\end{itemize}

\subsection{Equations}
Number equations consecutively. To make your 
equations more compact, you may use the solidus (~/~), the exp function, or 
appropriate exponents. Italicize Roman symbols for quantities and variables, 
but not Greek symbols. Use a long dash rather than a hyphen for a minus 
sign. Punctuate equations with commas or periods when they are part of a 
sentence, as in:
\begin{equation}
a+b=\gamma\label{eq}
\end{equation}

Be sure that the 
symbols in your equation have been defined before or immediately following 
the equation. Use ``\eqref{eq}'', not ``Eq.~\eqref{eq}'' or ``equation \eqref{eq}'', except at 
the beginning of a sentence: ``Equation \eqref{eq} is . . .''

\subsection{\LaTeX-Specific Advice}

Please use ``soft'' (e.g., \verb|\eqref{Eq}|) cross references instead
of ``hard'' references (e.g., \verb|(1)|). That will make it possible
to combine sections, add equations, or change the order of figures or
citations without having to go through the file line by line.

Please don't use the \verb|{eqnarray}| equation environment. Use
\verb|{align}| or \verb|{IEEEeqnarray}| instead. The \verb|{eqnarray}|
environment leaves unsightly spaces around relation symbols.

Please note that the \verb|{subequations}| environment in {\LaTeX}
will increment the main equation counter even when there are no
equation numbers displayed. If you forget that, you might write an
article in which the equation numbers skip from (17) to (20), causing
the copy editors to wonder if you've discovered a new method of
counting.

{\BibTeX} does not work by magic. It doesn't get the bibliographic
data from thin air but from .bib files. If you use {\BibTeX} to produce a
bibliography you must send the .bib files.

{\LaTeX} can't read your mind. If you assign the same label to a
subsubsection and a table, you might find that Table I has been cross
referenced as Table IV-B3.

{\LaTeX} does not have precognitive abilities. If you put a
\verb|\label| command before the command that updates the counter it's
supposed to be using, the label will pick up the last counter to be
cross referenced instead. In particular, a \verb|\label| command
should not go before the caption of a figure or a table.

Do not use \verb|\nonumber| inside the \verb|{array}| environment. It
will not stop equation numbers inside \verb|{array}| (there won't be
any anyway) and it might stop a wanted equation number in the
surrounding equation.

\subsection{Some Common Mistakes}\label{SCM}
\begin{itemize}
\item The word ``data'' is plural, not singular.
\item The subscript for the permeability of vacuum $\mu_{0}$, and other common scientific constants, is zero with subscript formatting, not a lowercase letter ``o''.
\item In American English, commas, semicolons, periods, question and exclamation marks are located within quotation marks only when a complete thought or name is cited, such as a title or full quotation. When quotation marks are used, instead of a bold or italic typeface, to highlight a word or phrase, punctuation should appear outside of the quotation marks. A parenthetical phrase or statement at the end of a sentence is punctuated outside of the closing parenthesis (like this). (A parenthetical sentence is punctuated within the parentheses.)
\item A graph within a graph is an ``inset'', not an ``insert''. The word alternatively is preferred to the word ``alternately'' (unless you really mean something that alternates).
\item Do not use the word ``essentially'' to mean ``approximately'' or ``effectively''.
\item In your paper title, if the words ``that uses'' can accurately replace the word ``using'', capitalize the ``u''; if not, keep using lower-cased.
\item Be aware of the different meanings of the homophones ``affect'' and ``effect'', ``complement'' and ``compliment'', ``discreet'' and ``discrete'', ``principal'' and ``principle''.
\item Do not confuse ``imply'' and ``infer''.
\item The prefix ``non'' is not a word; it should be joined to the word it modifies, usually without a hyphen.
\item There is no period after the ``et'' in the Latin abbreviation ``et al.''.
\item The abbreviation ``i.e.'' means ``that is'', and the abbreviation ``e.g.'' means ``for example''.
\end{itemize}
An excellent style manual for science writers is \cite{b7}.

\subsection{Authors and Affiliations}
\textbf{The class file is designed for, but not limited to, six authors.} A 
minimum of one author is required for all conference articles. Author names 
should be listed starting from left to right and then moving down to the 
next line. This is the author sequence that will be used in future citations 
and by indexing services. Names should not be listed in columns nor group by 
affiliation. Please keep your affiliations as succinct as possible (for 
example, do not differentiate among departments of the same organization).

\subsection{Identify the Headings}
Headings, or heads, are organizational devices that guide the reader through 
your paper. There are two types: component heads and text heads.

Component heads identify the different components of your paper and are not 
topically subordinate to each other. Examples include Acknowledgments and 
References and, for these, the correct style to use is ``Heading 5''. Use 
``figure caption'' for your Figure captions, and ``table head'' for your 
table title. Run-in heads, such as ``Abstract'', will require you to apply a 
style (in this case, italic) in addition to the style provided by the drop 
down menu to differentiate the head from the text.

Text heads organize the topics on a relational, hierarchical basis. For 
example, the paper title is the primary text head because all subsequent 
material relates and elaborates on this one topic. If there are two or more 
sub-topics, the next level head (uppercase Roman numerals) should be used 
and, conversely, if there are not at least two sub-topics, then no subheads 
should be introduced.

\subsection{Figures and Tables}
\paragraph{Positioning Figures and Tables} Place figures and tables at the top and 
bottom of columns. Avoid placing them in the middle of columns. Large 
figures and tables may span across both columns. Figure captions should be 
below the figures; table heads should appear above the tables. Insert 
figures and tables after they are cited in the text. Use the abbreviation 
``Fig.~\ref{fig}'', even at the beginning of a sentence.

\begin{table}[htbp]
\caption{Table Type Styles}
\begin{center}
\begin{tabular}{|c|c|c|c|}
\hline
\textbf{Table}&\multicolumn{3}{|c|}{\textbf{Table Column Head}} \\
\cline{2-4} 
\textbf{Head} & \textbf{\textit{Table column subhead}}& \textbf{\textit{Subhead}}& \textbf{\textit{Subhead}} \\
\hline
copy& More table copy$^{\mathrm{a}}$& &  \\
\hline
\multicolumn{4}{l}{$^{\mathrm{a}}$Sample of a Table footnote.}
\end{tabular}
\label{tab1}
\end{center}
\end{table}

\begin{figure}[htbp]
\centerline{\includegraphics{fig1.png}}
\caption{Example of a figure caption.}
\label{fig}
\end{figure}

Figure Labels: Use 8 point Times New Roman for Figure labels. Use words 
rather than symbols or abbreviations when writing Figure axis labels to 
avoid confusing the reader. As an example, write the quantity 
``Magnetization'', or ``Magnetization, M'', not just ``M''. If including 
units in the label, present them within parentheses. Do not label axes only 
with units. In the example, write ``Magnetization (A/m)'' or ``Magnetization 
\{A[m(1)]\}'', not just ``A/m''. Do not label axes with a ratio of 
quantities and units. For example, write ``Temperature (K)'', not 
``Temperature/K''.

\section*{Acknowledgment}

The preferred spelling of the word ``acknowledgment'' in America is without 
an ``e'' after the ``g''. Avoid the stilted expression ``one of us (R. B. 
G.) thanks $\ldots$''. Instead, try ``R. B. G. thanks$\ldots$''. Put sponsor 
acknowledgments in the unnumbered footnote on the first page.

\section*{References}

Please number citations consecutively within brackets \cite{b1}. The 
sentence punctuation follows the bracket \cite{b2}. Refer simply to the reference 
number, as in \cite{b3}---do not use ``Ref. \cite{b3}'' or ``reference \cite{b3}'' except at 
the beginning of a sentence: ``Reference \cite{b3} was the first $\ldots$''

Number footnotes separately in superscripts. Place the actual footnote at 
the bottom of the column in which it was cited. Do not put footnotes in the 
abstract or reference list. Use letters for table footnotes.

Unless there are six authors or more give all authors' names; do not use 
``et al.''. Papers that have not been published, even if they have been 
submitted for publication, should be cited as ``unpublished'' \cite{b4}. Papers 
that have been accepted for publication should be cited as ``in press'' \cite{b5}. 
Capitalize only the first word in a paper title, except for proper nouns and 
element symbols.

For papers published in translation journals, please give the English 
citation first, followed by the original foreign-language citation \cite{b6}.

\begin{thebibliography}{00}
\bibitem{b1} G. Eason, B. Noble, and I. N. Sneddon, ``On certain integrals of Lipschitz-Hankel type involving products of Bessel functions,'' Phil. Trans. Roy. Soc. London, vol. A247, pp. 529--551, April 1955.
\bibitem{b2} J. Clerk Maxwell, A Treatise on Electricity and Magnetism, 3rd ed., vol. 2. Oxford: Clarendon, 1892, pp.68--73.
\bibitem{b3} I. S. Jacobs and C. P. Bean, ``Fine particles, thin films and exchange anisotropy,'' in Magnetism, vol. III, G. T. Rado and H. Suhl, Eds. New York: Academic, 1963, pp. 271--350.
\bibitem{b4} K. Elissa, ``Title of paper if known,'' unpublished.
\bibitem{b5} R. Nicole, ``Title of paper with only first word capitalized,'' J. Name Stand. Abbrev., in press.
\bibitem{b6} Y. Yorozu, M. Hirano, K. Oka, and Y. Tagawa, ``Electron spectroscopy studies on magneto-optical media and plastic substrate interface,'' IEEE Transl. J. Magn. Japan, vol. 2, pp. 740--741, August 1987 [Digests 9th Annual Conf. Magnetics Japan, p. 301, 1982].
\bibitem{b7} M. Young, The Technical Writer's Handbook. Mill Valley, CA: University Science, 1989.
\end{thebibliography}
\vspace{12pt}
\color{red}
IEEE conference templates contain guidance text for composing and formatting conference papers. Please ensure that all template text is removed from your conference paper prior to submission to the conference. Failure to remove the template text from your paper may result in your paper not being published.

\end{document}

### 2024-02-28 paper/main.tex
Update main.tex

@@
 Stereo matching and depth estimation are pivotal in applications such as autonomous driving and robotic navigation.
-Traditional methods, as outlined by Scharstein and Szeliski (2002), have evolved with the adoption of Deep Neural Network (DNN) techniques, significantly enhancing accuracy (Zbontar & LeCun, 2016).
+Traditional methods, as outlined by Scharstein and Szeliski (2002), have evolved with the adoption of Deep Neural Network (DNN) techniques, significantly enhancing accuracy (Zbontar \& LeCun, 2016).
 However, the computational demands of these DNN solutions often outpace the capabilities of onboard systems in real-time scenarios (Shamsafar et al., 2022; Rahim et al., 2021).
@@
 \section{Technical Approach}
+%% In this section explain the each step of the processing detaily.
+
+Our approach is based on the PG method that is commonly used for kayıplı sinyal kestirimi in communications systems.
+The method is applicable for 1D signals, originally.
+The 2D implementation of it is required for us.
+There are some works on implementation to 2D of PG.
+
+An application of this algorithm is used for image inpainting by \cite{pg-image-inpaint-2009}.
+They introduce the PG for this specific scenerio but it has some precautions to use for us.
+The images they used are extremely dense when comparing with the Lidar.
+Also, they have single sensory image but we need to optain a fusion.
+
+Another work done by \cite{ozbay2015high} that introduce an extended version of PG for 3D point cloud data that is driven by a Lidar sensor.
+They use this as an upsampling to cope with delayed points when using Lidar in high resolution mode due to the low scan rates of cylindirical scanning ones.
+But appearantly, it also use single sensory system
+
+We introduce our novel approach my putting another extension of PG to the litarature that is the first one uses multi-sensory system. The nature of PG consists some elements.
+
+%% PG açıklaması başta olsa daha iyi olabilir.
+
 \subsection{Pre-Processing}
 \subsubsection{Data Augmentation}
-\subsubsection{Stereo Infill}
-\subsubsection{Lidar Null Mask}
+Due to the fundamental difference in the representation of depth information between liDAR measurements (point cloud) and stereo depth estimation (dense disparity maps), both data modalities necessitate preprocessing steps involving transformation, conversion, and upsampling to facilitate their integration and subsequent processing.
+
+REVISE - To make these two sensor data processable in a domain, we determine a common modality for these images.
+This modality is single channel depth image format.
+The RGB integration wont necessite for using our algorithm.
+
+\paragraph{Obtaining Depth Frame From Lidar Data}
+
+After making the calibrations for stereo camera and the Lidar, we know all the transformation matrices to generate the depth image from point cloud data.
+To project the 3D points into the 2D depth frame space we use the intrinsic parameters and the known transformation matrices.
+
+\paragraph{Lidar Null Mask}
+
+\paragraph{Stereo Infill}
+
 \subsection{PG: Papoulis GerchBerg}
-\subsubsection{PG Background}
-\subsubsection{our Approach}
+%% \subsubsection{PG Background}
+%% Cite other papers about papoulis gerchberg, super resolution
+
+\subsubsection{Our Approach}
 \subsubsection{Use of Band Reject Filter in PG}
@@
 \subsubsection{Sensor Suit}
-\subsubsubsection{Upsampling of lidar data} %% Further improvements dynamic lidar upsampling with kalman
+\paragraph{Upsampling of Lidar Data}
+%% Further improvements dynamic lidar upsampling with kalman
 
@@
 \end{thebibliography}
+
+\bibliographystyle{plain} % We choose the "plain" reference style
+\bibliography{refs} % Entries are in the refs.bib file
+
 \vspace{12pt}

### 2024-02-28 paper/main.tex
Update main.tex

@@
 
-\title{Bismillahirrahmanirrahim iSLAM\\
+\title{Enhancing Depth Estimation with Iterative Stereo-LiDAR Fusion:An Application of the Papoulis-Gerchberg Algorithm\\
 {\footnotesize \textsuperscript{*}Note: Sub-titles are not captured in Xplore and
@@
 
-Our approach is based on the PG method that is commonly used for kayıplı sinyal kestirimi in communications systems.
+The Papoulis-Gerchberg (PG) algorithm is an iterative technique employed in signal processing to perform band-limited signal extrapolation.
+This method enables the estimation of missing portions of a signal based on a limited segment of the original data and the knowledge of its bandwidth.
+
+The PG algorithm operates through an iterative process involving two key steps:
+
+\begin{enumerate}
+    \item Constraint Enforcement in the Spatial Domain:
+    The initial step involves enforcing the known samples of the signal.
+This is achieved by transforming the estimated spectrum (obtained through techniques like the Fourier transform) of the signal and modifying it to ensure it aligns with the known Fourier Transform values corresponding to the available samples.
+
+    \item Constraint Enforcement in the Frequency Domain:
+    The second step focuses on enforcing the known bandwidth constraint.
+This is implemented by modifying the signal's estimated version in the spatial domain (e.g., time domain for time-based signals) to guarantee it adheres to the specified bandwidth limitations.
+\end{enumerate}
+These two steps are repeated iteratively.
+As each iteration progresses, the estimated signal progressively converges towards the original signal, ultimately resulting in an extrapolated version that replenishes the missing data points.
+
+Further works opened the papoulis-gerchberg algorithm new using areas like signal restoration and image processing.
+%% Our approach is based on the PG method that is commonly used for estimation and reconstruction of lossy signals in communications systems.
 The method is applicable for 1D signals, originally.
@@
 
-An application of this algorithm is used for image inpainting by \cite{pg-image-inpaint-2009}.
+%% Paper about signal restoration
+%% Paper about super resolution
+Another application of this algorithm is used for image inpainting by \cite{pg-image-inpaint-2009}.
 They introduce the PG for this specific scenerio but it has some precautions to use for us.
@@
 After making the calibrations for stereo camera and the Lidar, we know all the transformation matrices to generate the depth image from point cloud data.
-To project the 3D points into the 2D depth frame space we use the intrinsic parameters and the known transformation matrices.
+To project the 3D points into the 2D depth frame space we use the pinhole camera model.
+%% TODO: Pinhole camera model daha mu uygun? use the intrinsic parameters and the known transformation matrices.
 
@@
 
-\section{Implementation}
+\section{Implementation} \label{implementation}
 \subsection{Adopting the Algorithm to Linear Algebra}
@@
 \section{Optimization}
+
+As defined in the previous section \ref{implementation}, our method is developed baed on the linear algebraic methods.
+Using the nature of the linear algebra, we introduce a parallel processing approach for our extended PG method.
+This introduce us a workspace that we can competite with our fellows (Buraya DNN gibi yapıları kullanarak GPU üzerinde koşan sensor fusion paperlarına alıntı yapılacak)
+
 \subsection{GPU Acceleration}

### 2024-02-28 paper/main.tex
Update main.tex

@@
 
-\title{Enhancing Depth Estimation with Iterative Stereo-LiDAR Fusion:An Application of the Papoulis-Gerchberg Algorithm\\
+\title{Enhancing Depth Estimation with Iterative Stereo-LiDAR Fusion: An Application of the Papoulis-Gerchberg Algorithm\\
 {\footnotesize \textsuperscript{*}Note: Sub-titles are not captured in Xplore and
@@
 
+Stereo depth augmentation requirement is originated from the deficit nature of the disparity maps that are used in depth estimation from stereo cameras.
+The depth images, due to the disparity maps, have some null values at blind spots that one of the cameras have no view at all.
+These null values at the depth images should be pre-processed before the PG algorithm to avoid the distortion of the well estimated depth entities (pixels).
+
+We proposed a method to overcome the null value problem.
+This method fills the gaps (null values at the depth images) by using the lidar measurements.
+Firstly, a segmentation of the gaps are made and then the lidar measurements are taken into account to fill the gap segments that are overlapped with lidar measurements.
+Again, due to the sparse nature of the lidar measurements, we made a interpolation for filling the huge gaps by using lidar measurements that have small area.
+
 \paragraph{Obtaining Depth Frame From Lidar Data}
@@
 \paragraph{Lidar Null Mask}
+
+The lidar depth frames will have null values do to the nature of the projection of sparse point cloud data into a plane.
+This null values should be considered before using the lidar depth map in the algorithm.
+
+Since, our method use the stereo depth maps as a starting point and lidar depth map as rectifier of depth information of these stereo depth images, we need a mask of null values to eliminate the spurious pixels (depth entities) at the post-processing part of our algorithm.
 

### 2024-02-28 paper/main.tex
Update main.tex

@@
 They introduce the PG for this specific scenerio but it has some precautions to use for us.
-The images they used are extremely dense when comparing with the Lidar.
+The images they used are extremely dense when comparing with the LiDAR.
 Also, they have single sensory image but we need to optain a fusion.
 
-Another work done by \cite{ozbay2015high} that introduce an extended version of PG for 3D point cloud data that is driven by a Lidar sensor.
-They use this as an upsampling to cope with delayed points when using Lidar in high resolution mode due to the low scan rates of cylindirical scanning ones.
+Another work done by \cite{ozbay2015high} that introduce an extended version of PG for 3D point cloud data that is driven by a LiDAR sensor.
+They use this as an upsampling to cope with delayed points when using LiDAR in high resolution mode due to the low scan rates of cylindirical scanning ones.
 But appearantly, it also use single sensory system
@@
 
-%% PG açıklaması başta olsa daha iyi olabilir.
+\subsection{Pre-Processing}
 
-\subsection{Pre-Processing}
-\subsubsection{Data Augmentation}
-Due to the fundamental difference in the representation of depth information between liDAR measurements (point cloud) and stereo depth estimation (dense disparity maps), both data modalities necessitate preprocessing steps involving transformation, conversion, and upsampling to facilitate their integration and subsequent processing.
+Due to the fundamental difference in the representation of depth information between LiDAR measurements (point cloud) and stereo depth estimation (dense disparity maps), both data modalities necessitate preprocessing steps involving transformation, conversion, and upsampling to facilitate their integration and subsequent processing.
 
@@
 The RGB integration wont necessite for using our algorithm.
+
+\subsubsection{Data Cropping}
+
+After projecting the LiDAR image onto the stereo depth image plane, there are some remaining problems.
+One of them is the boundaries of the depth regions.
+Since the vertical scan area of the lidar sensor is pretty small (like 30 degrees) we have to crop the stereo depth image also.
+On the other hand, the stereo depth image have a bound smaller than the lidar scan at the on the horzontal axis.
+This is appearent, because the lidar horizontal scan area is 360 degrees.
+Since our algorithm is based on depth images and there are no memories for previously detected points for the frame sensor fusioning, we just ignore this non-overlapping depth data regions.
+
+\subsubsection{Data Augmentation}
 
@@
 We proposed a method to overcome the null value problem.
-This method fills the gaps (null values at the depth images) by using the lidar measurements.
-Firstly, a segmentation of the gaps are made and then the lidar measurements are taken into account to fill the gap segments that are overlapped with lidar measurements.
-Again, due to the sparse nature of the lidar measurements, we made a interpolation for filling the huge gaps by using lidar measurements that have small area.
+This method fills the gaps (null values at the depth images) by using the LiDAR measurements.
+Firstly, a segmentation of the gaps are made and then the LiDAR measurements are taken into account to fill the gap segments that are overlapped with LiDAR measurements.
+Again, due to the sparse nature of the LiDAR measurements, we made a interpolation for filling the huge gaps by using LiDAR measurements that have small area.
 
-\paragraph{Obtaining Depth Frame From Lidar Data}
+\paragraph{Obtaining Depth Frame From LiDAR Data}
 
-After making the calibrations for stereo camera and the Lidar, we know all the transformation matrices to generate the depth image from point cloud data.
+After making the calibrations for stereo camera and the LiDAR, we know all the transformation matrices to generate the depth image from point cloud data.
 To project the 3D points into the 2D depth frame space we use the pinhole camera model.
-%% TODO: Pinhole camera model daha mu uygun? use the intrinsic parameters and the known transformation matrices.
 
-\paragraph{Lidar Null Mask}
+\paragraph{LiDAR Null Mask}
 
-The lidar depth frames will have null values do to the nature of the projection of sparse point cloud data into a plane.
-This null values should be considered before using the lidar depth map in the algorithm.
+The LiDAR depth frames will have null values do to the nature of the projection of sparse point cloud data into a plane.
+This null values should be considered before using the LiDAR depth map in the algorithm.
 
-Since, our method use the stereo depth maps as a starting point and lidar depth map as rectifier of depth information of these stereo depth images, we need a mask of null values to eliminate the spurious pixels (depth entities) at the post-processing part of our algorithm.
+Since, our method use the stereo depth maps as a starting point and LiDAR depth map as rectifier of depth information of these stereo depth images, we need a mask of null values to eliminate the spurious pixels (depth entities) at the post-processing part of our algorithm.
 
@@
 
-\subsection{PG: Papoulis GerchBerg}
-%% \subsubsection{PG Background}
-%% Cite other papers about papoulis gerchberg, super resolution
+\subsection{PG: Papoulis Gerchberg}
+We first intended to use Papoulish-Gerchberg algorithm as a pre-processing method to up sample the only LiDAR data. But the sparsity of the LiDAR data made it challenging to work with.
 
+We tried to apply zero-padding to empty regions of the depth frame obtained by the LiDAR samples.
+Since the Papoulish-Gerchberg method is an iterative method, zero background filling caused high amount of iterations to get reasonable results.
+And due to the sparse data especially in the vertical axis, the regions far from original LiDAR samples never get good results like the closer ones.
+%% add an image to represent the papoulis with zero-padding
+
+Then to decrease number of iterations, we tried to initialize the background with mean of actual LiDAR samples.
+Since the mean value is much closer to the actual depth of the neighboring samples, it significantly reduced the iterations.
+But the problem with the sparsity of LiDAR data was still present.
+%% add another figure for mean filling
+
+We need more significant data to overcome this problem.
+The initial solution was the using previous work of our colleague @ismail-uyanik, {Enhancing 3D range image measurement density via dynamic Papoulis-Gerchberg algorithm} along with the mean filling method.
+%% cite the paper from kuzucu-elvan
+%% add another image explaining the plan
 \subsubsection{Our Approach}
-\subsubsection{Use of Band Reject Filter in PG}
-\subsubsection{Evaluated PG for Our Work}
+As the purpose of this study is to investigate the potential of integrating data obtained from stereo cameras and LiDAR measurements, we also have the depth information from stereo image pair.
+Our motivation of this paper's work turned to merge both processes, up sampling of LiDAR data and fusing the depth information obtained from both sources.
+This method extremely decreased the need of processing power, and allowed real-time application of the sensor fusion even in mobile computing devices.
+Basically the method is nothing but proposing the background of lidar measurements as depth frame from stereo pair instead of using only a mean value.
+Since the background filling is not non-significant data, we have mostly overcame the problem with sparsity of LiDAR measurements.
+We also made some modifications and improvements to the standard 2D Papoulish-Gerchberg algorithm to match our needs better.
+\paragraph{Use of Band Reject Filter in PG}
+\paragraph{Evaluated PG for Our Work}
+
 \subsection{Post-Proc}
-\subsubsection{Removal of Trivial Aumented Data}
+
+The post-processing consinst of two major parts
+
+\subsubsection{Removal of Trivial Augmented Data}
 
 \section{Implementation} \label{implementation}
+
+The implementation of a proposed method is essential for experiments.
+Because of the robotic fields requirements, there are some constraints about the frequency of the algorithm, data rate of te sensors, and some other important facts.
+To accomplish all of this constraints, we use the power of the mathematics and the linear algebra.
+
+\subsection{Building the Algorithm}
+
 \subsection{Adopting the Algorithm to Linear Algebra}
@@
 
-\subsubsection{Effect of Shape and Format of Lidar Data}
+\subsubsection{Effect of Shape and Format of LiDAR Data}
 \subsubsection{Experiment Scenarios}
@@
 \subsubsection{Sensor Suit}
-\paragraph{Upsampling of Lidar Data}
-%% Further improvements dynamic lidar upsampling with kalman
+\paragraph{Upsampling of LiDAR Data}
+%% Further improvements dynamic LiDAR upsampling with kalman
 
@@
 \section{Results}
-
-\subsection{Abbreviations and Acronyms}\label{AA}
-
-\subsection{Units}
-\begin{itemize}
-\item Use either SI (MKS) or CGS as primary units.
-(SI units are encouraged.) English units may be used as secondary units (in parentheses).
-An exception would be the use of English units as identifiers in trade, such as ``3.5-inch disk drive''.
-\item Avoid combining SI and CGS units, such as current in amperes and magnetic field in oersteds.
-This often leads to confusion because equations do not balance dimensionally.
-If you must use mixed units, clearly state the units for each quantity that you use in an equation.
-\item Do not mix complete spellings and abbreviations of units: ``Wb/m\textsuperscript{2}'' or ``webers per square meter'', not ``webers/m\textsuperscript{2}''.
-Spell out units when they appear in text: ``.
-.
-.
-a few henries'', not ``.
-.
-.
-a few H''.
-\item Use a zero before decimal points: ``0.25'', not ``.25''. Use ``cm\textsuperscript{3}'', not ``cc''.)
-\end{itemize}
-
-\subsection{Equations}
-Number equations consecutively. To make your 
-equations more compact, you may use the solidus (~/~), the exp function, or 
-appropriate exponents. Italicize Roman symbols for quantities and variables, 
-but not Greek symbols. Use a long dash rather than a hyphen for a minus 
-sign. Punctuate equations with commas or periods when they are part of a 
-sentence, as in:
-\begin{equation}
-a+b=\gamma\label{eq}
-\end{equation}
-
-Be sure that the 
-symbols in your equation have been defined before or immediately following 
-the equation. Use ``\eqref{eq}'', not ``Eq.~\eqref{eq}'' or ``equation \eqref{eq}'', except at 
-the beginning of a sentence: ``Equation \eqref{eq} is . . .''
-
-\subsection{\LaTeX-Specific Advice}
-
-Please use ``soft'' (e.g., \verb|\eqref{Eq}|) cross references instead
-of ``hard'' references (e.g., \verb|(1)|). That will make it possible
-to combine sections, add equations, or change the order of figures or
-citations without having to go through the file line by line.
-
-Please don't use the \verb|{eqnarray}| equation environment. Use
-\verb|{align}| or \verb|{IEEEeqnarray}| instead. The \verb|{eqnarray}|
-environment leaves unsightly spaces around relation symbols.
-
-Please note that the \verb|{subequations}| environment in {\LaTeX}
-will increment the main equation counter even when there are no
-equation numbers displayed. If you forget that, you might write an
-article in which the equation numbers skip from (17) to (20), causing
-the copy editors to wonder if you've discovered a new method of
-counting.
-
-{\BibTeX} does not work by magic. It doesn't get the bibliographic
-data from thin air but from .bib files. If you use {\BibTeX} to produce a
-bibliography you must send the .bib files.
-
-{\LaTeX} can't read your mind. If you assign the same label to a
-subsubsection and a table, you might find that Table I has been cross
-referenced as Table IV-B3.
-
-{\LaTeX} does not have precognitive abilities. If you put a
-\verb|\label| command before the command that updates the counter it's
-supposed to be using, the label will pick up the last counter to be
-cross referenced instead. In particular, a \verb|\label| command
-should not go before the caption of a figure or a table.
-
-Do not use \verb|\nonumber| inside the \verb|{array}| environment. It
-will not stop equation numbers inside \verb|{array}| (there won't be
-any anyway) and it might stop a wanted equation number in the
-surrounding equation.
-
-\subsection{Some Common Mistakes}\label{SCM}
-\begin{itemize}
-\item The word ``data'' is plural, not singular.
-\item The subscript for the permeability of vacuum $\mu_{0}$, and other common scientific constants, is zero with subscript formatting, not a lowercase letter ``o''.
-\item In American English, commas, semicolons, periods, question and exclamation marks are located within quotation marks only when a complete thought or name is cited, such as a title or full quotation.
-When quotation marks are used, instead of a bold or italic typeface, to highlight a word or phrase, punctuation should appear outside of the quotation marks.
-A parenthetical phrase or statement at the end of a sentence is punctuated outside of the closing parenthesis (like this).
-(A parenthetical sentence is punctuated within the parentheses.)
-\item A graph within a graph is an ``inset'', not an ``insert''. The word alternatively is preferred to the word ``alternately'' (unless you really mean something that alternates).
-\item Do not use the word ``essentially'' to mean ``approximately'' or ``effectively''.
-\item In your paper title, if the words ``that uses'' can accurately replace the word ``using'', capitalize the ``u''; if not, keep using lower-cased.
-\item Be aware of the different meanings of the homophones ``affect'' and ``effect'', ``complement'' and ``compliment'', ``discreet'' and ``discrete'', ``principal'' and ``principle''.
-\item Do not confuse ``imply'' and ``infer''.
-\item The prefix ``non'' is not a word; it should be joined to the word it modifies, usually without a hyphen.
-\item There is no period after the ``et'' in the Latin abbreviation ``et al.''.
-\item The abbreviation ``i.e.'' means ``that is'', and the abbreviation ``e.g.'' means ``for example''.
-\end{itemize}
-An excellent style manual for science writers is \cite{b7}.
-
-\subsection{Authors and Affiliations}
-\textbf{The class file is designed for, but not limited to, six authors.} A 
-minimum of one author is required for all conference articles. Author names 
-should be listed starting from left to right and then moving down to the 
-next line. This is the author sequence that will be used in future citations 
-and by indexing services. Names should not be listed in columns nor group by 
-affiliation. Please keep your affiliations as succinct as possible (for 
-example, do not differentiate among departments of the same organization).
-
-\subsection{Identify the Headings}
-Headings, or heads, are organizational devices that guide the reader through 
-your paper. There are two types: component heads and text heads.
-
-Component heads identify the different components of your paper and are not 
-topically subordinate to each other. Examples include Acknowledgments and 
-References and, for these, the correct style to use is ``Heading 5''. Use 
-``figure caption'' for your Figure captions, and ``table head'' for your 
-table title. Run-in heads, such as ``Abstract'', will require you to apply a 
-style (in this case, italic) in addition to the style provided by the drop 
-down menu to differentiate the head from the text.
-
-Text heads organize the topics on a relational, hierarchical basis. For 
-example, the paper title is the primary text head because all subsequent 
-material relates and elaborates on this one topic. If there are two or more 
-sub-topics, the next level head (uppercase Roman numerals) should be used 
-and, conversely, if there are not at least two sub-topics, then no subheads 
-should be introduced.
-
-\subsection{Figures and Tables}
-\paragraph{Positioning Figures and Tables} Place figures and tables at the top and 
-bottom of columns. Avoid placing them in the middle of columns. Large 
-figures and tables may span across both columns. Figure captions should be 
-below the figures; table heads should appear above the tables. Insert 
-figures and tables after they are cited in the text. Use the abbreviation 
-``Fig.~\ref{fig}'', even at the beginning of a sentence.
-
-\begin{table}[htbp]
-\caption{Table Type Styles}
-\begin{center}
-\begin{tabular}{|c|c|c|c|}
-\hline
-\textbf{Table}&\multicolumn{3}{|c|}{\textbf{Table Column Head}} \\
-\cline{2-4} 
-\textbf{Head} & \textbf{\textit{Table column subhead}}& \textbf{\textit{Subhead}}& \textbf{\textit{Subhead}} \\
-\hline
-copy& More table copy$^{\mathrm{a}}$& &  \\
-\hline
-\multicolumn{4}{l}{$^{\mathrm{a}}$Sample of a Table footnote.}
-\end{tabular}
-\label{tab1}
-\end{center}
-\end{table}
-
-\begin{figure}[htbp]
-\centerline{\includegraphics{fig1.png}}
-\caption{Example of a figure caption.}
-\label{fig}
-\end{figure}
-
-Figure Labels: Use 8 point Times New Roman for Figure labels. Use words 
-rather than symbols or abbreviations when writing Figure axis labels to 
-avoid confusing the reader. As an example, write the quantity 
-``Magnetization'', or ``Magnetization, M'', not just ``M''. If including 
-units in the label, present them within parentheses. Do not label axes only 
-with units. In the example, write ``Magnetization (A/m)'' or ``Magnetization 
-\{A[m(1)]\}'', not just ``A/m''. Do not label axes with a ratio of 
-quantities and units. For example, write ``Temperature (K)'', not 
-``Temperature/K''.
-
-\section*{Acknowledgment}
-
-The preferred spelling of the word ``acknowledgment'' in America is without 
-an ``e'' after the ``g''. Avoid the stilted expression ``one of us (R. B. 
-G.) thanks $\ldots$''. Instead, try ``R. B. G. thanks$\ldots$''. Put sponsor 
-acknowledgments in the unnumbered footnote on the first page.
-
-\section*{References}
-
-Please number citations consecutively within brackets \cite{b1}. The 
-sentence punctuation follows the bracket \cite{b2}. Refer simply to the reference 
-number, as in \cite{b3}---do not use ``Ref. \cite{b3}'' or ``reference \cite{b3}'' except at 
-the beginning of a sentence: ``Reference \cite{b3} was the first $\ldots$''
-
-Number footnotes separately in superscripts. Place the actual footnote at 
-the bottom of the column in which it was cited. Do not put footnotes in the 
-abstract or reference list. Use letters for table footnotes.
-
-Unless there are six authors or more give all authors' names; do not use 
-``et al.''. Papers that have not been published, even if they have been 
-submitted for publication, should be cited as ``unpublished'' \cite{b4}. Papers 
-that have been accepted for publication should be cited as ``in press'' \cite{b5}. 
-Capitalize only the first word in a paper title, except for proper nouns and 
-element symbols.
-
-For papers published in translation journals, please give the English 
-citation first, followed by the original foreign-language citation \cite{b6}.
-
-\begin{thebibliography}{00}
-\bibitem{b1} G.
-Eason, B.
-Noble, and I.
-N.
-Sneddon, ``On certain integrals of Lipschitz-Hankel type involving products of Bessel functions,'' Phil.
-Trans.
-Roy.
-Soc.
-London, vol.
-A247, pp.
-529--551, April 1955.
-\bibitem{b2} J. Clerk Maxwell, A Treatise on Electricity and Magnetism, 3rd ed., vol. 2. Oxford: Clarendon, 1892, pp.68--73.
-\bibitem{b3} I. S. Jacobs and C. P. Bean, ``Fine particles, thin films and exchange anisotropy,'' in Magnetism, vol. III, G. T. Rado and H. Suhl, Eds. New York: Academic, 1963, pp. 271--350.
-\bibitem{b4} K. Elissa, ``Title of paper if known,'' unpublished.
-\bibitem{b5} R. Nicole, ``Title of paper with only first word capitalized,'' J. Name Stand. Abbrev., in press.
-\bibitem{b6} Y.
-Yorozu, M.
-Hirano, K.
-Oka, and Y.
-Tagawa, ``Electron spectroscopy studies on magneto-optical media and plastic substrate interface,'' IEEE Transl.
-J.
-Magn.
-Japan, vol.
-2, pp.
-740--741, August 1987 [Digests 9th Annual Conf.
-Magnetics Japan, p.
-301, 1982].
-\bibitem{b7} M. Young, The Technical Writer's Handbook. Mill Valley, CA: University Science, 1989.
-\end{thebibliography}
 

### 2024-02-28 paper/main.tex
Technical approach introductory revised

@@
 %% Paper about super resolution
-Another application of this algorithm is used for image inpainting by \cite{pg-image-inpaint-2009}.
-They introduce the PG for this specific scenerio but it has some precautions to use for us.
-The images they used are extremely dense when comparing with the LiDAR.
-Also, they have single sensory image but we need to optain a fusion.
+An application of this algorithm for image inpainting was proposed by \cite{pg-image-inpaint-2009}, wherein the Papoulis-Gerchberg (PG) algorithm is adapted for this particular scenario.
+However, there are several caveats to its application within our context.
+Primarily, the images utilized in their study exhibit significantly higher density compared to LiDAR data.
+Furthermore, their methodology is predicated upon the utilization of a singular sensory modality, whereas our objective necessitates the achievement of fusion across multiple sensors.
 
-Another work done by \cite{ozbay2015high} that introduce an extended version of PG for 3D point cloud data that is driven by a LiDAR sensor.
-They use this as an upsampling to cope with delayed points when using LiDAR in high resolution mode due to the low scan rates of cylindirical scanning ones.
-But appearantly, it also use single sensory system
+Another noteworthy contribution is made by \cite{ozbay2015high}, presenting an enhanced iteration of the PG algorithm tailored for 3D point cloud data derived from LiDAR sensors.
+Their method focuses on upsampling to mitigate the latency in point acquisition observed when operating LiDAR at high-resolution modes, attributable to the inherently slow scan rates of cylindrical scanning systems.
+Nevertheless, it is important to note that this approach, too, relies on a singular sensory system.
 
-We introduce our novel approach my putting another extension of PG to the litarature that is the first one uses multi-sensory system. The nature of PG consists some elements.
+The utilization of the Papoulis-Gerchberg (PG) algorithm within our framework is primarily motivated by its deterministic nature, setting it apart from machine learning-based methodologies that often rely on probabilistic outcomes.
+A closer examination of the PG algorithm's underlying principles reveals two pivotal components: a densely populated base image intended for gap-filling using points of known accuracy, and the employment of these high-confidence points for iterative refinement.
+This paradigm bears a striking resemblance to the challenges inherent in sensor fusion.
+In our methodology, stereo depth images serve as the densely populated base analogous to the PG algorithm's framework, while LiDAR sparse data represent the high-confidence points.
+This analogy forms the cornerstone of our approach, leveraging the PG algorithm's architecture to integrate multi-sensorial data, thereby pioneering a novel application of PG in the realm of sensor fusion.
 

### 2024-02-28 paper/main.tex
Pre-Processing subsection introduction revised

@@
 \end{enumerate}
+
 These two steps are repeated iteratively.
@@
 
-Due to the fundamental difference in the representation of depth information between LiDAR measurements (point cloud) and stereo depth estimation (dense disparity maps), both data modalities necessitate preprocessing steps involving transformation, conversion, and upsampling to facilitate their integration and subsequent processing.
+The fusion of data from heterogeneous sensor systems, such as LiDAR and stereo cameras, presents a unique set of challenges, primarily due to the fundamental differences in their respective depth information representations.
+LiDAR sensors provide sparse but highly accurate point clouds, whereas stereo vision techniques yield dense disparity maps with varying degrees of precision.
+To effectively integrate these diverse data modalities, a comprehensive pre-processing regimen is indispensable.
+This regimen encompasses transformation, conversion, and upsampling procedures designed to standardize the data, thereby laying the groundwork for their combined analysis.
 
-REVISE - To make these two sensor data processable in a domain, we determine a common modality for these images.
-This modality is single channel depth image format.
-The RGB integration wont necessite for using our algorithm.
+Transformation and Conversion: The initial step involves transforming the LiDAR-generated point clouds and stereo vision-produced disparity maps into a uniform format.
+We opt for the single-channel depth image format as this common ground facilitates the merging process.
+This transformation is critical not only for aligning the data spatially but also for ensuring that subsequent processing stages operate on data that are consistent in terms of dimensionality and scale.
+
+Upsampling: Given the sparse nature of LiDAR data compared to the relatively denser outputs from stereo vision, upsampling is employed to increase the resolution of the LiDAR measurements.
+This step is crucial for achieving a density comparable to that of the stereo disparity maps, thereby enhancing the reliability of the fused depth information.
+Upsampling is carefully executed to preserve the integrity of the LiDAR data, ensuring that the enhanced resolution does not compromise the accuracy inherent to LiDAR measurements.
+
+RGB Integration Exclusion: It is worth noting that our approach does not necessitate the integration of RGB data.
+This decision streamlines the preprocessing phase, focusing solely on depth information, which is paramount for our application.
+The exclusion of RGB data underscores our methodology's emphasis on depth accuracy and computational efficiency, aligning with the objective of developing a lightweight and practical fusion algorithm.
+
+By meticulously addressing these preprocessing steps, we ensure that the LiDAR and stereo depth data are optimally prepared for fusion.
+This preparation is vital for the successful application of our novel algorithm, which leverages the strengths of both sensor types to produce depth estimations of unparalleled accuracy and density.
 
@@
 
-\paragraph{Obtaining Depth Frame From LiDAR Data}
+\subsubsection{Obtaining Depth Frame From LiDAR Data}
 
@@
 
-\paragraph{LiDAR Null Mask}
+\subsubsection{LiDAR Null Mask}
 

### 2024-02-28 paper/main.tex
Detailed explainations added to the pre-proc section

@@
 
+\subsubsection{Obtaining Depth Frame From LiDAR Data}
+
+Subsequent to the calibration of the stereo camera and LiDAR sensor, the transformation matrices essential for converting point cloud data into a depth image are established.
+The conversion employs the pinhole camera model to accurately project three-dimensional points onto a two-dimensional depth frame space.
+This process is fundamental in translating the spatial coordinates of LiDAR measurements into a format that is compatible with stereo vision data, thereby facilitating their integration.
+
 \subsubsection{Data Cropping}
 
-After projecting the LiDAR image onto the stereo depth image plane, there are some remaining problems.
-One of them is the boundaries of the depth regions.
-Since the vertical scan area of the lidar sensor is pretty small (like 30 degrees) we have to crop the stereo depth image also.
-On the other hand, the stereo depth image have a bound smaller than the lidar scan at the on the horzontal axis.
-This is appearent, because the lidar horizontal scan area is 360 degrees.
-Since our algorithm is based on depth images and there are no memories for previously detected points for the frame sensor fusioning, we just ignore this non-overlapping depth data regions.
+Projection of LiDAR data onto the stereo depth image plane reveals discrepancies in the boundary extents of the depth regions.
+This is particularly evident in the vertical scan range of the LiDAR sensor, which is relatively narrow (approximately 30 degrees) compared to the broader field of view captured by the stereo camera.
+Consequently, it becomes necessary to crop the stereo depth image to match the LiDAR's scan area.
+Moreover, the stereo depth image exhibits a narrower horizontal boundary compared to the LiDAR's 360-degree scanning capability.
+Given our algorithm's reliance on depth images and the absence of memory for previously detected points in frame sensor fusion, non-overlapping depth data regions are disregarded to maintain algorithmic integrity and focus on the fusion of congruent depth information.
 
-\subsubsection{Data Augmentation}
+\subsubsection{Data Augmentation (Stereo Infill)}
 
-Stereo depth augmentation requirement is originated from the deficit nature of the disparity maps that are used in depth estimation from stereo cameras.
-The depth images, due to the disparity maps, have some null values at blind spots that one of the cameras have no view at all.
-These null values at the depth images should be pre-processed before the PG algorithm to avoid the distortion of the well estimated depth entities (pixels).
+The augmentation of stereo depth data addresses the challenge posed by inherent limitations in stereo vision disparity maps, which frequently exhibit null values in occluded regions not visible to one or both of the stereo cameras.
+These occlusions result in incomplete depth information, manifesting as gaps in the depth images derived from stereo vision.
+To rectify this and ensure a comprehensive depth field, our method incorporates a strategic process of gap filling that leverages high-confidence LiDAR measurements.
 
-We proposed a method to overcome the null value problem.
-This method fills the gaps (null values at the depth images) by using the LiDAR measurements.
-Firstly, a segmentation of the gaps are made and then the LiDAR measurements are taken into account to fill the gap segments that are overlapped with LiDAR measurements.
-Again, due to the sparse nature of the LiDAR measurements, we made a interpolation for filling the huge gaps by using LiDAR measurements that have small area.
+Gap Segmentation and Identification: The initial phase involves a meticulous analysis of the stereo depth images to segment and identify areas devoid of depth information—these are the regions where disparity maps fail to provide reliable depth due to occlusions or insufficient texture.
 
-\subsubsection{Obtaining Depth Frame From LiDAR Data}
+LiDAR Measurement Integration: Upon identifying these gaps, the process proceeds by overlaying LiDAR data onto the stereo depth images.
+This step utilizes the precise depth information from LiDAR measurements to fill in the identified gaps, prioritizing regions where LiDAR data overlaps with the gaps in stereo images.
+This integration capitalizes on the accuracy and reliability of LiDAR measurements to enhance the fidelity of the depth information in areas previously marked by uncertainty.
 
-After making the calibrations for stereo camera and the LiDAR, we know all the transformation matrices to generate the depth image from point cloud data.
-To project the 3D points into the 2D depth frame space we use the pinhole camera model.
+Interpolation for Comprehensive Coverage: Recognizing the sparse nature of LiDAR data and its inability to cover all gaps fully, we employ interpolation techniques as a subsequent measure.
+This approach extrapolates the available LiDAR measurements to fill larger gaps, ensuring that even areas without direct LiDAR coverage receive enhanced depth estimations.
+The interpolation considers the spatial distribution of LiDAR points and the surrounding depth context to generate a coherent depth field that bridges the disparity-induced gaps with plausible depth values.
+
+This augmented approach to stereo depth infill not only rectifies the limitations inherent in stereo disparity maps but also enhances the overall depth accuracy of the fused stereo-LiDAR image.
+By meticulously integrating and interpolating LiDAR data within the stereo depth framework, we achieve a more robust and complete representation of the scene's depth, paving the way for improved performance in subsequent processing stages of our sensor fusion algorithm.
+
+This phase represents a pivotal component of our preprocessing strategy, primarily due to its critical role in safeguarding the integrity of the Papoulis-Gerchberg (PG) algorithm's iterative process.
+In the absence of this meticulous data augmentation step, the iterative nature of the PG algorithm could inadvertently exacerbate depth estimation errors with each subsequent iteration.
+Specifically, unaddressed gaps in the stereo depth images could lead to erroneous propagation and amplification of inaccuracies, undermining the algorithm's ability to refine and converge towards a precise depth map.
+Therefore, by ensuring that these gaps are effectively filled through the integration and interpolation of LiDAR data, we substantially enhance the stability and accuracy of the depth estimation process, particularly as the number of iterations increases.
+This augmentation is not merely a preparatory measure but a fundamental prerequisite for harnessing the full potential of the PG algorithm in achieving high-fidelity depth estimations from fused stereo-LiDAR data.
 
@@
 
-The LiDAR depth frames will have null values do to the nature of the projection of sparse point cloud data into a plane.
-This null values should be considered before using the LiDAR depth map in the algorithm.
-
-Since, our method use the stereo depth maps as a starting point and LiDAR depth map as rectifier of depth information of these stereo depth images, we need a mask of null values to eliminate the spurious pixels (depth entities) at the post-processing part of our algorithm.
-
-\paragraph{Stereo Infill}
+The projection of sparse LiDAR point cloud data onto a two-dimensional plane inherently results in the presence of null values within the depth frames.
+These null values must be meticulously accounted for prior to the integration of LiDAR depth maps into our fusion algorithm.
+To address this, we employ a null value mask for the LiDAR depth data, which serves to identify and subsequently exclude spurious depth entities from the fused depth estimation.
+This preprocessing step is crucial for refining the depth information derived from stereo depth maps, utilizing the LiDAR data as a corrective measure to enhance the fidelity of the resultant depth image.
 
@@
 \paragraph{Use of Band Reject Filter in PG}
+As the nature of the low pass filters used in papoulis-gerchberg algorithm, they all eliminate the high frequency components as they meant to be.
+But in our case, our data also need to be having high frequency components
 \paragraph{Evaluated PG for Our Work}

### 2024-02-28 paper/main.tex
Processing part introduced up to Band-Reject section

@@
 As the nature of the low pass filters used in papoulis-gerchberg algorithm, they all eliminate the high frequency components as they meant to be.
-But in our case, our data also need to be having high frequency components
+But in our case, the data shouldnt be losing these high frequency components.
+Normally papoulis-gerchberg algorithm assumes a continuous signal can't have sharp transitions.
+Sharp transitions meant to be losses of the signal, so papoulis-gerchberg algorithm tries to reconstruct the image by eliminating high frequency components.
+In our case the high frequency components or sharp transitions meant to be the transition between foreground to background, so we don't want to lose this information.
+To overcome this problem we proposed to use band-reject filter that allows higher and lower frequency components to pass, only mid frequencies to reject.
+mid frequencies assumed to be the differences between LiDAR measurements and estimated depth information from stereo pair.
+Iteratively applying this filter and placing Lidar points as anchors makes the estimation closer to the lidar's accuracy.
+
 \paragraph{Evaluated PG for Our Work}

### 2024-02-29 paper/main.tex
Update main.tex

--- 2024-02-28 paper/main.tex
@@
 \section{Technical Approach}
-%% In this section explain the each step of the processing detaily.
+%% In this section explain each step of the processing detaily.
 
@@
 
-The utilization of the Papoulis-Gerchberg (PG) algorithm within our framework is primarily motivated by its deterministic nature, setting it apart from machine learning-based methodologies that often rely on probabilistic outcomes.
+The utilization of the Papoulis-Gerchberg (PG) algorithm within our framework is primarily motivated by its deterministic nature, setting it apart from DNN-based methodologies that often rely on probabilistic outcomes.
 A closer examination of the PG algorithm's underlying principles reveals two pivotal components: a densely populated base image intended for gap-filling using points of known accuracy, and the employment of these high-confidence points for iterative refinement.
@@
 \subsection{PG: Papoulis Gerchberg}
-We first intended to use Papoulish-Gerchberg algorithm as a pre-processing method to up sample the only LiDAR data. But the sparsity of the LiDAR data made it challenging to work with.
 
-We tried to apply zero-padding to empty regions of the depth frame obtained by the LiDAR samples.
-Since the Papoulish-Gerchberg method is an iterative method, zero background filling caused high amount of iterations to get reasonable results.
-And due to the sparse data especially in the vertical axis, the regions far from original LiDAR samples never get good results like the closer ones.
+Initially, our approach sought to leverage the Papoulis-Gerchberg algorithm as a method for pre-processing, specifically aimed at upsampling LiDAR data exclusively.
+However, the intrinsic sparsity of LiDAR measurements presented substantial challenges.
+The iterative nature of the Papoulis-Gerchberg method, when applied to areas devoid of data—filled with zero-padding—resulted in an excessively high number of iterations before achieving satisfactory outcomes.
+Moreover, the sparse nature of the data, particularly along the vertical axis, led to inadequate enhancement in regions distant from the original LiDAR samples, in stark contrast to areas proximal to these samples.
+
+In an effort to mitigate the iteration count, we experimented with initializing the empty regions with the mean depth value derived from the actual LiDAR samples.
+This approach, leveraging a more representative initial state, markedly reduced the number of necessary iterations by providing a closer approximation to the true depth values of the surrounding area.
+Nonetheless, the fundamental issue posed by the sparsity of the LiDAR data persisted.
 %% add an image to represent the papoulis with zero-padding
 
-Then to decrease number of iterations, we tried to initialize the background with mean of actual LiDAR samples.
-Since the mean value is much closer to the actual depth of the neighboring samples, it significantly reduced the iterations.
-But the problem with the sparsity of LiDAR data was still present.
+Addressing this challenge required a more robust solution.
+We turned to the work of our colleague, Ismail Uyanik, specifically his pioneering method, "Enhancing 3D range image measurement density via the dynamic Papoulis-Gerchberg algorithm." This approach, coupled with our mean value initialization strategy, offered a comprehensive solution.
+By incorporating a dynamic iteration of the PG algorithm, we significantly enhanced our ability to refine the depth data, particularly in sparsely measured areas, thereby overcoming the limitations previously encountered with LiDAR data sparsity.
 %% add another figure for mean filling
 
-We need more significant data to overcome this problem.
+% We need more significant data to overcome this problem.
 The initial solution was the using previous work of our colleague @ismail-uyanik, {Enhancing 3D range image measurement density via dynamic Papoulis-Gerchberg algorithm} along with the mean filling method.
+
+% BU DA OLABİLİR. İKİSİNİ DE GPT YAZDI
+% Initially, our strategy entailed leveraging the Papoulis-Gerchberg (PG) algorithm predominantly as a pre-processing technique to upscale LiDAR data exclusively.
+However, the inherent sparsity of LiDAR measurements posed significant challenges, particularly in maintaining the fidelity of depth information across the depth frame.
+
+% In an attempt to address the sparse nature of the LiDAR data, we explored the application of zero-padding to the vacant regions within the depth frames derived from LiDAR samples.
+Despite the iterative prowess of the PG algorithm, the introduction of a zero-value background substantially increased the number of iterations required to achieve satisfactory outcomes.
+More critically, due to the pronounced sparsity along the vertical axis, areas distant from the original LiDAR points consistently yielded inferior results compared to those in closer proximity.
+
+% To mitigate the excessive iteration counts, an alternative initialization strategy was considered, involving the use of the mean depth value from actual LiDAR samples as a background filler.
+This approach, by aligning the initial background estimation closer to the real depth values of neighboring points, markedly reduced the necessary iterations for convergence.
+Nonetheless, the fundamental issue of data sparsity remained inadequately addressed.
+
+% The breakthrough came with the incorporation of a technique developed by our colleague, Ismail Uyanik, titled "Enhancing 3D Range Image Measurement Density via Dynamic Papoulis-Gerchberg Algorithm." By integrating this method with our mean filling approach, we were able to significantly augment the density of the depth data.
+This synergy not only addressed the sparsity challenge but also empowered the PG algorithm to reconstruct depth frames with enhanced accuracy and reduced computational overhead, laying a robust foundation for the subsequent fusion process.
+
 %% cite the paper from kuzucu-elvan
@@
 \subsubsection{Our Approach}
-As the purpose of this study is to investigate the potential of integrating data obtained from stereo cameras and LiDAR measurements, we also have the depth information from stereo image pair.
-Our motivation of this paper's work turned to merge both processes, up sampling of LiDAR data and fusing the depth information obtained from both sources.
-This method extremely decreased the need of processing power, and allowed real-time application of the sensor fusion even in mobile computing devices.
-Basically the method is nothing but proposing the background of lidar measurements as depth frame from stereo pair instead of using only a mean value.
-Since the background filling is not non-significant data, we have mostly overcame the problem with sparsity of LiDAR measurements.
-We also made some modifications and improvements to the standard 2D Papoulish-Gerchberg algorithm to match our needs better.
+
+The primary objective of this research is to explore the integration of depth information derived from stereo cameras and LiDAR measurements.
+Our approach innovatively combines the upsampling of LiDAR data with the fusion of depth information from both sources.
+This methodology significantly reduces the computational demands, enabling the real-time application of sensor fusion on mobile computing platforms.
+
+At the core of our method is the strategy of utilizing the depth frame obtained from the stereo image pair as the background for LiDAR measurements, rather than relying solely on a mean value for background estimation.
+This approach effectively addresses the issue of LiDAR data sparsity by providing a rich, relevant context for the depth information.
+Consequently, this significantly enhances the depth frame's completeness and accuracy.
+
+Furthermore, we have adapted and refined the traditional 2D Papoulis-Gerchberg algorithm to better suit the unique requirements of our sensor fusion task.
+
 \paragraph{Use of Band Reject Filter in PG}
-As the nature of the low pass filters used in papoulis-gerchberg algorithm, they all eliminate the high frequency components as they meant to be.
-But in our case, our data also need to be having high frequency components
+
+A critical modification in our adaptation of the Papoulis-Gerchberg algorithm involves the application of a band-reject filter.
+The inherent design of low-pass filters within the PG algorithm is to attenuate high-frequency components, under the assumption that a continuous signal does not encompass abrupt transitions.
+Such transitions are typically interpreted as signal loss, prompting the PG algorithm to attempt image reconstruction by excluding high-frequency details.
+
+However, in the context of depth estimation, these high-frequency components or sharp transitions are indicative of the boundaries between foreground and background elements, carrying essential spatial information that we aim to preserve.
+To retain this critical detail, we introduce a band-reject filter that selectively allows the passage of both higher and lower frequency components while filtering out mid-frequency ranges.
+These mid-frequencies are presumed to represent discrepancies between LiDAR measurements and the depth estimates from the stereo pair.
+
+Through iterative application of this filter, coupled with the strategic placement of LiDAR points as accuracy anchors, our method progressively refines the depth estimation, aligning it more closely with the inherent precision of LiDAR data.
+This tailored filtering approach, therefore, plays a pivotal role in mitigating the fusion process's challenges, ensuring that the transition zones between different depth regions are accurately maintained, thereby enhancing the overall fidelity of the fused depth map.
+
 \paragraph{Evaluated PG for Our Work}
@@
 \subsubsection{Sensor Suit}
+We conducted experiments using the data we collected from our sensors.
+We used Velodyne VLP-16 as 3D LiDAR and Zed Stereo Camera as main sensing devices.
+We also developed a suit that mounts the sensors we use for experimenting in fixed positions.
+The sensor suit has also an Nvidia Jetson Nano 4GB computer.
+The computer plays different roles in different experiment scenarios.
+% put images of sensor suit only
+To obtain real-life like conditions by mounting the sensor suit on top of a quadruped robot. 
+% put images of sensor suit with robot
 \paragraph{Upsampling of LiDAR Data}
@@
 \subsubsection{Experiment Scenarios}
+\paragraph{Cloud Computing (Performance):}
+\paragraph{Mobile Computing (Performance):}
+\paragraph{SLAM (Quality):}
+% Put three slam maps and visual odometry visualization images {Only Stereo}, {Only Lidar}, {Fused}. AIM:Compare map quality and odomerty accuracy.
 

### 2024-02-29 paper/main.tex
Post-proc section modified

@@
 
-\subsubsection{LiDAR Null Mask}
+\subsubsection{LiDAR Null Mask} \label{lidar-null-mask}
 
@@
 %% add another image explaining the plan
-\subsubsection{Our Approach}
+\paragraph{Our Approach}
 
@@
 
-\paragraph{Evaluated PG for Our Work}
+% \paragraph{Evaluated PG for Our Work}
 
-\subsection{Post-Proc}
+\subsection{Post-Processing}
 
-The post-processing consinst of two major parts
+The post-processing phase is designed to refine the fused depth data, ensuring its accuracy and consistency. This phase is divided into two primary components:
 
 \subsubsection{Removal of Trivial Augmented Data}
+
+A key step involves utilizing the stereo null mask to identify and eliminate any trivial or less reliable data that was introduced during the data augmentation process.
+Specifically, depth estimations that fall within areas masked by the stereo null mask are removed from the final depth frame.
+This mask effectively delineates regions where the stereo data lacked sufficient information, ensuring that only depth data with a high confidence level, either directly measured or accurately estimated, is retained in the final output.
+
+\subsubsection{Enhancement and Smoothing}
+
+To further improve the quality of the fused depth maps, an enhancement and smoothing process is applied.
+This involves the use of filtering techniques to reduce noise and smooth out rough edges, enhancing the overall visual quality of the depth map.
+Techniques such as Gaussian smoothing or median filtering can be particularly effective in achieving a more coherent depth representation, especially in areas where the fusion of stereo and LiDAR data may produce discontinuities or artifacts.
+
+\subsubsection{Final Depth Consistency Check}
+
+Another critical component of the post-processing phase is the final depth consistency check.
+This step involves comparing the processed depth map against the original LiDAR and stereo data to identify and correct any discrepancies.
+Discrepancy correction can be performed through a combination of automated threshold-based adjustments and manual inspections, ensuring that the final depth map faithfully represents the real-world scene.
+
+\subsubsection{Edge Sharpening}
+
+Lastly, edge sharpening techniques are applied to restore and emphasize the definition of edges that may have been softened during the smoothing process.
+This is particularly important for maintaining the structural integrity of objects within the scene, allowing for clear delineation between different entities and the background.
+Techniques such as unsharp masking or adaptive sharpening filters can be employed to enhance edge clarity without introducing noise or artifacts.
 
@@
 
+\vspace{12pt}
+\color{red}
+Tüm referanslar kontrol edilmeli (label referansları dahil)
+
 \end{document}

### 2024-02-29 paper/main.tex
Implementation and optimization parts revised

@@
 
+To provide a comprehensive explanation of the algorithm's implementation and its adaptation to linear algebra, let's expand on the structure you've outlined, detailing both the building and adaptation phases with clarity and technical depth.
+
 \section{Implementation} \label{implementation}
 
-The implementation of a proposed method is essential for experiments.
-Because of the robotic fields requirements, there are some constraints about the frequency of the algorithm, data rate of te sensors, and some other important facts.
-To accomplish all of this constraints, we use the power of the mathematics and the linear algebra.
+The successful implementation of our proposed sensor fusion method is paramount, particularly given the stringent performance and efficiency requirements characteristic of robotic applications.
+These requirements include, but are not limited to, the need for high-frequency processing capabilities, efficient handling of data rates from multiple sensors, and adherence to computational constraints.
+To navigate these challenges, we harness the computational efficiency of linear algebra and the precision of mathematical formulations.
 
@@
 
+The construction of our algorithm is underpinned by a modular design philosophy, facilitating flexibility and scalability. The core components include:
+
+\paragraph{Data Preprocessing Module} 
+This initial module is responsible for standardizing and preparing the input data from stereo cameras and LiDAR sensors.
+It includes procedures for data transformation, cropping, augmentation, and the application of null masks, as detailed in the pre-processing section.
+
+\paragraph{Fusion Core}
+At the heart of our method lies the fusion core, which employs the adapted Papoulis-Gerchberg algorithm.
+This component integrates the pre-processed stereo and LiDAR data, leveraging iterative refinement to enhance the depth map's resolution and accuracy.
+
+\paragraph{Post-Processing Module}
+Following the fusion process, this module applies several refinement techniques, including the removal of trivial augmented data, smoothing, consistency checks, and edge sharpening, ensuring the output depth map is of the highest quality and ready for application use.
+
 \subsection{Adopting the Algorithm to Linear Algebra}
+
+The adaptation of our algorithm to the principles of linear algebra is critical for achieving the desired computational efficiency and scalability. This adaptation involves:
+
+\paragraph{Matrix Representation of Data} 
+All input data, including depth frames from stereo images and LiDAR point clouds, are represented as matrices.
+This format facilitates the application of linear transformations and operations, streamlining the fusion process.
+
+\paragraph{Linear Transformations for Data Fusion}
+The core of the data fusion process relies on linear transformations, which are applied to align, scale, and integrate the stereo and LiDAR data within a unified depth framework.
+These transformations are mathematically defined to ensure precision and efficiency.
+
+\paragraph{Iterative Refinement via Linear Operations} 
+The iterative refinement process within the Papoulis-Gerchberg algorithm is implemented through a series of linear operations.
+This includes the use of linear filters for smoothing and the application of linear equations to model the iterative update mechanism.
+
+% Optimization Techniques: To further enhance performance, linear algebra-based optimization techniques are employed.
+These techniques are designed to reduce computational complexity, improve convergence speed, and ensure the algorithm can be executed in real-time on hardware with limited processing capabilities.
+
+% Parallel Processing and Hardware Acceleration: The algorithm is optimized for parallel processing, taking advantage of hardware acceleration features available in modern computing platforms.
+This optimization is facilitated by the decomposition of linear algebra operations into sub-tasks that can be executed concurrently, significantly reducing processing time.
 
@@
 
-As defined in the previous section \ref{implementation}, our method is developed baed on the linear algebraic methods.
-Using the nature of the linear algebra, we introduce a parallel processing approach for our extended PG method.
-This introduce us a workspace that we can competite with our fellows (Buraya DNN gibi yapıları kullanarak GPU üzerinde koşan sensor fusion paperlarına alıntı yapılacak)
+Building upon the foundational linear algebraic principles outlined in the implementation section (\ref{implementation}), we have further optimized our extended Papoulis-Gerchberg (PG) method to harness parallel processing capabilities.
+This optimization enables our approach to achieve significant computational efficiency, making it competitive with contemporary sensor fusion methods that utilize deep neural networks (DNNs) and other computationally intensive techniques, often reliant on GPU acceleration.
+(Buraya DNN gibi yapıları kullanarak GPU üzerinde koşan sensor fusion paperlarına alıntı yapılacak)
 
 \subsection{GPU Acceleration}
+
+The cornerstone of our optimization strategy is GPU acceleration, which is particularly well-suited to the linear algebra operations at the heart of our method.
+By translating our algorithm's core processes into GPU-optimized computations, we can leverage the massive parallelism offered by modern GPUs.
+This allows for the simultaneous processing of multiple data elements, drastically reducing the time required for depth map fusion and refinement.
+
+Parallel Data Processing: Key operations, such as matrix multiplications and linear transformations, are implemented to take full advantage of GPU threads, ensuring that data from stereo images and LiDAR sensors can be processed concurrently.
+
+Iterative Refinement Optimization: The iterative nature of the PG algorithm benefits immensely from GPU acceleration.
+Each iteration, involving the update and refinement of depth estimations, is executed across thousands of GPU cores, significantly speeding up convergence towards the optimal depth map.
+
+Memory Management and Data Transfer: Efficient memory management techniques are employed to minimize the overhead associated with transferring data between CPU and GPU memory spaces.
+This includes the use of pinned memory and asynchronous data transfer operations to maintain high throughput and low latency.
+
 \subsection{Cloud Computing}

### 2024-09-10 paper/codit2024/main.tex
Paper moved under folder

--- 2024-02-29 paper/main.tex
@@
 
-Parallel Data Processing: Key operations, such as matrix multiplications and linear transformations, are implemented to take full advantage of GPU threads, ensuring that data from stereo images and LiDAR sensors can be processed concurrently.
+\paragraph{Parallel Data Processing} Key operations, such as matrix multiplications and linear transformations, are implemented to take full advantage of GPU threads, ensuring that data from stereo images and LiDAR sensors can be processed concurrently.
 
-Iterative Refinement Optimization: The iterative nature of the PG algorithm benefits immensely from GPU acceleration.
+\paragraph{Iterative Refinement Optimization} The iterative nature of the PG algorithm benefits immensely from GPU acceleration.
 Each iteration, involving the update and refinement of depth estimations, is executed across thousands of GPU cores, significantly speeding up convergence towards the optimal depth map.
 
-Memory Management and Data Transfer: Efficient memory management techniques are employed to minimize the overhead associated with transferring data between CPU and GPU memory spaces.
+\paragraph{Memory Management and Data Transfer} Efficient memory management techniques are employed to minimize the overhead associated with transferring data between CPU and GPU memory spaces.
 This includes the use of pinned memory and asynchronous data transfer operations to maintain high throughput and low latency.
@@
 \subsection{Cloud Computing}
+
+To further enhance the scalability and accessibility of our sensor fusion method, we integrate cloud computing resources.
+This integration not only extends the computational capabilities beyond the confines of local hardware but also facilitates the processing of large datasets and the deployment of our solution in distributed systems.
+
+\paragraph{Scalable Computing Resources} By leveraging cloud-based GPUs and distributed computing environments, our method can dynamically allocate resources based on the complexity of the data and the required processing power, ensuring efficient handling of varying workloads.
+
+\paragraph{Collaborative Processing} The cloud computing model enables collaborative sensor fusion tasks, where data from multiple sources can be integrated and processed in a centralized manner.
+This is particularly advantageous for applications requiring real-time analysis of data from distributed sensor networks.
+
+\paragraph{Deployment Flexibility} The adoption of cloud computing allows for the deployment of our sensor fusion solution across a range of platforms, from mobile devices with limited processing capabilities to enterprise-level systems requiring high throughput and real-time performance.
 

### 2024-09-10 paper/icra2025/main.tex
CoDIT paper forked as ICRA init

--- 2024-09-10 paper/codit2024/main.tex
@@
 \section{Experiment}
+To test the performance and quality of our work, we developed some test methods.
+Although there are widely used datasets to test the stereo and flow depth estimation algorithms, due to the fundamental difference between well known depth estimation algorithms and our work, we had to design custom tests and benchmarks.
+The traditional depth estimation algorithms are mainly focuses on generating disparity maps using stereo pair or flow image pairs, the datasets and benchmarks provide stereo pairs as input data and LiDAR measurements as ground truth.
+Unlike a standard depth estimation method, our multi-sensory system uses also the LiDAR measurements to predict depth.
+We designed different tests to both compare our method with traditional single sensory methods and test it with real life conditions.
+\subsection{Experimenting our Algorithm With KITTI Benchmark Suite}
+The KITTI Vision Benchmark Suite is a well-established resource for researchers and developers working on computer vision algorithms, particularly those related to autonomous driving.
+It provides a diverse set of datasets and evaluation tools for various tasks, including:
 
-\subsection{Experimenting our Algorithm With Kitti Benchmark Suite}
+\paragraph{Stereo vision:} This involves reconstructing 3D information from two slightly different camera viewpoints. KITTI offers datasets for both static and dynamic scenarios.
+\paragraph{Optical flow:} This measures the apparent motion of pixels between consecutive video frames, which is crucial for tasks like motion estimation and scene understanding.
+\paragraph{Visual odometry:} This estimates the 3D motion of a camera solely based on visual information, which is essential for robot navigation.
+3D object detection and tracking: This involves identifying and locating objects of interest (e.g., cars, pedestrians) in 3D space, while tracking their movement over time.
+The suite leverages data captured from a specially equipped vehicle driving in various environments, including urban streets, rural areas, and highways. This data includes:
+
+High-resolution color and grayscale camera images
+Velodyne laser scanner data (LiDAR) for accurate depth information
+GPS localization data
+Researchers can use the KITTI datasets to train and evaluate their algorithms, and compare their performance against others on the provided leaderboards.
+This fosters advancements in computer vision techniques with real-world applications, particularly in the field of autonomous driving.
 %% Mention the dataset preparation picking data from different datasets
@@
 % put images of sensor suit with robot
-\paragraph{Upsampling of LiDAR Data}
-%% Further improvements dynamic LiDAR upsampling with kalman
 
 \subsubsection{Experiment Scenarios}
+We conducted different experiment scenarios to measure different metrics.
 \paragraph{Cloud Computing (Performance):}
+In  this scenario, the 
 \paragraph{Mobile Computing (Performance):}
@@
 \section{Results}
+
+\subsection{Results Obtained From KITTI Benchmark Suite Experiments}
+
+\begin{table}[htbp]
+    \caption{Evaluation of our method and standard mono-sensory methods}
+    \begin{center}
+        \begin{tabular}{|c|c|c|}
+            \hline
+            \textbf{Table}&\multicolumn{2}{|c|}{\textbf{Metrics}} \\
+            \cline{2-3} 
+            \textbf{Head} & \textbf{\textit{MAe}}& \textbf{\textit{RMSe}} \\
+            \hline
+            Stereo Depth& &   \\
+            \hline
+            LiDAR Measurement& More table copy$^{\mathrm{a}}$& \\
+            \hline
+            Ours& More table copy$^{\mathrm{a}}$& \\
+            \hline
+            \multicolumn{4}{l}{$^{\mathrm{a}}$Sample of a Table footnote.}
+        \end{tabular}
+        \label{tab1}
+    \end{center}
+\end{table}
+
+\subsection{Results Obtained From Sensor Suit Experiments}
+
+\begin{table}[htbp]
+\caption{Table Type Styles}
+\begin{center}
+\begin{tabular}{|c|c|c|c|}
+\hline
+\textbf{Table}&\multicolumn{3}{|c|}{\textbf{Table Column Head}} \\
+\cline{2-4} 
+\textbf{Head} & \textbf{\textit{Table column subhead}}& \textbf{\textit{Subhead}}& \textbf{\textit{Subhead}} \\
+\hline
+copy& More table copy$^{\mathrm{a}}$& &  \\
+\hline
+\multicolumn{4}{l}{$^{\mathrm{a}}$Sample of a Table footnote.}
+\end{tabular}
+\label{tab1}
+\end{center}
+\end{table}
 

### 2024-09-10 paper/icra2025/main.tex
ICRA update

@@
 \title{Enhancing Depth Estimation with Iterative Stereo-LiDAR Fusion: An Application of the Papoulis-Gerchberg Algorithm\\
-{\footnotesize \textsuperscript{*}Note: Sub-titles are not captured in Xplore and
-should not be used}
-\thanks{Identify applicable funding agency here. If none, delete this.}
+% {\footnotesize \textsuperscript{*}Note: Sub-titles are not captured in Xplore and should not be used}
+% \thanks{Identify applicable funding agency here. If none, delete this.}
 }
@@
 \author{
-\IEEEauthorblockN{1\textsuperscript{st} Assc. Prof. İsmail Uyanık}
-\IEEEauthorblockA{\textit{Department of Electrical and Electronics Engineering} \\
-\textit{Hacettepe University}\\
-Ankara, Turkiye \\
-uyanik@ee.hacettepe.edu.tr}
-\and
     \IEEEauthorblockN{2\textsuperscript{st} Ahmet Yusuf Şirin}
@@
     ertugrultyk@gmail.com}
+    \and
+    \IEEEauthorblockN{1\textsuperscript{st} Assc. Prof. İsmail Uyanık}
+    \IEEEauthorblockA{\textit{Department of Electrical and Electronics Engineering} \\
+    \textit{Hacettepe University}\\
+    Ankara, Turkiye \\
+    uyanik@ee.hacettepe.edu.tr}
 }
@@
 \begin{abstract}
-Depth perception in autonomous systems is critically enhanced by sensor fusion techniques, which traditionally rely on the combination of stereo vision and LiDAR data.
-However, the challenge of integrating sparse LiDAR data with dense stereo imagery remains.
-This paper introduces a novel fusion approach that applies the Papoulis-Gerchberg algorithm to iteratively refine stereo depth maps with LiDAR data.
-Our method uses LiDAR measurements as anchor points to guide the enhancement of depth information, resulting in superior resolution and accuracy.
-We evaluate our approach using the KITTI benchmark suite, demonstrating significant improvements over existing methods such as FastFusion, particularly in edge definition and object boundary sharpness.
-The results suggest our method not only achieves higher fidelity in depth estimation but also offers a new perspective on sensor data integration for real-time applications in autonomous navigation.
+Depth perception in autonomous systems is critically enhanced through sensor fusion, where the integration of stereo vision and LiDAR data plays a pivotal role.
+However, effectively merging sparse LiDAR measurements with dense stereo imagery remains a key challenge.
+This paper introduces a novel fusion method based on the Papoulis-Gerchberg algorithm, which iteratively refines stereo depth maps using LiDAR data as anchor points.
+By leveraging these anchor points, our approach enhances depth map resolution and accuracy, particularly in areas where traditional methods struggle, such as object boundaries and edge definition.
+
+We validate our method using the KITTI benchmark suite, demonstrating substantial improvements over existing approaches like FastFusion, particularly in edge sharpness and object boundary clarity.
+These results highlight the potential of our method to not only achieve superior fidelity in depth estimation but also to set a new standard for sensor data integration in real-time autonomous navigation.
+
+To test our approach in real-world conditions, we introduce a versatile mobile sensor suite, comprising a Velodyne VLP16 LiDAR, ZED Stereo Camera, and an NVIDIA Jetson Nano for on-board processing.
+This sensor suite is mounted on a legged robot platform to simulate vibrations and disturbances common in military and civilian applications.
+The collected data is processed using the ROS framework, and SLAM algorithms are employed to evaluate the system’s performance, providing a complete test loop for autonomous navigation tasks.
 \end{abstract}
@@
 To effectively integrate these diverse data modalities, a comprehensive pre-processing regimen is indispensable.
-This regimen encompasses transformation, conversion, and upsampling procedures designed to standardize the data, thereby laying the groundwork for their combined analysis.
+This regimen encompasses transformation, conversion, and history based infill procedures designed to standardize the data, thereby laying the groundwork for their combined analysis.
 
@@
 
-Upsampling: Given the sparse nature of LiDAR data compared to the relatively denser outputs from stereo vision, upsampling is employed to increase the resolution of the LiDAR measurements.
-This step is crucial for achieving a density comparable to that of the stereo disparity maps, thereby enhancing the reliability of the fused depth information.
-Upsampling is carefully executed to preserve the integrity of the LiDAR data, ensuring that the enhanced resolution does not compromise the accuracy inherent to LiDAR measurements.
+History Based LiDAR Infill: We use a history based data augmentation for sparse LiDAR data.
+The previous point cloud frames are translated by the negative derivative of the current velocity to infill the empty regions in the point cloud.
+This makes us to integrate already known points to the future estimations.
 
@@
 
-\paragraph{Use of Band Reject Filter in PG}
+% \paragraph{Use of Band Reject Filter in PG}
 
-A critical modification in our adaptation of the Papoulis-Gerchberg algorithm involves the application of a band-reject filter.
+% A critical modification in our adaptation of the Papoulis-Gerchberg algorithm involves the application of a band-reject filter.
 The inherent design of low-pass filters within the PG algorithm is to attenuate high-frequency components, under the assumption that a continuous signal does not encompass abrupt transitions.
@@
 
-However, in the context of depth estimation, these high-frequency components or sharp transitions are indicative of the boundaries between foreground and background elements, carrying essential spatial information that we aim to preserve.
+% However, in the context of depth estimation, these high-frequency components or sharp transitions are indicative of the boundaries between foreground and background elements, carrying essential spatial information that we aim to preserve.
 To retain this critical detail, we introduce a band-reject filter that selectively allows the passage of both higher and lower frequency components while filtering out mid-frequency ranges.
@@
 
-Through iterative application of this filter, coupled with the strategic placement of LiDAR points as accuracy anchors, our method progressively refines the depth estimation, aligning it more closely with the inherent precision of LiDAR data.
+% Through iterative application of this filter, coupled with the strategic placement of LiDAR points as accuracy anchors, our method progressively refines the depth estimation, aligning it more closely with the inherent precision of LiDAR data.
 This tailored filtering approach, therefore, plays a pivotal role in mitigating the fusion process's challenges, ensuring that the transition zones between different depth regions are accurately maintained, thereby enhancing the overall fidelity of the fused depth map.
@@
 
-\subsubsection{Edge Sharpening}
+% \subsubsection{Edge Sharpening}
 
-Lastly, edge sharpening techniques are applied to restore and emphasize the definition of edges that may have been softened during the smoothing process.
+% Lastly, edge sharpening techniques are applied to restore and emphasize the definition of edges that may have been softened during the smoothing process.
 This is particularly important for maintaining the structural integrity of objects within the scene, allowing for clear delineation between different entities and the background.
@@
 \paragraph{Post-Processing Module}
-Following the fusion process, this module applies several refinement techniques, including the removal of trivial augmented data, smoothing, consistency checks, and edge sharpening, ensuring the output depth map is of the highest quality and ready for application use.
+Following the fusion process, this module applies several refinement techniques, including the removal of trivial augmented data, smoothing, and consistency checks, ensuring the output depth map is of the highest quality and ready for application use.
 
